### 2020-01-30 paper/paper.tex
export paper from LaTeX

\documentclass[article, shortnames]{jss}

%% -- LaTeX packages and custom commands --------------

%% recommended packages
\usepackage{thumbpdf,lmodern}

%% another package (only for this demo article)
\usepackage{framed}
\usepackage{subcaption} %continuefloat
\usepackage{xfrac}
\usepackage{mathtools}
\usepackage{natbib}

\usepackage[acronym, xindy, toc, acronym, style=super,  nogroupskip, nonumberlist, nopostdot]{glossaries}
\setlength{\glsdescwidth}{0.75\textwidth}
\include{glossary}
\makeglossaries

\usepackage{placeins}

%\usepackage[colorlinks,
%           linkcolor={black},
%           citecolor={black},
%           urlcolor={blue!80!white}]{hyperref}
           %allbordercolors={0 0 0},
           %pdfborderstyle={/S/U/W 1}

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

%% - Article metainformation (author, title, ...) ----------

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{S. Thomas Kelly\\University of Otago
   \And Michael A. Black\\University of Otago}
\Plainauthor{Thomas Kelly, Michael Black}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{\pkg{graphsim}: An \proglang{R} package for simulating gene expression data from graph structures of biological pathways}
\Plaintitle{graphsim: An R package for simulating gene expression data from graph structures of biological pathways}
\Shorttitle{Simulating gene expression from pathways in R}

%% - \Abstract{} almost as usual
\Abstract{
Transcriptomic analysis is used to capture the molecular state of a cell or sample in many biological and medical applications. In addition to identifying alterations in activity at the level of individual genes, understanding changes in the gene networks that regulate fundamental biological mechanisms is also an important objective of molecular analysis. As a result, databases that describe biological pathways are increasingly relied on to assist with the interpretation of results from large-scale genomics studies. Incorporating information from biological pathways and gene regulatory networks into a genomic data analysis is a popular strategy, and there are many methods that provide this functionality for gene expression data. When developing or comparing such methods, it is important to gain an accurate assessment of their performance, with simulation-based validation studies a popular choice.  This necessitates the use of simulated data that correctly accounts for pathway relationships and correlations. Here we present a versatile statistical framework to simulate correlated gene expression data from biological pathways, by sampling from a multivariate normal distribution derived from a graph structure. This procedure has been released as the \pkg{graphsim} \proglang{R} package (\url{https://github.com/TomKellyGenetics/graphsim}) and is compatible with any graph structure that can be described using the \pkg{igraph} package.
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{gene-expression, simulation, genomics, pathway, network, \proglang{R}}
\Plainkeywords{gene-expression, simulation, genomics, pathway, network, R}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  S. Thomas Kelly\\
  Division of Genomic Medicine
  RIKEN Center for Integrative Medical Sciences\\
  Yokohama, Japan \\
  E-mail: \email{tom.kelly@riken.jp}\\
  \emph{and}\\
  Department of Biochemistry\\
  University of Otago\\
  Dunedin, New Zealand\\
  %E-mail: \email{tom.kelly@postgrad.otago.ac.nz}%\\
  %URL: \url{https://eeecon.uibk.ac.at/~zeileis/}

  Michael Black\\
  Department of Biochemistry\\
  University of Otago\\
  PO Box 56\\
  Dunedin, New Zealand\\
  E-mail: \email{mik.black@otago.ac.nz}%\\
  %URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
}

\begin{document}

%% - Introduction ---------------------

%% - In principle "as usual".
%% - But should typically have some discussion of both _software_ and _methods_.
%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
%% - If such markup is in (sub)section titles, a plain text version has to be
%%   added as well.
%% - All software mentioned should be properly \cite-d.
%% - All abbreviations should be introduced.
%% - Unless the expansions of abbreviations are proper names (like "Journal
%%   of Statistical Software" above) they should be in sentence case (like
%%   "generalized linear models" below).

\section[Introduction: inference and modelling of biological networks]{Introduction: inference and modelling of biological networks} \label{sec:intro}

\shortcites{Perou2000, Arner2015, Reactome, Komatsu2013, Yamaguchi2007, Shimamura2009, Hirose2008, Li2015, Genz2016}

Network analysis of molecular biological pathways has the potential to lead to new insights into biology and medical genetics \citep{Barabasi2004, Hu2016}. Since gene expression profiles capture a consistent signature of the regulatory state of a cell \citep{Perou2000, Ozsolak2011, Svensson2018}, they can be used to analyse complex molecular states with genome-scale data. However, biological pathways are often analysed in a reductionist paradigm as amorphous sets of genes involved in particular functions, despite the fact that the relationships defined by pathway structure could further inform gene expression analyses. In many cases, the pathway relationships are well-defined, experimentally-validated, and are available in public databases \citep{Reactome}. 
%While pathway analysis over-representation analysis using the hypergeometric test is an effective approach to identify enriched functional groups \citep{Mooney2015, Werner2008}, it is limited by the accuracy and coverage of the biological databases used. Furthermore, pathway enrichment does not utilise the pathway structure information that has been gathered for many pathways and is supported by databases such as Reactome \citep{Reactome}. 
As a result, network analysis techniques could play an important role in furthering our understanding of biological pathways and aiding in the interpretation of genomics studies. %Furthermore with modelling and simulation procedures to evaluate them are needed.

Gene networks provide insights into how cells are regulated, by mapping regulatory interactions between target genes and transcription factors, enhancers, and sites of epigenetic marks or chromatin structures \citep{Barabasi2004, Yamaguchi2007}. Inference of these regulatory interactions for genomics investigations has the potential to radically expand the range of candidate biological pathways to be further explored, or to improve the accuracy of bioinformatics and functional genomic analysis. %which relies on these pathways being curated.
A number of methods have already been developed to utilise timecourse gene expression data \citep{Arner2015, Yamaguchi2007} using gene regulatory modules in state-space models and recursive vector autoregressive models \citep{Hirose2008, Shimamura2009}. Various approaches to gene regulation and networks at the genome-wide scale have lead to novel biological insights \citep{Arner2015, Komatsu2013}. However, inference of regulatory networks has thus far relied on experimental validation or resampling-based approaches to estimate the likelihood of specific network modules being predicted \citep{Markowetz2007,Hawe2019}.

There is a need, therefore, for a systematic framework for statistical modelling and simulation of gene expression data derived from hypothetical, inferred or known gene networks. Here we present an \proglang{R} package to achieve this, where samples from a multivariate normal distribution are used to generate normally-distributed log-expression data, with correlations between genes derived from the structure of the underlying pathway or gene regulatory network. This methodology enables simulation of expression profiles that approximate the log-transformed and normalised data from microarray and bulk or single-cell RNA-Seq experiments. This procedure has been released as the \pkg{graphsim} \proglang{R} package to enable the generation of simulated gene expression datasets containing pathway relationships from a known underlying network. These simulated datasets can be used to evaluate various bioinformatics methodologies, including statistical and network inference procedures.

%This procedure enables modelling of genes within a biological pathway and the investigation of the impact of pathway structure on identification of biomarkers from gene expression analysis techniques. It was developed for biological pathway members with correlated gene expression (higher than the background of genes in other pathways) but it may also be applicable to modelling protein levels (e.g, in a kinase regulation cascade) or substrates and products (e.g., in a metabolic pathway). The simulation procedure was designed to use graph structures (such as those shown in Figure~\ref{fig:simple_graph}) to inform development of a ``Sigma'' variance-covariance matrix ($\Sigma$) for sampling from a multivariate normal distribution. % (using the \pkg{mvtnorm} \proglang{R} package).

\iffalse
\begin{leftbar}
The introduction is in principle ``as usual''. However, it should usually embed
both the implemented \emph{methods} and the \emph{software} into the respective
relevant literature. For the latter both competing and complementary software
should be discussed (within the same software environment and beyond), bringing
out relative (dis)advantages. All software mentioned should be properly
\verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
\textsc{Bib}{\TeX}.)

For writing about software JSS requires authors to use the markup
\verb|\proglang{}| (programming languages and large programmable systems),
\verb|\pkg{}| (software packages), \verb|\code{}| (functions, commands,
arguments, etc.). If there is such markup in (sub)section titles (as above), a
plain text version has to be provided in the {\LaTeX} command as well. Below we
also illustrate how abbreviations should be introduced and citation commands can
be employed. See the {\LaTeX} code for more details.
\end{leftbar}

Modeling count variables is a common task in economics and the social sciences.
The classical Poisson regression model for count data is often of limited use in
these disciplines because empirical count data sets typically exhibit
overdispersion and/or an excess number of zeros. The former issue can be
addressed by extending  the plain Poisson regression model in various
directions: e.g., using sandwich covariances or estimating an additional
dispersion parameter (in a so-called quasi-Poisson model). Another more formal
way is to use a negative binomial (NB) regression. All of these models belong to
the family of generalized linear models (GLMs). However, although these models
typically can capture overdispersion rather well, they are in many applications
not sufficient for  modeling excess zeros. Since \citep{Mullahy:1986} there is
increased interest in zero-augmented models that address this issue by a second
model component capturing zero counts. An overview of count data models in
econometrics, including  hurdle and zero-inflated models, is provided in
\citep{Cameron+Trivedi:2013}.

In \proglang{R} \citep{R}, GLMs are provided by the model fitting functions
\fct{glm} in the \pkg{stats} package and \fct{glm.nb} in the \pkg{MASS} package
\citep[][Chapter~7.4]{Venables+Ripley:2002} along with associated methods for
diagnostics and inference. The manuscript that this document is based on
\citep{Zeileis+Kleiber+Jackman:2008} then introduced hurdle and zero-inflated
count models in the functions \fct{hurdle} and \fct{zeroinfl} in the \pkg{pscl}
package \citep{Jackman:2015}. Of course, much more software could be discussed
here, including (but not limited to) generalized additive models for count data
as available in the \proglang{R} packages \pkg{mgcv} \citep{Wood:2006}, 
\pkg{gamlss} \citep{Stasinopoulos+Rigby:2007}, or \pkg{VGAM} \citep{Yee:2009}.
\fi

%% - Manuscript ---------------------

%% - In principle "as usual" again.
%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
%%   avoid empty lines before and after the equation (which would signal a new
%%   paragraph.
%% - When describing longer chunks of code that are _not_ meant for execution
%%   (e.g., a function synopsis or list of arguments), the environment {Code}
%%   is recommended. Alternatively, a plain {verbatim} can also be used.
%%   (For executed code see the next section.)

\section{Methodology and software} \label{sec:methods}

\begin{figure*}[!htb]
%\begin{mdframed}
%  \resizebox{\textwidth}{!}{
         \begin{center}
%
        \subcaptionbox{Activating pathway structure  \label{fig:simple_graph:first}}{%
            \fbox{
            \includegraphics[width=0.45\textwidth]{{simple_graph.png}}
            }
        }%
        \subcaptionbox{Pathway structure with inhibitions \label{fig:simple_graph:second}}{%
            \fbox{
           \includegraphics[width=0.45\textwidth]{{simple_graph_inhibiting.png}}
           }
        }%
%
    \end{center}
   \caption[Simulated graph structures]{\small \textbf{\textbf{Simulated graph structures.}} A constructed graph structure used as an example to demonstrate the simulation procedure in Figures~\ref{fig:simulation_activating} and~\ref{fig:simulation_inhibiting}. Activating links are denoted by black arrows and inhibiting links by red edges. Inhibiting edges have been highlighted in red.}
%}
\label{fig:simple_graph}
%\end{mdframed}
\end{figure*}

Here we present a procedure to simulate gene expression data with correlation structure derived from a known graph structure.  This procedure assumes that transcriptomic data have been generated and follow a log-normal distribution (i.e., $log(X_{ij}) \sim MVN({\bf\mu}, \Sigma)$, where ${\bf\mu}$ and $\Sigma$ are the mean vector  and variance-covariance matrix respectively, for gene expression data derived from a biological pathway) after appropriate normalisation \citep{Law2014, Li2015}. Log-normality of gene expression matches the assumptions of the popular \pkg{limma} package, which is often used for the analysis of intensity-based data from gene expression microarray studies and count-based data from RNA-Seq experiments. This approach has also been applied for modelling UMI-based count data from single-cell RNA-Seq experiments in the \pkg{DESCEND} package \citep{Wang2018}.
%rather than correlated blocks for $\Sigma$ in \pkg{mvtnorm} \citet{mvtnorm}
%This procedure produces a ``Sigma'' variance-covariance matrix ($\Sigma$) for sampling from a multivariate normal distribution (using the \pkg{mvtnorm} \proglang{R} package).

In order to simulate transcriptomic data, a pathway is first constructed as a graph structure, using the \pkg{igraph} \proglang{R} package \citep{igraph}, with the status of the edge relationships defined (i.e, whether they activate or inhibit downstream pathway members). 
%
\textcolor{black}{This procedure uses} a graph structure such as that presented in Figure~\ref{fig:simple_graph:first}. The graph can be defined by an adjacency matrix, \textbf{$A$} (with elements \textcolor{black}{$A_{ij}$}), where  \textcolor{black}{
%
\[
A_{ij} = 
\begin{dcases*}
   1                         & if genes $i$ and $j$ are adjacent \\
   0                         & otherwise
\end{dcases*}
\]}
%
\noindent
A matrix, \textbf{$R$}, with elements \textcolor{black}{$R_{ij}$}, is calculated based on distance (i.e., number of edges contained in the shortest path) between nodes, such that closer nodes are given more weight than more distant nodes, to define inter-node relationships. A geometrically-decreasing (relative) distance weighting is used to achieve this: \textcolor{black}{
%
\[
R_{ij} = 
\begin{dcases*}
   1                            & if genes $i$ and $j$ are adjacent \\
   (\sfrac{1}{2})^{d_{ij}}      & if a path can be found  between genes $i$ and $j$ \\
   0                            & if no path exists between genes $i$ and $j$ 
\end{dcases*}
\]}
%
\noindent
where $d_{ij}$ is the length of the shortest path (i.e., minimum number of edges traversed) between genes (nodes) $i$ and $j$ in graph $G$. Each more distant node is thus related by $\sfrac{1}{2}$ compared to the next nearest, as shown in Figure~\ref{fig:simulation_activating:second}.
An arithmetically-decreasing (absolute) distance weighting is also supported in the \pkg{graphsim} \proglang{R} package which implements this procedure: \textcolor{black}{
%
\[
R_{ij} = 
\begin{dcases*}
   1                            & if genes $i$ and $j$ are adjacent \\
   1-\frac{d_{ij}}{diam(G)}     & if a path can be found  between genes $i$ and $j$ \\
   0                            & if no path exists between genes $i$ and $j$ 
\end{dcases*}
\]
%
}

Assuming a unit variance for each gene, these values can be used to derive a $\Sigma$ matrix:
%
\[
\Sigma_{ij} = 
\begin{dcases*}
   1                            & if $i=j$ \\
   \rho R_{ij}  & otherwise
\end{dcases*}
\]
%
\noindent
where $\rho$ is the correlation between adjacent nodes. 
%The $\Sigma$ matrix (with a diagonal of $1$) would thus have a standard deviation of 1 for each node. 
Thus covariances between adjacent nodes are assigned by a correlation parameter ($\rho$) and the remaining off-diagonal values in the matrix are based on scaling these correlations by the geometrically weighted relationship matrix (or the nearest positive definite matrix for $\Sigma$ with negative correlations).\\

Computing the nearest positive definite matrix is necessary to ensure that the variance-covariance matrix could be inverted when used as a parameter in multivariate normal simulations, particularly when negative correlations are included for inhibitions (as shown below). Matrices that could not be inverted occurred rarely with biologically plausible graph structures but this approach allows for the computation of a plausible correlation matrix when the graph structure given is incomplete or contains loops. When required, the nearest positive definite matrix is computed using the \texttt{nearPD} function of the \pkg{Matrix} \proglang{R} package \citep{Matrix} to perform Higham's algorithm \citep{Higham2002} on variance-covariance matrices. The \pkg{graphsim} package gives a warning when this occurs.

\section{Illustrations} \label{sec:illustrations}
\subsection{Generating a Graph Structure} \label{sec:plot_graph}

The graph structure in Figure~\ref{fig:simple_graph:first} was used to simulate correlated gene expression data by sampling from a multivariate normal distribution using the \pkg{mvtnorm} \proglang{R} package \citep{Genz2009, mvtnorm}.
%This simulation procedure will be demonstrated with the relatively simple constructed graph structure shown in Figure~\ref{fig:simple_graph}. 
The graph structure visualisation in Figure~\ref{fig:simple_graph} was specifically developed for (directed) iGraph objects in \proglang{R} and is available in the \pkg{plot.igraph} and \pkg{igraph.extensions} packages. The \texttt{plot\_directed} function enables customisation of plot parameters for each node or edge, and mixed (directed) edge types for indicating activation or inhibition. These inhibition links (which occur frequently in biological pathways) are demonstrated in Figure~\ref{fig:simple_graph:second}.

A graph structure can be generated and plotted using the following commands in R:

\begin{CodeChunk}
\begin{CodeInput}
#install packages required (once per machine)
install.packages("igraph")
install.packages("devtools")
library("devtools")
devtools::install\_github("TomKellyGenetics/graphsim")
#install custom plotting package
devtools::install\_github("TomKellyGenetics/plot.igraph")
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
#load required packages (once per R instance)
library("igraph")
library("graphsim")
library("plot.igraph")
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
#generate graph structure
graph\_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"), c("D", "E"), 
    c("D", "F"), c("F", "G"), c("F", "I"), c("H", "I"))
graph <- graph.edgelist(graph\_edges, directed = T)

#plot graph structure (Figure 1)
plot\_directed(graph, state = "activating", layout = layout.kamada.kawai, 
    cex.node=3, cex.arrow=5, arrow\_clip = 0.2)
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
#generate parameters for inhibitions
state <-  c(1, 1, -1, 1, 1, 1, 1, -1, 1)

#plot graph structure with inhibitions (Figure 2)
plot\_directed(graph, state=state,  layout = layout.kamada.kawai, 
    cex.node=3, cex.arrow=5, arrow\_clip = 0.2)
\end{CodeInput}
\end{CodeChunk}

\subsection{Generating a Simulated Expression Dataset} \label{sec:graphsim_demo}

\begin{figure*}[!hp]
%\begin{mdframed}
%  \resizebox{\textwidth}{!}{
         \begin{center}
%
        \subcaptionbox{Activating pathway structure \label{fig:simulation_activating:first}}{%
            \includegraphics[width=0.35\textwidth]{{simple_graph.png}}
        }%
        \subcaptionbox{Relationship matrix \label{fig:simulation_activating:second}}{%
            \includegraphics[width=0.35\textwidth]{{dist_mat.png}}
        }%

        \subcaptionbox{$\Sigma$ (covariance matrix) \label{fig:simulation_activating:third}}{%
           \includegraphics[width=0.35\textwidth]{{sigma_mat.png}}
        }%
	\subcaptionbox{Simulated correlation\label{fig:simulation_activating:fifth}}{%
           \includegraphics[width=0.35\textwidth]{{expr_cor_mat.png}}
        }%

	\subcaptionbox{Simulated expression data (log scale) \label{fig:simulation_activating:fourth}}{%
            \includegraphics[width=1\textwidth]{{expr_mat_wide.png}}
        }%
        %\subcaptionbox{Simulated gene function calls \label{fig:simulation_activating:sixth}}{%
        %   \includegraphics[width=0.35\textwidth]{{expr_disc_mat.png}}
        %}%
    \end{center}
   \caption[Simulating expression from a graph structure]{\small \textbf{\textbf{Simulating expression from a graph structure.}} An example of a graph structure (a) that has been used to derive a relationship matrix (b), $\Sigma$ matrix (c) and correlation structure (d) from the relative distances between the nodes. Non-negative values are coloured white to red from $0$ to $1$. This $\Sigma$ matrix has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution. Here genes with closer relationships in the pathway structure show higher correlation between simulated values.}
%}
\label{fig:simulation_activating}
%\end{mdframed}
\end{figure*}

The correlation parameter of $\rho = 0.8$ is used to demonstrate the inter-correlated datasets using a geometrically-generated relationship matrix (as used for the example in Figure~\ref{fig:simulation_activating:third}). This $\Sigma$ matrix was then used to sample from a multivariate normal distribution such that each gene had a mean of $0$, standard deviation $1$, and covariance within the range $[0,1]$ so that the off-diagonal elements of $\Sigma$ represent correlations. This procedure generated a simulated (continuous normally-distributed) log-expression profile for each node (Figure~\ref{fig:simulation_activating:fourth}) with a corresponding correlation structure (Figure~\ref{fig:simulation_activating:fifth}). The simulated correlation structure closely resembled the expected correlation structure ($\Sigma$ in Figure~\ref{fig:simulation_activating:third}) even for the relatively modest sample size ($N=100$) illustrated in Figure~\ref{fig:simulation_activating}. Once a gene expression dataset comprising multiple pathways has been generated (as in Figure~\ref{fig:simulation_activating:fourth}), 
%then a discrete matrix of gene function was constructed with a functional threshold quantile to simulate functional relationships of synthetic lethality (Figure~\ref{fig:SL_Model_Expression}). %Throughout this thesis, this threshold is the 0.3 quantile (as discussed in Section~\ref{methods:SL_Model}) which generates functional discrete matrices such as those used for synthetic lethal simulation in Section~\ref{methods:simulating_SL} (as shown Figure~\ref{fig:simulation_activating:sixth}).
it can then be used to test procedures designed for analysis of empirical gene expression data (such as those generated by microarrays or RNA-Seq) that have been normalised on a log-scale.

The simulated dataset can be generated using the following \proglang{R} code:

\begin{CodeChunk}
\begin{CodeInput}
#adjacency matrix
adj\_mat <- make\_adjmatrix\_graph(graph)

#relationship matrix
dist\_mat <- make\_distance\_graph(graph\_test4, absolute = F)

#sigma matrix directly from graph
sigma\_mat <- make\_sigma\_mat\_dist\_graph(graph, 0.8, absolute = F)

#show shortest paths of graph
shortest\_paths <- shortest.paths(graph)

#generate expression data directly from graph
expr <- generate\_expression(100, graph, cor = 0.8, mean = 0, comm = F,
                            dist = T, absolute = F, state = state)
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
##plot steps

#plot adjacency matrix
heatmap.2(make\_adjmatrix\_graph(graph), scale = "none", trace = "none", 
          col = colorpanel(3, "grey75", "white", "blue"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))

#plot relationship matrix
heatmap.2(make\_distance\_graph(graph\_test4, absolute = F), scale = "none",
          trace = "none", col = bluered(50), colsep = 1:length(V(graph)), 
        rowsep = 1:length(V(graph)))

#plot sigma matrix
heatmap.2(make\_sigma\_mat\_dist\_graph(graph, 0.8, absolute = F), scale = "none", 
          trace = "none", col = bluered(50), colsep = 1:length(V(graph)), 
          rowsep = 1:length(V(graph)))
expr <- generate\_expression(100, graph, cor = 0.8, mean = 0, comm = F, dist = T,
                            absolute = F, state = state)

#plot simulated expression data
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50), 
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))

#plot simulated correlations
heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = bluered(50), 
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
\end{CodeInput}
\end{CodeChunk}

\begin{figure*}[!hp]
%\begin{mdframed}
%  \resizebox{\textwidth}{!}{
         \begin{center}
%
        \subcaptionbox{Inhibiting pathway structure\label{fig:simulation_inhibiting:first}}{%
            \includegraphics[width=0.35\textwidth]{{simple_graph_inhibiting.png}}
        }%
        \subcaptionbox{Relationship matrix \label{fig:simulation_inhibiting:second}}{%
            \includegraphics[width=0.35\textwidth]{{dist_mat.png}}
        }%

        \subcaptionbox{$\Sigma$ (covariance matrix) \label{fig:simulation_inhibiting:third}}{%
           \includegraphics[width=0.35\textwidth]{{sigma_mat_inhibiting.png}}
        }%
	\subcaptionbox{Simulated correlation\label{fig:simulation_inhibiting:fifth}}{%
           \includegraphics[width=0.35\textwidth]{{expr_cor_mat_inhibiting.png}}
        }%

        \subcaptionbox{Simulated expression data (log scale) \label{fig:simulation_inhibiting:fourth}}{%

            \includegraphics[width=1\textwidth]{{expr_mat_inhibiting_wide.png}}
        }%
        %\subcaptionbox{Simulated gene function calls \label{fig:simulation_inhibiting:sixth}}{%
        %   \includegraphics[width=0.35\textwidth]{{expr_inhib_disc_mat.png}}
        %}%
    \end{center}
    \caption[Simulating expression from graph structure with inhibitions]{\small \textbf{\textbf{Simulating expression from graph structure with inhibitions.}} An example of a graph structure (a), that has been used to derive a relationship matrix (b), $\Sigma$ matrix (c), and correlation structure (d), from the relative distances between the nodes. These values are coloured blue to red from $-1$ to $1$. This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution. Here the inhibitory relationships between genes are reflected in negatively correlated simulated  values.}
%}
\label{fig:simulation_inhibiting}
%\end{mdframed}
\end{figure*}

The simulation procedure (Figure~\ref{fig:simulation_activating}) can similarly be used for pathways containing inhibitory links (Figure~\ref{fig:simulation_inhibiting}) with several refinements. With the inhibitory links (Figure~\ref{fig:simulation_inhibiting:first}), distances are calculated in the same manner as before (Figure~\ref{fig:simulation_inhibiting:second}) with inhibitions accounted for by iteratively multiplying downstream nodes by $-1$ to form modules with negative correlations between them (Figures~\ref{fig:simulation_inhibiting:third} and~\ref{fig:simulation_inhibiting:fifth}). A multivariate normal distribution with these negative correlations can be sampled to generate simulated data (Figure~\ref{fig:simulation_inhibiting:fourth}).% and~\ref{fig:simulation_inhibiting:sixth}).

\begin{figure*}[!hp]
%\begin{mdframed}
%  \resizebox{\textwidth}{!}{
         \begin{center}
%
        \subcaptionbox{TGF-$\beta$ activates SMADs\label{fig:simulation_smad:first}}{%
            \includegraphics[width=0.35\textwidth]{{graph_smad_simple.pdf}}
        }%
        \subcaptionbox{Relationship matrix \label{fig:simulation_smad:second}}{%
            \includegraphics[width=0.35\textwidth]{{dist_mat_smad_simple.png}}
        }%

        \subcaptionbox{$\Sigma$ (covariance matrix) \label{fig:simulation_smad:third}}{%
           \includegraphics[width=0.35\textwidth]{{sigma_mat_inhibiting_smad_simple.png}}
        }%
	\subcaptionbox{Simulated correlation\label{fig:simulation_smad:fifth}}{%
           \includegraphics[width=0.35\textwidth]{{expr_cor_mat_inhibiting_smad_simple.png}}
        }%

        \subcaptionbox{Simulated expression data (log scale) \label{fig:simulation_smad:fourth}}{%

            \includegraphics[width=0.8\textwidth]{{expr_mat_inhibiting_wide_smad_simple.png}}
        }%
        %\subcaptionbox{Simulated gene function calls \label{fig:simulation_smad:sixth}}{%
        %   \includegraphics[width=0.35\textwidth]{{expr_inhib_disc_mat.png}}
        %}%
    \end{center}
   \caption[Simulating expression from a biological pathway graph structure]{\small \textbf{\textbf{Simulating expression from graph structure with inhibitions.}} The graph structure (a) of a known biological pathway, the TGF-$\beta$ receptor signaling activates SMADs (R-HSA-2173789), was used to derive a relationship matrix (b), $\Sigma$ matrix (c) and correlation structure (d) from the relative distances between the nodes. These values are coloured blue to red from $-1$ to $1$. This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution. Here modules of genes with correlated expression can be clearly discerned.}
%}
\label{fig:simulation_smad}
%\end{mdframed}
\end{figure*}

The simulation procedure is also demonstrated here (Figure~\ref{fig:simulation_smad}) on a pathway structure for a known biological pathway (from reactome R-HSA-2173789) of TGF-$\beta$ receptor signaling activates SMADs (Figure~\ref{fig:simulation_smad:first}) derived from the Reactome database version 52 \citep{Reactome}. Distances are calculated in the same manner as before (Figure~\ref{fig:simulation_smad:second}) producing blocks of correlated genes (Figures~\ref{fig:simulation_inhibiting:third} and~\ref{fig:simulation_inhibiting:fifth}). This shows that multivariate normal distribution can be sampled to generate simulated data to represent expression with the complexity of a biological pathway (Figure~\ref{fig:simulation_inhibiting:fourth}). Here \textit{SMAD7} exhibits negative correlations with the other SMADs consistent with it's functions as as an ``inhibitor SMAD'' with competitively inhibits \textit{SMAD4}. % and~\ref{fig:simulation_inhibiting:sixth}).

\iffalse
These simulated datasets could then be used for simulating synthetic lethal partners of a query gene within a graph network. The query gene was assumed to be separate from the graph network pathway and was added to the dataset using the procedure in Section~\ref{methods:simulating_SL}. Thus I can simulate known synthetic lethal partner genes within a synthetic lethal partner pathway structure.
\fi

\iffalse

	\includegraphics{{dist_mat.png}}
	\includegraphics{{sigma_mat.png}}
		\includegraphics{{expr_mat.png}}
			\includegraphics{{expr_cor_mat.png}}
		\includegraphics{{expr_disc_mat.png}}

	\includegraphics{{state_matrix_inhibiting.png}}
	\includegraphics{{dist_mat.png}}
		\includegraphics{{sigma_mat_inhibiting.png}}
		\includegraphics{{expr_inhib_mat.png}}
			\includegraphics{{expr_inhib_cor_mat.png}}
		\includegraphics{{expr_inhib_disc_mat.png}}

\fi

%\FloatBarrier

%% - Illustrations --------------------

%% - Virtually all JSS manuscripts list source code along with the generated
%%   output. The style files provide dedicated environments for this.
%% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
%%   or knitr using the render_sweave() hook - are used (without the need to
%%   load Sweave.sty).
%% - Equivalently, {CodeInput} and {CodeOutput} can be used.
%% - The code input should use "the usual" command prompt in the respective
%%   software system.
%% - For \proglang{R} code, the prompt "R> " should be used with "+  " as the
%%   continuation prompt.
%% - Comments within the code chunks should be avoided - these should be made
%%   within the regular LaTeX text.

\iffalse

\section{Illustrations} \label{sec:illustrations}

For a simple illustration of basic Poisson and NB count regression the
\code{quine} data from the \pkg{MASS} package is used. This provides the number
of \code{Days} that children were absent from school in Australia in a
particular year, along with several covariates that can be employed as regressors.
The data can be loaded by
%
\begin{CodeChunk}
\begin{CodeInput}
R> data("quine", package = "MASS")
\end{CodeInput}
\end{CodeChunk}
%
and a basic frequency distribution of the response variable is displayed in
Figure~\ref{fig:quine}.

\begin{leftbar}
For code input and output, the style files provide dedicated environments.
Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
hook. Please make sure that all code is properly spaced, e.g., using
\code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
use ``the usual'' command prompt in the respective software system. For
\proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
the continuation prompt. Generally, comments within the code chunks should be
avoided - and made in the regular {\LaTeX} text instead. Finally, empty lines
before and after code input/output should be avoided (see above).
\end{leftbar}

\begin{figure}[t!]
\centering
\includegraphics{article-visualization}
\caption{\label{fig:quine} Frequency distribution for number of days absent
from school.}
\end{figure}

As a first model for the \code{quine} data, we fit the basic Poisson regression
model. (Note that JSS prefers when the second line of code is indented by two
spaces.)
%
\begin{CodeChunk}
\begin{CodeInput}
R> m\_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
+    family = poisson)
\end{CodeInput}
\end{CodeChunk}
%
To account for potential overdispersion we also consider a negative binomial
GLM.
%
\begin{CodeChunk}
\begin{CodeInput}
R> library("MASS")
R> m\_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
\end{CodeInput}
\end{CodeChunk}
%
In a comparison with the BIC the latter model is clearly preferred.
%
\begin{CodeChunk}
\begin{CodeInput}
R> BIC(m\_pois, m\_nbin)
\end{CodeInput}
\begin{CodeOutput}
       df      BIC
m\_pois 18 2046.851
m\_nbin 19 1157.235
\end{CodeOutput}
\end{CodeChunk}
%
Hence, the full summary of that model is shown below.
%
\begin{CodeChunk}
\begin{CodeInput}
R> summary(m\_nbin)
\end{CodeInput}
\begin{CodeOutput}
Call:
glm.nb(formula = Days ~ (Eth + Sex + Age + Lrn)^2, data = quine, 
    init.theta = 1.60364105, link = log)

Deviance Residuals: 
    Min       1Q   Median       3Q      Max  
-3.0857  -0.8306  -0.2620   0.4282   2.0898

Coefficients: (1 not defined because of singularities)
            Estimate Std. Error z value Pr(>|z|)    
(Intercept)  3.00155    0.33709   8.904  < 2e-16 ***
EthN        -0.24591    0.39135  -0.628  0.52977    
SexM        -0.77181    0.38021  -2.030  0.04236 *  
AgeF1       -0.02546    0.41615  -0.061  0.95121    
AgeF2       -0.54884    0.54393  -1.009  0.31296    
AgeF3       -0.25735    0.40558  -0.635  0.52574    
LrnSL        0.38919    0.48421   0.804  0.42153    
EthN:SexM    0.36240    0.29430   1.231  0.21818    
EthN:AgeF1  -0.70000    0.43646  -1.604  0.10876    
EthN:AgeF2  -1.23283    0.42962  -2.870  0.00411 ** 
EthN:AgeF3   0.04721    0.44883   0.105  0.91622    
EthN:LrnSL   0.06847    0.34040   0.201  0.84059    
SexM:AgeF1   0.02257    0.47360   0.048  0.96198    
SexM:AgeF2   1.55330    0.51325   3.026  0.00247 ** 
SexM:AgeF3   1.25227    0.45539   2.750  0.00596 ** 
SexM:LrnSL   0.07187    0.40805   0.176  0.86019    
AgeF1:LrnSL -0.43101    0.47948  -0.899  0.36870    
AgeF2:LrnSL  0.52074    0.48567   1.072  0.28363    
AgeF3:LrnSL       NA         NA      NA       NA    
-
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

(Dispersion parameter for Negative Binomial(1.6036) family taken to be 1)

    Null deviance: 235.23  on 145  degrees of freedom
Residual deviance: 167.53  on 128  degrees of freedom
AIC: 1100.5

Number of Fisher Scoring iterations: 1

              Theta:  1.604 
          Std. Err.:  0.214

 2 x log-likelihood:  -1062.546 
\end{CodeOutput}
\end{CodeChunk}
\fi

%% - Summary/conclusions/discussion ---------------

\FloatBarrier

\section{Summary and discussion} \label{sec:summary}

\begin{leftbar}
Biological pathways are of fundamental importance to understanding molecular biology. In order to translate findings from genomics studies into real-world applications such as improved healthcare, the roles of genes must be studied in the context of molecular pathways. Here we present a statistical framework to simulate gene expression from biological pathways, and provide the \pkg{graphsim} package in \proglang{R} to generate these simulated datasets. This approach is versatile and can be fine-tuned for modelling existing biological pathways or for testing whether constructed pathways can be detected by other means. In particular, methods to infer biological pathways and gene regulatory networks from gene expression data can be tested on simulated datasets using this framework. The package also enables simulation of complex gene expression datasets to test how these pathways impact on statistical analysis of gene expression data using existing methods or novel statistical methods being developed for gene expression data analysis. 
\end{leftbar}

%% - Optional special unnumbered sections -------------

\section*{Computational details}

\iffalse
\begin{leftbar}
If necessary or useful, information about certain computational details
such as version numbers, operating systems, or compilers could be included
in an unnumbered section. Also, auxiliary packages (say, for visualizations,
maps, tables, \dots) that are not cited in the main text can be credited here.
\end{leftbar}
\fi

The results in this paper were obtained using
\proglang{R}~3.6.1 with the
\pkg{igraph}~1.2.4.1 \pkg{Matrix}~1.2-17, \pkg{matrixcalc}~1.0-3, and \pkg{mvtnorm}~1.0-11 packages. \proglang{R} itself
and all dependent packages used are available from the Comprehensive
\proglang{R} Archive Network (CRAN) at
\url{https://CRAN.R-project.org/}. 
The \pkg{graphsim} and \pkg{plot.igraph} packages presented can be installed from \url{https://github.com/TomKellyGenetics/graphsim} 
and \url{https://github.com/TomKellyGenetics/plot.igraph} respectively. These functions can also be installed using the \pkg{igraph.extensions} library at \url{https://github.com/TomKellyGenetics/igraph.extensions} which includes other plotting functions used. This software is cross-platform and compatible with \proglang{R} installations on Windows, Mac, and Linux operating systems. The package GitHub repository also contains Vignettes with more information and examples on running functions released in the \proglang{R} package. The package (\pkg{graphsim}~0.1.0) meets CRAN submission criteria and will be released.

\section*{Acknowledgements} %change to British spelling from American: ``Acknowledgments''

\begin{leftbar}
This package was developed as part of a PhD research project funded by the Postgraduate Tassell Scholarship in Cancer Research Scholarship awarded to STK. We thank members of the Laboratory of Professor Satoru Miyano at the University of Tokyo, Institute for Medical Science, Professor Seiya Imoto, Associate Professor Rui Yamaguchi, and Dr Paul Sheridan (Assistant Professor at Hirosaki University,CSO at Tupac Bio) for helpful discussions in this field. We also thank Professor Parry Guilford at the University of Otago, Professor Cristin Print at the University of Auckland, and Dr Erik Arner at the RIKEN Center for Integrative Medical Sciences for their excellent advice during this project.
\end{leftbar}

%% - Bibliography ---------------------
%% - References need to be provided in a .bib BibTeX database.
%% - All references should be made with \cite, \citet, \cite, \citealp etc.
%%   (and never hard-coded). See the FAQ for details.
%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
%% - Titles in the .bib should be in title case.
%% - DOIs should be included where available.

\bibliography{refs}

%% - Appendix (if any) -------------------
%% - After the bibliography with page break.
%% - With proper section titles and _not_ just "Appendix".

\newpage

%\begin{appendix}

%\section{More technical details} \label{app:technical}

\begin{leftbar}
%Would it be appropriate to include the Vignette for the package here?
\iffalse
Appendices can be included after the bibliography (with a page break). Each
section within the appendix should have a proper section title (rather than
just \emph{Appendix}).

For more technical style details, please check out JSS's style FAQ at
\url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
which includes the following topics:
\begin{itemize}
  \item Title vs.\ sentence case.
  \item Graphics formatting.
  \item Naming conventions.
  \item Turning JSS manuscripts into \proglang{R} package vignettes.
  \item Trouble shooting.
  \item Many other potentially helpful details\dots
\end{itemize}
\fi
\end{leftbar}

\iffalse
\section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}

\begin{leftbar}
References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
\verb|\citealp| etc.\ (and never hard-coded). This commands yield different
formats of author-year citations and allow to include additional details (e.g.,
pages, chapters, \dots) in brackets. In case you are not familiar with these
commands see the JSS style FAQ for details.

Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task - especially
when acquiring the entries automatically from mixed online sources. However,
it is important that informations are complete and presented in a consistent
style to avoid confusions. JSS requires the following format.
\begin{itemize}
  \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
    be used in the references.
  \item Titles should be in title case.
  \item Journal titles should not be abbreviated and in title case.
  \item DOIs should be included where available.
  \item Software should be properly cited as well. For \proglang{R} packages
    \code{citation("pkgname")} typically provides a good starting point.
\end{itemize}
\end{leftbar}
\fi

%\end{appendix}

%% --------------------------

\end{document}

### 2020-03-02 tests/testthat/test-cpp.R
expand unit tests

@@ -1,0 +1,4 @@
+context("C++")
+test_that("Catch unit tests pass", {
+    expect_cpp_tests_pass("graphsim")
+})

### 2020-03-02 tests/testthat/test_adjmatrix.R

@@ -12,4 +12,5 @@ test_that("Generate adjacency matrix from graph structure", {
   expect_equal(ncol(adjacency_matrix1), length(V(graph_test1)))
   expect_equal(sum(adjacency_matrix1), length(E(graph_test1))*2)
   expect_equal(all(is.matrix(adjacency_matrix1)), TRUE)
+  expect_true(all(adjacency_matrix1 == cbind(c(0, 1, 0, 0), c(1, 0, 1, 1), c(0, 1, 0, 0), c(0, 1, 0, 0))))
 })

### 2020-03-02 tests/testthat/test_commonlink.R

@@ -13,10 +13,10 @@ test_that("Generate common link matrix from adjacency matrix", {
   expect_equal(nrow(common_link_matrix1), length(V(graph_test1)))
   expect_equal(ncol(common_link_matrix1), length(V(graph_test1)))
   expect_equal(sum(common_link_matrix1), length(E(graph_test1))*4)
+  expect_true(all(adjacency_matrix1 == cbind(c(0, 1, 0, 0), c(1, 0, 1, 1), c(0, 1, 0, 0), c(0, 1, 0, 0))))
   expect_equal(all(is.matrix(common_link_matrix1)), TRUE)
+  expect_true(all(common_link_matrix1 == cbind(c(1, 0, 1, 1), c(0, 3, 0, 0), c(1, 0, 1, 1), c(1, 0, 1, 1))))
 })
-
-
 
 test_that("Generate common link matrix from graph structure", {
   graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
@@ -29,4 +29,5 @@ test_that("Generate common link matrix from graph structure", {
   expect_equal(ncol(common_link_matrix1), length(V(graph_test1)))
   expect_equal(sum(common_link_matrix1), length(E(graph_test1))*4)
   expect_equal(all(is.matrix(common_link_matrix1)), TRUE)
+  expect_true(all(common_link_matrix1 == cbind(c(1, 0, 1, 1), c(0, 3, 0, 0), c(1, 0, 1, 1), c(1, 0, 1, 1))))
 })

### 2020-03-02 tests/testthat/test_distance.R

@@ -1,0 +1,52 @@
+library("graphsim")
+library("igraph")
+context("Make Distance Matrix")
+
+test_that("Generate distance relation matrix from adjacency matrix", {
+  graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
+  graph_test1 <- graph.edgelist(graph_test1_edges, directed = T)
+  adjacency_matrix1 <- make_adjmatrix_graph(graph_test1)
+  distance_matrix1 <- make_distance_adjmat(adjacency_matrix1)
+  expect_equal(isSymmetric(distance_matrix1), TRUE)
+  expect_true(all(diag(distance_matrix1) == rep(1, nrow(adjacency_matrix1))))
+  expect_equal(nrow(distance_matrix1), length(V(graph_test1)))
+  expect_equal(ncol(distance_matrix1), length(V(graph_test1)))
+  expect_equal(sum(distance_matrix1), 8.5)
+  expect_true(all(adjacency_matrix1 == cbind(c(0, 1, 0, 0), c(1, 0, 1, 1), c(0, 1, 0, 0), c(0, 1, 0, 0))))
+  expect_true(all(distance_matrix1 == 1/cbind(c(1, 2, 4, 4), c(2, 1, 2, 2), c(4, 2, 1, 4), c(4, 2, 4, 1))))
+  expect_true(all(distance_matrix1 %in% c(0, 1/2^c(0:100))))
+})
+
+test_that("Generate distance relation matrix from commonlink matrix", {
+  graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
+  graph_test1 <- graph.edgelist(graph_test1_edges, directed = T)
+  adjacency_matrix1 <- make_adjmatrix_graph(graph_test1)
+  common_link_matrix1 <- make_commonlink_adjmat(adjacency_matrix1)
+  distance_matrix1 <- make_distance_adjmat(common_link_matrix1)
+  expect_equal(isSymmetric(distance_matrix1), TRUE)
+  expect_true(all(diag(distance_matrix1) == rep(1, nrow(adjacency_matrix1))))
+  expect_equal(nrow(distance_matrix1), length(V(graph_test1)))
+  expect_equal(ncol(distance_matrix1), length(V(graph_test1)))
[…]

### 2020-03-02 tests/testthat/test_generate.R

@@ -1,0 +1,55 @@
+library("graphsim")
+library("igraph")
+context("Generate Expression Data")
+
+test_that("Generate Expression matrix from graph object", {
+  graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
+  graph_test1 <- graph.edgelist(graph_test1_edges, directed = T)
+  adjacency_matrix1 <- make_adjmatrix_graph(graph_test1)
+  n <- 100
+  suppressWarnings(expression_matrix1 <- generate_expression(n, graph_test1, cor = 0.8))
+  expect_equal(nrow(expression_matrix1), length(V(graph_test1)))
+  expect_equal(ncol(expression_matrix1), n)
+  expect_equal(round(mean(expression_matrix1), 0), 0)
+  expect_equal(round(sd(expression_matrix1), 0), 1)
+  expect_true(is.matrix(expression_matrix1))
+  })
+
+test_that("Generate Expression matrix from graph object by commonlinks", {
+  graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
+  graph_test1 <- graph.edgelist(graph_test1_edges, directed = T)
+  adjacency_matrix1 <- make_adjmatrix_graph(graph_test1)
+  n <- 100
+  suppressWarnings(expression_matrix1 <- generate_expression(n, graph_test1, cor = 0.8, comm = TRUE))
+  expect_equal(nrow(expression_matrix1), length(V(graph_test1)))
+  expect_equal(ncol(expression_matrix1), n)
+  expect_equal(round(mean(expression_matrix1), 0), 0)
+  expect_equal(round(sd(expression_matrix1), 0), 1)
+  expect_true(is.matrix(expression_matrix1))
+})
+
+test_that("Generate Expression matrix from graph object by relative distance", {
+  graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
+  graph_test1 <- graph.edgelist(graph_test1_edges, directed = T)
+  adjacency_matrix1 <- make_adjmatrix_graph(graph_test1)
+  n <- 100
+  suppressWarnings(expression_matrix1 <- generate_expression(n, graph_test1, cor = 0.8, dist = TRUE))
+  expect_equal(nrow(expression_matrix1), length(V(graph_test1)))
+  expect_equal(ncol(expression_matrix1), n)
[…]

### 2020-03-02 tests/testthat/test_plot.R

@@ -1,0 +1,54 @@
+library("graphsim")
+library("igraph")
+library("vdiffr")
+context("Plot Directed Graphs")
+
+use_catch("../..")
+
+#generate fixed layout
+graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
+graph_test1 <- graph.edgelist(graph_test1_edges, directed = T)
+## note that it is not possible to pass a set.seed to layout
+# coords <- layout_with_fr(graph_test1, niter=50, start.temp=sqrt(10)/10)
+coords <- structure(c(0.697871079806861, -0.229514654942849, -1.43395965899018, 
+                      0.0584109525510336, 1.04730615951541, 0.189948108158827, 0.556504819830214, 
+                      -1.03768519984881), .Dim = c(4L, 2L))
+
+
+test_that("Plot an undirected graph structure", {
+  graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
+  graph_test1 <- graph.edgelist(graph_test1_edges, directed = T)
+  graph_test1 <- as.undirected(graph_test1)
+  set.seed = 9000
+  expect_doppelganger("test plotting undirected with igraph", plot(graph_test1, layout = coords))
+  set.seed = 9000
+  expect_doppelganger("test plotting undirected with plot_directed", plot_directed(graph_test1, layout = coords))
+})
+
+test_that("Plot a directed graph structure", {
+  graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
+  graph_test1 <- graph.edgelist(graph_test1_edges, directed = T)
+  set.seed = 9000
+  expect_doppelganger("test plotting directed with igraph", plot(graph_test1, layout = coords))
+  set.seed = 9000
+  expect_doppelganger("test plotting directed with plot_directed", plot_directed(graph_test1, layout = coords))
+  set.seed = 9000
+  expect_doppelganger("test plotting directed with plot_directedas state vector", plot_directed(graph_test1, layout = coords, state = c(1, 0, 1)))
+})
+
+test_that("Plot a graph structure with plotting parameters", {
+  graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
+  graph_test1 <- graph.edgelist(graph_test1_edges, directed = T)
+  set.seed = 9000
[…]

### 2020-03-02 tests/testthat/test_sigma.R

@@ -1,0 +1,127 @@
+library("graphsim")
+library("igraph")
+context("Make Sigma Matrix")
+
+test_that("Generate Sigma matrix from adjacency matrix", {
+    graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
+    graph_test1 <- graph.edgelist(graph_test1_edges, directed = T)
+    adjacency_matrix1 <- make_adjmatrix_graph(graph_test1)
+    sigma_matrix1 <- make_sigma_mat_adjmat(adjacency_matrix1, cor = 0.8)
+    expect_equal(isSymmetric(sigma_matrix1), TRUE)
+    expect_true(all(diag(sigma_matrix1) == rep(1, nrow(adjacency_matrix1))))
+    expect_equal(nrow(sigma_matrix1), length(V(graph_test1)))
+    expect_equal(ncol(sigma_matrix1), length(V(graph_test1)))
+    expect_equal(sum(sigma_matrix1), 8.8)
+    expect_true(all(adjacency_matrix1 == cbind(c(0, 1, 0, 0), c(1, 0, 1, 1), c(0, 1, 0, 0), c(0, 1, 0, 0))))
+    expect_true(all(sigma_matrix1 == cbind(c(1, 0.8, 0, 0), c(0.8, 1, 0.8, 0.8), c(0, 0.8, 1, 0), c(0, 0.8, 0, 1))))
+    expect_true(all(sigma_matrix1 %in% c(0, 1, 0.8/2^c(0:100))))
+  })
+
+test_that("Generate Sigma matrix from distances derived from adjacency matrix", {
+  graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("C", "D"), c("A", "D"), c("D", "C"), c("A", "C"))
+  graph_test1 <- graph.edgelist(graph_test1_edges, directed = T)
+  adjacency_matrix1 <- make_adjmatrix_graph(graph_test1)
+  expect_error(make_sigma_mat_dist_adjmat(adjacency_matrix1, cor = 0.8, absolute = TRUE), "distance matrix must have diagonal of zero")
+  diag(adjacency_matrix1) <- 1
+  expect_error(make_sigma_mat_dist_adjmat(adjacency_matrix1, cor = 0.8, absolute = TRUE), "distance matrix expected, not adjacency matrix")
+})
+
[…]

### 2020-03-02 tests/testthat/test_state.R

@@ -12,6 +12,8 @@ test_that("Default to activating matrix", {
   expect_equal(ncol(state_matrix1), length(V(graph_test1)))
   expect_equal(sum(state_matrix1), length(V(graph_test1))**2)
   expect_equal(all(state_matrix1 > 0), TRUE)
+  expect_equal(all(is.matrix(state_matrix1)), TRUE)
+  expect_true(all(state_matrix1 == 1))
 })
 
 test_that("text input to activating matrix", {
@@ -24,6 +26,7 @@ test_that("text input to activating matrix", {
   expect_equal(ncol(state_matrix1), length(V(graph_test1)))
   expect_equal(sum(state_matrix1), length(V(graph_test1))**2)
   expect_equal(all(state_matrix1 > 0), TRUE)
+  expect_true(all(state_matrix1 == 1))
 })
 
 test_that("text input to activating matrix", {
@@ -36,6 +39,7 @@ test_that("text input to activating matrix", {
   expect_equal(ncol(state_matrix1), length(V(graph_test1)))
   expect_equal(sum(state_matrix1), length(V(graph_test1))**2)
   expect_equal(all(state_matrix1 > 0), TRUE)
+  expect_true(all(state_matrix1 == 1))
   state_matrix1 <- make_state_matrix(graph_test1, state = 1)
   expect_equal(isSymmetric(state_matrix1), TRUE)
   expect_equal(sum(diag(state_matrix1)), length(V(graph_test1)))
@@ -43,6 +47,7 @@ test_that("text input to activating matrix", {
   expect_equal(ncol(state_matrix1), length(V(graph_test1)))
   expect_equal(sum(state_matrix1), length(V(graph_test1))**2)
   expect_equal(all(state_matrix1 > 0), TRUE)
+  expect_true(all(state_matrix1 == 1))
 })
 
 test_that("text input to inhibiting matrix", {
@@ -56,6 +61,8 @@ test_that("text input to inhibiting matrix", {
   expect_equal(sum(state_matrix1), length(V(graph_test1)))
   expect_equal(length(V(graph_test1))+sum(state_matrix1 < 0)-sum(state_matrix1 > 0) == 0, TRUE)
   expect_equal(all(state_matrix1 > 0), FALSE)
+  expect_true(all(state_matrix1 == 1 || state_matrix1 == -1))
+  expect_true(all(diag(state_matrix1) == 1))
 })
 
[…]

### 2020-03-02 R/catch-routine-registration.R
improved handling to support testing

@@ -1,0 +1,6 @@
+# This dummy function definition is included with the package to ensure that
+# 'tools::package_native_routine_registration_skeleton()' generates the required
+# registration info for the 'run_testthat_tests' symbol.
+(function() {
+  .Call("run_testthat_tests", PACKAGE = "graphsim")
+})

### 2020-03-02 R/generate_expression_graph.R

@@ -14,10 +14,25 @@
 ##' @param comm,absolute logical. Parameters for Sigma matrix generation. Passed on to \code{\link[graphsim]{make_sigma_mat_dist_graph}} or \code{\link[graphsim]{make_sigma_mat_graph}}.
 ##' @keywords graph network igraph mvtnorm simulation
 ##' @import igraph mvtnorm
+##' @importFrom igraph is_igraph
 ##' @importFrom Matrix nearPD
 ##' @importFrom matrixcalc is.symmetric.matrix is.positive.definite
 ##' @export
 generate_expression <- function(n, graph, state = NULL, cor = 0.8, mean = 0, comm = F, dist = F, absolute = F){
+  if(!is.integer(n)){
+    if(is.numeric(n)){
+      if(floor(n) == n){
+        n <- as.integer(n)
+      } else{
+        n <- floor(n)
+        print(paste("rounding to sample size", n))
+      }
+    } else{
+      stop("sample size n must be an integer of length 1")
+    }
+  }
+  if(length(n) > 1) stop("sample size n must be an integer of length 1")
+  if(!is_igraph(graph)) stop("graph must be an igraph class")
   if(is.vector(state) || length(state) == 1) state <- make_state_matrix(graph, state)
   if(!(is.vector(mean)) || length(mean) == 1 ) mean <- rep(mean,length(V(graph)))
   if(dist){
@@ -35,7 +50,7 @@ generate_expression <- function(n, graph, state = NULL, cor = 0.8, mean = 0, com
     sig <- as.matrix(nearPD(sig, corr=T, keepDiag = T)$mat) #postive definite correction
   }
   expr_mat <- t(rmvnorm(n,mean=mean, sigma=sig))
-  rownames(expr_mat)<-names(V(graph))
-  colnames(expr_mat)<-paste0("sample_", 1:n)
+  rownames(expr_mat) <-names(V(graph))
+  colnames(expr_mat) <-paste0("sample_", 1:n)
   return(expr_mat)
 }

### 2020-03-02 R/make_distance_graph.R

@@ -4,6 +4,7 @@
 ##'
 ##' @description Compute the distance matrix of using shortest paths of a (directed) \code{\link[igraph]{igraph}} structure, normalising by the diameter of the network, preserving node/column/row names (and direction).
 ##'
+##' @param mat precomputed adjacency or commonlink matrix.
 ##' @param graph An \code{\link[igraph]{igraph}} object. May be directed or weighted.
 ##' @param directed logical. Whether directed information is passed to the distance matrix.
 ##' @param absolute logical. Whether distances are scaled as the absolute difference from the diameter (maximum possible). Defaults to TRUE. The alternative is to calculate a relative difference from the diameter for a geometric decay in distance.
@@ -12,7 +13,24 @@
 ##' @export
 make_distance_graph <- function(graph, directed = T, absolute = F){
   if(directed == F) graph <- as.undirected(graph)
-  diam <-diameter(graph)
+  diam <- diameter(graph)
+  if (absolute){
+    mat <- (diam-shortest.paths(graph))/diam
+  } else {
+    mat <- 1^-diam/(diam*shortest.paths(graph))
+    diag(mat) <- 1
+  }
+  rownames(mat) <- colnames(mat) <- names(V(graph))
+  return(mat)
+}
+
+##' @rdname make_distance
+##' @importFrom igraph graph_from_adjacency_matrix
+##' @export
+make_distance_adjmat <- function(mat, directed = T, absolute = F){
+  diag(mat) <- 0
+  graph <- graph_from_adjacency_matrix(mat, weighted = NULL, mode = "undirected")
+  diam <- diameter(graph)
   if (absolute){
     mat <- (diam-shortest.paths(graph))/diam
   } else {

### 2020-03-02 R/make_sigma_graph.R

@@ -38,6 +38,8 @@ make_sigma_mat_graph <- function(graph, cor, comm = F, directed = F){
 ##' @rdname make_sigma
 ##' @export
 make_sigma_mat_dist_adjmat <- function(mat, cor, absolute = F){
+  if(!(all(diag(mat) == 1))) stop("distance matrix must have diagonal of zero")
+  if(!(max(mat[mat!=1]) > 0) || !(max(mat[mat!=1]) <= 1)) stop("distance matrix expected, not adjacency matrix")
   sig <- mat/max(mat[mat!=1]) * cor
   sig <- ifelse(sig>0, sig, 0)
   diag(sig)<-1

### 2020-03-02 R/plot_directed.R

@@ -17,7 +17,7 @@
 ##' @param graph An \code{\link[igraph]{igraph}} object. Must be directed with known states.
 ##' @param state character or integer. Defaults to "activating" if no "state" edge attribute found. May be applied a scalar across all edges or as a vector for each edge respectively. Accepts non-integer values for weighted edges provided that the sign indicates whether links are activating (positive) or inhibiting (negative). May also be entered as text for "activating" or "inhibiting" or as integers for activating (0,1) or inhibiting (-1,2). Compatible with inputs for make_state_matrix or generate_expression_graph in the graphsim package \url{https://github.com/TomKellyGenetics/graphsim}.
 ##' @param labels character vector. For labels to plot nodes. Defaults to vertex names in graph object. Entering "" would yield unlabelled nodes.
-##' @param layout function. Layout function as selected from \code{\link[igraph]{layout_}}. Defaults to layout.fruchterman.reingold. Alternatives include layout.kamada.kawai, layout.reingold.tilford, layout.sugiyama, and layout.davidson.harel.
+##' @param layout function. Layout function as selected from \code{\link[igraph]{layout_}}. Defaults to layout.fruchterman.reingold. Alternatives include layout.kamada.kawai, layout.reingold.tilford, layout.sugiyama, and layout.davidson.harel. A 2-column layout matrix giving x and y co-ordinates of each node can be given.
 ##' @param cex.node numeric. Defaults to 1.
 ##' @param cex.label numeric. Defaults to 0.75.
 ##' @param cex.main numeric. Defaults to 0.8.
@@ -62,7 +62,15 @@
 ##' plot_directed(graph_test4, state=c(1, 1, -1, 1, -1, 1, -1, 1), layout = layout.kamada.kawai)
 ##' @export
 plot_directed <- function(graph, state = NULL, labels = NULL, layout = layout.fruchterman.reingold, cex.node = 1, cex.label = 0.75, cex.arrow=1.25, cex.main=0.8, arrow_clip = 0.075, pch=21, border.node="grey33", fill.node="grey66", col.label = NULL, col.arrow=NULL, main=NULL, sub=NULL, xlab="", ylab="", frame.plot=F){
-  L <- layout(graph)
+  if(is.function(layout)){
+    L <-  layout(graph)
+  } else {
+    if(is.matrix(layout) && nrow(layout) == length(V(graph)) && ncol(layout) == 2){
[…]

### 2020-03-02 R/plot_directed.R
checks before release

@@ -12,7 +12,7 @@
 ##'
 ##' @title Extensions to iGraph for Customising plots
 ##'
-##' @description Functions to plot_directed or graph structures including customised colors, layout, states, arrows. Uses graphs functions as an extension of \code{\link[igraph]{igraph}}. Designed for plotting directed graphs.
+##' @description Functions to plot_directed or graph structures including customised colours, layout, states, arrows. Uses graphs functions as an extension of \code{\link[igraph]{igraph}}. Designed for plotting directed graphs.
 ##'
 ##' @param graph An \code{\link[igraph]{igraph}} object. Must be directed with known states.
 ##' @param state character or integer. Defaults to "activating" if no "state" edge attribute found. May be applied a scalar across all edges or as a vector for each edge respectively. Accepts non-integer values for weighted edges provided that the sign indicates whether links are activating (positive) or inhibiting (negative). May also be entered as text for "activating" or "inhibiting" or as integers for activating (0,1) or inhibiting (-1,2). Compatible with inputs for make_state_matrix or generate_expression_graph in the graphsim package \url{https://github.com/TomKellyGenetics/graphsim}.
@@ -28,7 +28,7 @@
 ##' @param border.node character. Specifies the colours of node border passed to plot. Defaults to grey33. Applies to whole node shape if pch has only one colour.
 ##' @param fill.node character. Specfies the colours of node fill passed to plot. Defaults to grey66.
 ##' @param col.arrow character. Specfies the colours of arrows passed to plot. Defaults to par("fg").  May take a scalar applied to all edges or a vector with colours for each edge respectively.
-##' @param main,sub,xlab,ylab Plotting paramaters to specify plot titles or axes labels
+##' @param main,sub,xlab,ylab Plotting parameters to specify plot titles or axes labels
 ##' @param frame.plot logical. Whether to frame plot with a box. Defaults to FALSE.
 ##' @keywords graph igraph igraph plot
 ##' @import igraph graphics

### 2020-04-30 paper/paper.Rmd
migrate to Rmarkdown #13 (figure 1)

---
title: 'graphsim: An R package for simulating gene expression data from graph structures of biological pathways'
output:
  rmarkdown::pdf_document:
    fig_crop: no
    keep_md: TRUE
    fig_caption: yes
tags:
  - R
  - gene-expression
  - simulation
  - genomics
  - pathway
  - network
authors:
  - name: S. Thomas Kelly
    email: "tom.kelly@postgrad.otago.ac.nz, tom.kelly@riken.jp"
    orcid: 0000-0003-3904-6690
    affiliation: "1, 2" # (Multiple affiliations must be quoted)
  - name: Michael A. Black
    email: mik.black@otago.ac.nz
    orcid: 0000-0003-1174-6054
    affiliation: "1"
affiliations:
 - name: "Department of Biochemistry, University of Otago, PO Box 56, Dunedin 9054, New Zealand"
   index: 1
 - name: "RIKEN Center for Integrative Medical Sciences, Suehiro-cho-1-7-22, Tsurumi Ward, Yokohama, Kanagawa 230-0045, Japan"
   index: 2
date: 30 April 2020
bibliography: paper.bib
header-includes:
- \usepackage{caption}
- \usepackage{subfig}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", width = 68)
knitr::opts_chunk$set(fig.cap = "", fig.path = "Plot")
options(width = 68, cli.unicode = FALSE, cli.width = 68)
par(mai=c(1.02, 0.82, 0.82, 0.82)-0.82)
par(bty="o")
captioner::captioner()
```

### Summary
Transcriptomic analysis is used to capture the molecular state of a cell
or sample in many biological and medical applications. In addition to 
identifying alterations in activity at the level of individual genes, 
understanding changes in the gene networks that regulate fundamental
biological mechanisms is also an important objective of molecular 
analysis. As a result, databases that describe biological pathways 
are increasingly relied on to assist with the interpretation of results
from large-scale genomics studies. Incorporating information from 
biological pathways and gene regulatory networks into a genomic data
analysis is a popular strategy, and there are many methods that provide
this functionality for gene expression data. When developing or comparing
such methods, it is important to gain an accurate assessment of their 
performance, with simulation-based validation studies a popular choice.
This necessitates the use of simulated data that correctly accounts for
pathway relationships and correlations. Here we present a versatile
statistical framework to simulate correlated gene expression data from
biological pathways, by sampling from a multivariate normal distribution
derived from a graph structure. This procedure has been released as the
\texttt{graphsim} R package on CRAN and GitHub (\url{https://github.com/TomKellyGenetics/graphsim})
 and is compatible with any graph structure that can be described using
the `igraph` package. This package allows the simulation of biological
pathways from a graph structure based on a statistical model of gene expression.

Introduction: inference and modelling of biological networks {#sec:intro}
============================================================

Network analysis of molecular biological pathways has the potential to
lead to new insights into biology and medical genetics
[@Barabasi2004; @Hu2016]. Since gene expression profiles capture a
consistent signature of the regulatory state of a cell
[@Perou2000; @Ozsolak2011; @Svensson2018], they can be used to analyse
complex molecular states with genome-scale data. However, biological
pathways are often analysed in a reductionist paradigm as amorphous sets
of genes involved in particular functions, despite the fact that the
relationships defined by pathway structure could further inform gene
expression analyses. In many cases, the pathway relationships are
well-defined, experimentally-validated, and are available in public
databases [@Reactome]. As a result, network analysis techniques could
play an important role in furthering our understanding of biological
pathways and aiding in the interpretation of genomics studies.

Gene networks provide insights into how cells are regulated, by mapping
regulatory interactions between target genes and transcription factors,
enhancers, and sites of epigenetic marks or chromatin structures
[@Barabasi2004; @Yamaguchi2007]. Inference of these regulatory
interactions for genomics investigations has the potential to radically
expand the range of candidate biological pathways to be further
explored, or to improve the accuracy of bioinformatics and functional
genomic analysis. A number of methods have already been developed to
utilise timecourse gene expression data [@Arner2015; @Yamaguchi2007]
using gene regulatory modules in state-space models and recursive vector
autoregressive models [@Hirose2008; @Shimamura2009]. Various approaches
to gene regulation and networks at the genome-wide scale have led to
novel biological insights [@Arner2015; @Komatsu2013]. However, inference
of regulatory networks has thus far relied on experimental validation or
resampling-based approaches to estimate the likelihood of specific
network modules being predicted [@Markowetz2007; @Hawe2019].

There is a need, therefore, for a systematic framework for statistical
modelling and simulation of gene expression data derived from
hypothetical, inferred or known gene networks. Here we present a
package to achieve this, where samples from a multivariate normal
distribution are used to generate normally-distributed log-expression
data, with correlations between genes derived from the structure of the
underlying pathway or gene regulatory network. This methodology enables
simulation of expression profiles that approximate the log-transformed
and normalised data from microarray and bulk or single-cell RNA-Seq
experiments. This procedure has been released as the package to enable
the generation of simulated gene expression datasets containing pathway
relationships from a known underlying network. These simulated datasets
can be used to evaluate various bioinformatics methodologies, including
statistical and network inference procedures.

Methodology and software {#sec:methods}
========================

Here we present a procedure to simulate gene expression data with
correlation structure derived from a known graph structure. This
procedure assumes that transcriptomic data have been generated and
follow a log-normal distribution (i.e.,
$log(X_{ij}) \sim MVN({\bf\mu}, \Sigma)$, where ${\bf\mu}$ and $\Sigma$
are the mean vector and variance-covariance matrix respectively, for
gene expression data derived from a biological pathway) after
appropriate normalisation [@Law2014; @Li2015]. Log-normality of gene
expression matches the assumptions of the popular \texttt{limma} package [@limma], which is
often used for the analysis of intensity-based data from gene expression
microarray studies and count-based data from RNA-Seq experiments. This
approach has also been applied for modelling UMI-based count data from
single-cell RNA-Seq experiments in the package [@Wang2018].

In order to simulate transcriptomic data, a pathway is first constructed
as a graph structure, using the \texttt{igraph} R package [@igraph], with the status of
the edge relationships defined (i.e, whether they activate or inhibit
downstream pathway members). [This procedure uses]{style="color: black"}
a graph structure such as that presented in
Figure [1a](#fig:simple_graph:first){reference-type="ref"
reference="fig:simple_graph:first"}. The graph can be defined by an
adjacency matrix, **$A$** (with elements
$A_{ij}$), where $$A_{ij} = 
\begin{cases}
   1                         & \mbox{if genes } i \mbox{ and } j \mbox{ are adjacent} \\
   0                         & \mbox{otherwise}
\end{cases}$$

A matrix, **$R$**, with elements
[$R_{ij}$]{style="color: black"}, is calculated based on distance (i.e.,
number of edges contained in the shortest path) between nodes, such that
closer nodes are given more weight than more distant nodes, to define
inter-node relationships. A geometrically-decreasing (relative) distance
weighting is used to achieve this:

[ $$R_{ij} = 
\begin{cases}
   1                            & \mbox{if genes } i \mbox{ and } j \mbox{ are adjacent} \\
   (\frac{1}{2})^{d_{ij}}      & \mbox{if a path can be found between genes } i \mbox{ and } j \\
   0                            & \mbox{if no path exists between genes } i \mbox{ and } j 
\end{cases}$$]{style="color: black"} where $d_{ij}$ is the length of
the shortest path (i.e., minimum number of edges traversed) between
genes (nodes) $i$ and $j$ in graph $G$. Each more distant node is thus
related by $\frac{1}{2}$ compared to the next nearest, as shown in
Figure [2b](#fig:simulation_activating:second){reference-type="ref"
reference="fig:simulation_activating:second"}. An
arithmetically-decreasing (absolute) distance weighting is also
supported in the \texttt{graphsim} R package which implements this procedure: [ $$R_{ij} = 
\begin{cases}
   1                            & \mbox{if genes } i \mbox{ and } j \mbox{ are adjacent} \\
   1-\frac{d_{ij}}{diam(G)}     & \mbox{if a path can be found between genes } i \mbox{ and } j \\
   0                            & \mbox{if no path exists between genes } i \mbox{ and } j 
\end{cases}$$ ]{style="color: black"}

Assuming a unit variance for each gene, these values can be used to
derive a $\Sigma$ matrix: $$\Sigma_{ij} = 
\begin{cases}
   1                            & \mbox{if } i=j \\
   \rho R_{ij}  & \mbox{otherwise}
\end{cases}$$ where $\rho$ is the correlation between adjacent nodes.
Thus covariances between adjacent nodes are assigned by a correlation
parameter ($\rho$) and the remaining off-diagonal values in the matrix
are based on scaling these correlations by the geometrically weighted
relationship matrix (or the nearest positive definite matrix for
$\Sigma$ with negative correlations).\
Computing the nearest positive definite matrix is necessary to ensure
that the variance-covariance matrix could be inverted when used as a
parameter in multivariate normal simulations, particularly when negative
correlations are included for inhibitions (as shown below). Matrices
that could not be inverted occurred rarely with biologically plausible
graph structures but this approach allows for the computation of a
plausible correlation matrix when the given graph structure is
incomplete or contains loops. When required, the nearest positive
definite matrix is computed using the `nearPD` function of the \texttt{Matrix} R package
[@Matrix] to perform Higham's algorithm [@Higham2002] on
variance-covariance matrices. The \texttt{graphsim} package gives a warning when this
occurs.

Illustrations {#sec:illustrations}
=============

Generating a Graph Structure {#sec:plot_graph}
----------------------------

The graph structure in
Figure [1a](#fig:simple_graph:first){reference-type="ref"
reference="fig:simple_graph:first"} was used to simulate correlated gene
expression data by sampling from a multivariate normal distribution
using the R package [@Genz2009; @mvtnorm]. The graph structure
visualisation in
Figure [1](#fig:simple_graph){reference-type="ref"
reference="fig:simple_graph"} was specifically developed for (directed)
\texttt{igraph} objects in and is available in the and packages. The
\texttt{plot_directed} function enables customisation of plot parameters for
each node or edge, and mixed (directed) edge types for indicating
activation or inhibition. These inhibition links (which occur frequently
in biological pathways) are demonstrated in
Figure [1b](#fig:simple_graph:second){reference-type="ref"
reference="fig:simple_graph:second"}.

A graph structure can be generated and plotted using the following
commands in R:

```{r, eval = FALSE}
#install packages required (once per machine)

install.packages("graphsim")
```

```{r}
#load required packages (once per R instance)

library("igraph")
library("graphsim")
library("gplots")
```

```{r, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.455\\linewidth', out.width='.455\\linewidth', fig.retina=20, fig.margin = FALSE, fig.cap = '\\textbf{Simulated graph structures.}. A constructed graph structure used as an example to demonstrate the simulation procedure inFigures 2 and 3. Activating links are denoted by black arrows and inhibiting links by red edges. Inhibiting edges have been highlighted in red.', fig.subcap=c("Activating pathway structure", "Inhibiting pathway structure")}
#generate graph structure

graph_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"), c("D", "E"),
                     c("D", "F"), c("F", "G"), c("F", "I"), c("H", "I"))
graph <- graph.edgelist(graph_edges, directed = TRUE)

#plot graph structure (Figure 1a)

plot_directed(graph, state ="activating", layout = layout.kamada.kawai,
              cex.node=3, cex.arrow=5, arrow_clip = 0.2)

#generate parameters for inhibitions

state <- c(1, 1, -1, 1, 1, 1, 1, -1, 1)

#plot graph structure with inhibitions (Figure 1b)

plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node=3, cex.arrow=5, arrow_clip = 0.2)
```

![\small \textbf{\textbf{Simulating expression from a graph structure.}}
An example of a graph structure (a) that has been used to derive a
relationship matrix (b), $\Sigma$ matrix (c) and correlation structure
(d) from the relative distances between the nodes. Non-negative values
are coloured white to red from $0$ to $1$. This $\Sigma$ matrix has been
used to generate a simulated expression dataset of 100 samples (coloured
blue to red from low to high) via sampling from the multivariate
normal distribution. Here genes with closer relationships in the pathway
structure show a higher correlation between simulated values.](figure2.pdf){ width=95% #fig:simulation_activating label="simulation_activating"}

Generating a Simulated Expression Dataset {#sec:graphsim_demo}
-----------------------------------------

The correlation parameter of $\rho = 0.8$ is used to demonstrate the
inter-correlated datasets using a geometrically-generated relationship
matrix (as used for the example in
Figure [2c](#fig:simulation_activating:third){reference-type="ref"
reference="fig:simulation_activating:third"}). This $\Sigma$ matrix was
then used to sample from a multivariate normal distribution such that
each gene had a mean of $0$, standard deviation $1$, and covariance
within the range $[0,1]$ so that the off-diagonal elements of $\Sigma$
represent correlations. This procedure generated a simulated (continuous
normally-distributed) log-expression profile for each node
(Figure [2e](#fig:simulation_activating:fourth){reference-type="ref"
reference="fig:simulation_activating:fourth"}) with a corresponding
correlation structure (Figure [2d](#fig:simulation_activating:fifth){reference-type="ref"
reference="fig:simulation_activating:fifth"}). The simulated correlation
structure closely resembled the expected correlation structure ($\Sigma$
in Figure [2c](#fig:simulation_activating:third){reference-type="ref"
reference="fig:simulation_activating:third"}) even for the relatively
modest sample size ($N=100$) illustrated in
Figure [2](#fig:simulation_activating){reference-type="ref"
reference="fig:simulation_activating"}. Once a gene expression dataset
comprising multiple pathways has been generated (as in
Figure [2e](#fig:simulation_activating:fourth){reference-type="ref"
reference="fig:simulation_activating:fourth"}), it can then be used to
test procedures designed for analysis of empirical gene expression data
(such as those generated by microarrays or RNA-Seq) that have been
normalised on a log-scale.

The simulated dataset can be generated using the following code:

```r
# activating graph
state <- rep(length(E(graph)), 1)
plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node=3, cex.arrow=5, arrow_clip = 0.2)

#adjacency matrix

adj_mat <- make_adjmatrix_graph(graph)

#relationship matrix

dist_mat <- make_distance_graph(graph, absolute = FALSE)

#sigma matrix directly from graph

sigma_mat <- make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE)

#show shortest paths of graph

shortest_paths <- shortest.paths(graph)

#generate expression data directly from graph

expr <- generate_expression(100, graph, cor = 0.8, mean = 0, comm = FALSE,
          dist = TRUE, absolute = FALSE, state = state)

#plot adjacency matrix

library("gplots")

heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
          col = colorpanel(3, "grey75", "white", "blue"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))

#plot relationship matrix

heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))

#plot sigma matrix

heatmap.2(make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
```

![\small \textbf{\textbf{Simulating expression from graph structure with inhibitions.}}
Simulating expression from graph structure with inhibitions.}} An example of a graph
structure (a), that has been used to derive a relationship matrix (b), $\Sigma$ matrix
(c), and correlation structure (d), from the relative distances between the nodes.
These values are coloured blue to red from $-1$ to $1$. This has been used to generate
a simulated expression dataset of 100 samples (coloured blue to red from low to high)
via sampling from the multivariate normal distribution. Here the inhibitory relationships
between genes are reflected in negatively correlated simulated  values.](figure3.pdf){ width=95% #fig:simulation_inhibiting label="fig:simulation_inhibiting"}

![\small \textbf{\textbf{Simulating expression from a biological pathway graph structure.}}
Simulating expression from graph structure with inhibitions.}} The graph structure (a) of a
known biological pathway, the TGF-$\beta$ receptor signaling activates SMADs (R-HSA-2173789),
was used to derive a relationship matrix (b), $\Sigma$ matrix (c) and correlation structure
(d) from the relative distances between the nodes. These values are coloured blue to red from
$-1$ to $1$. This has been used to generate a simulated expression dataset of 100 samples
(coloured blue to red from low to high) via sampling from the multivariate normal distribution.
Here modules of genes with correlated expression can be clearly discerned.](figure4.pdf){ width=95% #fig:simulation_smad label="fig:simulation_smad"}

 \newpage

```r
expr <- generate_expression(100, graph, cor = 0.8, mean = 0,
          comm = FALSE, dist =TRUE, absolute = FALSE, state = state)

#plot simulated expression data

heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))

#plot simulated correlations

heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
```

The simulation procedure
(Figure [2](#fig:simulation_activating){reference-type="ref"
reference="fig:simulation_activating"}) can similarly be used for
pathways containing inhibitory links
(Figure [3](#fig:simulation_inhibiting){reference-type="ref"
reference="fig:simulation_inhibiting"}) with several refinements. With
the inhibitory links
(Figure [3a](#fig:simulation_inhibiting:first){reference-type="ref"
reference="fig:simulation_inhibiting:first"}), distances are calculated
in the same manner as before
(Figure [3b](#fig:simulation_inhibiting:second){reference-type="ref"
reference="fig:simulation_inhibiting:second"}) with inhibitions
accounted for by iteratively multiplying downstream nodes by $-1$ to
form modules with negative correlations between them
(Figures [3c
](#fig:simulation_inhibiting:third){reference-type="ref"
reference="fig:simulation_inhibiting:third"}
and [3d](#fig:simulation_inhibiting:fifth){reference-type="ref"
reference="fig:simulation_inhibiting:fifth"}). A multivariate normal
distribution with these negative correlations can be sampled to generate
simulated data
(Figure [3e](#fig:simulation_inhibiting:fourth){reference-type="ref"
reference="fig:simulation_inhibiting:fourth"}).

```r
state <- c(1, 1, -1, 1, 1, 1, 1, -1, 1)
plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node=3, cex.arrow=5, arrow_clip = 0.2)

#adjacency matrix
adj_mat <- make_adjmatrix_graph(graph)
#relationship matrix
dist_mat <- make_distance_graph(graph, absolute = FALSE)
#state matrix
state_mat <- make_state_matrix(graph, state)
#sigma matrix directly from graph
sigma_mat <- make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE)
#show shortest paths of graph
shortest_paths <- shortest.paths(graph)
#generate expression data directly from graph
expr <- generate_expression(100, graph, cor = 0.8, mean = 0, comm = FALSE,
                            dist = TRUE, absolute = FALSE, state = state)
#plot adjacency matrix
heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
          col = colorpanel(3, "grey75", "white", "blue"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
#plot state matrix
heatmap.2(make_state_matrix(graph, state),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
#plot relationship matrix
heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
#plot sigma matrix
heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0,
                            comm = FALSE, dist =TRUE, absolute = FALSE)
#plot simulated expression data
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
#plot simulated correlations
heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
```

The simulation procedure is also demonstrated here
(Figure [4](#fig:simulation_smad){reference-type="ref"
reference="fig:simulation_smad"}) on a pathway structure for a known
biological pathway (from reactome R-HSA-2173789) of TGF-$\beta$ receptor
signaling activates SMADs
(Figure [4a](#fig:simulation_smad:first){reference-type="ref"
reference="fig:simulation_smad:first"}) derived from the Reactome
database version 52 `[@Reactome]`. Distances are calculated in the same
manner as before
(Figure [4b](#fig:simulation_smad:second){reference-type="ref"
reference="fig:simulation_smad:second"}) producing blocks of correlated
genes
(Figures [4c](#fig:simulation_smad:third){reference-type="ref"
reference="fig:simulation_smad:third"}
and [4d](#fig:simulation_smad:fifth){reference-type="ref"
reference="fig:simulation_smad:fifth"}). This shows that
multivariate normal distribution can be sampled to generate simulated
data to represent expression with the complexity of a biological pathway
(Figure [4e](#fig:simulation_smad:fourth){reference-type="ref"
reference="fig:simulation_smad:fourth"}). Here *SMAD7* exhibits
negative correlations with the other SMADs consistent with it's
functions as as an "inhibitor SMAD" with competitively inhibits *SMAD4*.

These simulated datasets can also be used for simulating gene 
expression data within a graph network to test genomic analysis techniques.
Correlation structure can be included into datasets generated
when testing whether true positive genes or samples can be detected
in a sample with the background of complex pathway structure.

Summary and discussion {#sec:summary}
======================

Biological pathways are of fundamental importance to understanding
molecular biology. In order to translate findings from genomics studies
into real-world applications such as improved healthcare, the roles of
genes must be studied in the context of molecular pathways. Here we
present a statistical framework to simulate gene expression from
biological pathways, and provide the \texttt{graphsim} package in R to generate these
simulated datasets. This approach is versatile and can be fine-tuned for
modelling existing biological pathways or for testing whether
constructed pathways can be detected by other means. In particular,
methods to infer biological pathways and gene regulatory networks from
gene expression data can be tested on simulated datasets using this
framework. The package also enables simulation of complex gene
expression datasets to test how these pathways impact on statistical
analysis of gene expression data using existing methods or novel
statistical methods being developed for gene expression data analysis.

Computational details {#computational-details .unnumbered .unnumbered}
=====================

The results in this paper were obtained using R 3.6.1 with the \texttt{igraph} 1.2.4.1
\texttt{Matrix} 1.2-17, \texttt{matrixcalc} 1.0-3, and \texttt{mvtnorm} 1.0-11 packages. R itself and all dependent packages
used are available from the Comprehensive Archive Network (CRAN) at
\url{https://CRAN.R-project.org}. The \texttt{graphsim} package presented can be
installed from CRAN and the issues can  be reported to the development version on GitHub (\url{https://github.com/TomKellyGenetics/graphsim}). This package is included in the library on GitHub (\url{https://github.com/TomKellyGenetics/igraph.extensions}) which installs various
tools for \texttt{igraph} analysis. This software is cross-platform and
compatible with installations on Windows, Mac, and Linux operating
systems. The package GitHub repository also contains vignettes with more
information and examples on running functions released in the package.
The package (\texttt{graphsim} 0.1.2) has been released on CRAN and will be updated.

Acknowledgements {#acknowledgements .unnumbered .unnumbered}
================

This package was developed as part of a PhD research project funded by
the Postgraduate Tassell Scholarship in Cancer Research Scholarship
awarded to STK. We thank members of the Laboratory of Professor Satoru
Miyano at the University of Tokyo, Institute for Medical Science,
Professor Seiya Imoto, Associate Professor Rui Yamaguchi, and Dr Paul
Sheridan (Assistant Professor at Hirosaki University,CSO at Tupac Bio)
for helpful discussions in this field. We also thank Professor Parry
Guilford at the University of Otago, Professor Cristin Print at the
University of Auckland, and Dr Erik Arner at the RIKEN Center for
Integrative Medical Sciences for their excellent advice during this
project.

Author Contributions  {#contributions .unnumbered .unnumbered}
================

S.T.K. and M.A.B. conceived of the presented methodology. S.T.K. developed the theory and performed the computations.
M.A.B. provided guidance throughout the project and gave feedback on the package. All authors discussed the package and contributed to the final manuscript.

# References

### 2020-04-30 paper/paper.Rmd
build with subcaptions
build paper without subcaption or subfig package

@@
   - \usepackage{caption}
-- \usepackage{subfig}
+ # - \usepackage{subfig}
 ---
@@
 
-```{r, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.455\\linewidth', out.width='.455\\linewidth', fig.retina=20, fig.margin = FALSE, fig.cap = '\\textbf{Simulated graph structures.}.
+```{r, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.455\\linewidth', out.width='.455\\linewidth', fig.retina=10, fig.margin = FALSE, fig.cap = '\\textbf{Simulated graph structures.}.
 A constructed graph structure used as an example to demonstrate the simulation procedure inFigures 2 and 3.
 Activating links are denoted by black arrows and inhibiting links by red edges.
-Inhibiting edges have been highlighted in red.', fig.subcap=c("Activating pathway structure", "Inhibiting pathway structure")}
+Inhibiting edges have been highlighted in red.'}
+#, fig.subcap=c("Activating pathway structure", "Inhibiting pathway structure")}
 #generate graph structure
@@
 plot_directed(graph, state ="activating", layout = layout.kamada.kawai,
-              cex.node=3, cex.arrow=5, arrow_clip = 0.2)
+              cex.node = 2, cex.arrow = 4, arrow_clip = 0.2,
+              sub = "(a) Activating pathway structure", cex.sub = 1.5)
 
@@
 plot_directed(graph, state=state, layout = layout.kamada.kawai,
-              cex.node=3, cex.arrow=5, arrow_clip = 0.2)
+              cex.node = 2, cex.arrow = 4, arrow_clip = 0.2,
+              sub = "(b) Inhibiting pathway structure", cex.sub = 1.5)
 ```

### 2020-04-30 R/plot_directed.R

@@ -31,6 +31,7 @@
 ##' @param cex.node numeric. Defaults to 1.
 ##' @param cex.label numeric. Defaults to 0.75.
 ##' @param cex.main numeric. Defaults to 0.8.
+##' @param cex.sub numeric. Defaults to 0.8.
 ##' @param cex.arrow numeric Defaults to 1.25. May take a scalar applied to all edges 
 ##' or a vector with values for each edge respectively.
 ##' @param col.label character. Specfies the colours of node labels passed to plot. 
@@ -82,7 +83,7 @@
 ##' @return base R graphics
 ##' 
 ##' @export
-plot_directed <- function(graph, state = NULL, labels = NULL, layout = layout.fruchterman.reingold, cex.node = 1, cex.label = 0.75, cex.arrow=1.25, cex.main=0.8, arrow_clip = 0.075, pch=21, border.node="grey33", fill.node="grey66", col.label = NULL, col.arrow=NULL, main=NULL, sub=NULL, xlab="", ylab="", frame.plot=F){
+plot_directed <- function(graph, state = NULL, labels = NULL, layout = layout.fruchterman.reingold, cex.node = 1, cex.label = 0.75, cex.arrow=1.25, cex.main=0.8, cex.sub=0.8, arrow_clip = 0.075, pch=21, border.node="grey33", fill.node="grey66", col.label = NULL, col.arrow=NULL, main=NULL, sub=NULL, xlab="", ylab="", frame.plot=F){
   if(is.function(layout)){
     L <-  layout(graph)
   } else {
@@ -98,7 +99,7 @@ plot_directed <- function(graph, state = NULL, labels = NULL, layout = layout.fr
   Ne <- length(es[1]$V1)
   Xn <- L[,1]
   Yn <- L[,2]
-  plot(Xn, Yn, xaxt="n", yaxt="n", xlab=xlab, ylab=ylab, frame.plot=frame.plot, cex = 2 * cex.node, pch=1, col=par()$bg, main=main, sub=sub, cex.main=cex.main)
+  plot(Xn, Yn, xaxt="n", yaxt="n", xlab=xlab, ylab=ylab, frame.plot=frame.plot, cex = 2 * cex.node, pch=1, col=par()$bg, main=main, sub=sub, cex.main=cex.main, cex.sub=cex.sub)
   if(!is.null(get.edge.attribute(graph, "state"))){
     state <- get.edge.attribute(graph, "state")
   } else {

### 2020-04-30 paper/paper.Rmd
Rmarkdown figure 2 #13

@@
 
-```{r, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.455\\linewidth', out.width='.455\\linewidth', fig.retina=10, fig.margin = FALSE, fig.cap = '\\textbf{Simulated graph structures.}.
+```{r, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.400\\linewidth', out.width='.400\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulated graph structures}.
 A constructed graph structure used as an example to demonstrate the simulation procedure inFigures 2 and 3.
@@
 Inhibiting edges have been highlighted in red.'}
-#, fig.subcap=c("Activating pathway structure", "Inhibiting pathway structure")}
 #generate graph structure
@@
 ```
-
-![\small \textbf{\textbf{Simulating expression from a graph structure.}}
-An example of a graph structure (a) that has been used to derive a
-relationship matrix (b), $\Sigma$ matrix (c) and correlation structure
-(d) from the relative distances between the nodes. Non-negative values
-are coloured white to red from $0$ to $1$. This $\Sigma$ matrix has been
-used to generate a simulated expression dataset of 100 samples (coloured
-blue to red from low to high) via sampling from the multivariate
-normal distribution. Here genes with closer relationships in the pathway
-structure show a higher correlation between simulated values.](figure2.pdf){ width=95% #fig:simulation_activating label="simulation_activating"}
 
@@
 
-```r
+```{r, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.400\\linewidth', out.width='.400\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}.
+An example of a graph structure (a) that has been used to derive a relationship matrix (b), $\\Sigma$  matrix (c) and correlation structure (d) from the relative distances between the nodes.
+Non-negative values are coloured white to red from $0$ to $1$.
+This $\\Sigma$ matrix has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
+Here genes with closer relationships in the pathway structure show a higher correlation between simulated values.'}
 # activating graph
-state <- rep(length(E(graph)), 1)
+state <- rep(1, length(E(graph)))
 plot_directed(graph, state=state, layout = layout.kamada.kawai,
-              cex.node=3, cex.arrow=5, arrow_clip = 0.2)
+              cex.node=3, cex.arrow=5, arrow_clip = 0.2,
+              sub = "(a) Activating pathway structure", cex.sub = 1.5)
 
 #adjacency matrix
-
 adj_mat <- make_adjmatrix_graph(graph)
-
 #relationship matrix
-
 dist_mat <- make_distance_graph(graph, absolute = FALSE)
-
 #sigma matrix directly from graph
-
 sigma_mat <- make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE)
-
 #show shortest paths of graph
-
 shortest_paths <- shortest.paths(graph)
-
 #generate expression data directly from graph
-
 expr <- generate_expression(100, graph, cor = 0.8, mean = 0, comm = FALSE,
                             dist = TRUE, absolute = FALSE, state = state)
-
 #plot adjacency matrix
-
-library("gplots")
-
 heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
           col = colorpanel(3, "grey75", "white", "blue"),
-          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-
+          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
+              sub = "Adjacency matrix", cex.sub = 1.5)
 #plot relationship matrix
-
 heatmap.2(make_distance_graph(graph, absolute = FALSE),
           scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
-          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
+              sub = "(b) relationship matrix", cex.sub = 1.5)
 #plot sigma matrix
-
 heatmap.2(make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE),
 scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
+              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.5)
+expr <- generate_expression(100, graph, cor = 0.8, mean = 0,
+comm = FALSE, dist =TRUE, absolute = FALSE, state = state)
+#plot simulated expression data
+heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
+              sub = "(e) Simulated expressino data (log scale)", cex.sub = 1.5)
+#plot simulated correlations
+heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
 ```
+
+%![\small \textbf{\textbf{Simulating expression from a graph structure.}}
+%An example of a graph structure (a) that has been used to derive a
+%relationship matrix (b), $\Sigma$ matrix (c) and correlation structure
+%(d) from the relative distances between the nodes. Non-negative values
+%are coloured white to red from $0$ to $1$. This $\Sigma$ matrix has been
+%used to generate a simulated expression dataset of 100 samples (coloured
+%blue to red from low to high) via sampling from the multivariate
+%normal distribution. Here genes with closer relationships in the pathway
+%structure show a higher correlation between simulated values.](figure2.pdf){ width=95% #fig:simulation_activating %label="simulation_activating"}
 
@@
 Here modules of genes with correlated expression can be clearly discerned.](figure4.pdf){ width=95% #fig:simulation_smad label="fig:simulation_smad"}
-
- \newpage
-
-```r
-expr <- generate_expression(100, graph, cor = 0.8, mean = 0,
-          comm = FALSE, dist =TRUE, absolute = FALSE, state = state)
-
-#plot simulated expression data
-
-heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
-          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-
-#plot simulated correlations
-
-heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
-          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-```
 

### 2020-04-30 R/state.R

@@ -46,6 +46,7 @@ make_state_matrix <- function(graph, state = NULL){
   if(length(state) == 1) state <- rep(state, length(E(graph)))
   state[state == 2] <- -1
   state[state == 1 | state == 0] <- 1
+  if(is.numeric(state)) state[!(state %in% -1:2)] <- sign(state[!(state %in% -1:2)])
   if(is.character(state)){
     state <- as.list(state)
     state[grep("activating", state)] <- 1

### 2020-04-30 paper/paper.Rmd
Rmarkdown figure 3 #13

@@
 par(bty="o")
-captioner::captioner()
+#captioner::captioner(prefix = "Fig.")
 ```
@@
 
-```{r, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.400\\linewidth', out.width='.400\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulated graph structures}.
-A constructed graph structure used as an example to demonstrate the simulation procedure inFigures 2 and 3.
+```{r simple_graph, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulated graph structures}.
+A constructed graph structure used as an example to demonstrate the simulation procedure in Figures 2 and 3.
 Activating links are denoted by black arrows and inhibiting links by red edges.
@@
 
-```{r, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.400\\linewidth', out.width='.400\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}.
+```{r simulation_activating, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}.
 An example of a graph structure (a) that has been used to derive a relationship matrix (b), $\\Sigma$  matrix (c) and correlation structure (d) from the relative distances between the nodes.
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = "(e) Simulated expressino data (log scale)", cex.sub = 1.5)
+              sub = "(e) Simulated expression data (log scale)", cex.sub = 1.5)
 #plot simulated correlations
@@
 
-%![\small \textbf{\textbf{Simulating expression from a graph structure.}}
-%An example of a graph structure (a) that has been used to derive a
-%relationship matrix (b), $\Sigma$ matrix (c) and correlation structure
-%(d) from the relative distances between the nodes. Non-negative values
-%are coloured white to red from $0$ to $1$. This $\Sigma$ matrix has been
-%used to generate a simulated expression dataset of 100 samples (coloured
-%blue to red from low to high) via sampling from the multivariate
-%normal distribution. Here genes with closer relationships in the pathway
-%structure show a higher correlation between simulated values.](figure2.pdf){ width=95% #fig:simulation_activating %label="simulation_activating"}
+```{r simulation_inhibiting, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from graph structure with inhibitions}.
+An example of a graph structure (a), that has been used to derive a relationship matrix (b), $\\Sigma$ matrix (c), and correlation structure (d), from the relative distances between the nodes.
+These values are coloured blue to red from $-1$ to $1$.
+This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
+Here the inhibitory relationships between genes are reflected in negatively correlated simulated  values.'}
 
-![\small \textbf{\textbf{Simulating expression from graph structure with inhibitions.}}
-Simulating expression from graph structure with inhibitions.}} An example of a graph
-structure (a), that has been used to derive a relationship matrix (b), $\Sigma$ matrix
-(c), and correlation structure (d), from the relative distances between the nodes.
-These values are coloured blue to red from $-1$ to $1$. This has been used to generate
-a simulated expression dataset of 100 samples (coloured blue to red from low to high)
-via sampling from the multivariate normal distribution. Here the inhibitory relationships
-between genes are reflected in negatively correlated simulated  values.](figure3.pdf){ width=95% #fig:simulation_inhibiting label="fig:simulation_inhibiting"}
+#generate parameters for inhibitions
+state <- c(1, 1, -1, 1, 1, 1, 1, -1)
+plot_directed(graph, state=state, layout = layout.kamada.kawai,
+              cex.node=3, cex.arrow=5, arrow_clip = 0.2,
+              sub = "(a) Inhitibing pathway structure", cex.sub = 1.5)
+
+#adjacency matrix
+adj_mat <- make_adjmatrix_graph(graph)
+#relationship matrix
+dist_mat <- make_distance_graph(graph, absolute = FALSE)
+#sigma matrix directly from graph
+sigma_mat <- make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE)
+#show shortest paths of graph
+shortest_paths <- shortest.paths(graph)
+#generate expression data directly from graph
+expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0, comm = FALSE,
+                            dist = TRUE, absolute = FALSE)
+#plot adjacency matrix
+heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
+          col = colorpanel(3, "grey75", "white", "blue"),
+          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
+              sub = "Adjacency matrix", cex.sub = 1.5)
+# #plot relationship matrix
+heatmap.2(make_distance_graph(graph, absolute = FALSE),
+          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
+              sub = "(b) relationship matrix", cex.sub = 1.5)
+# #plot sigma matrix
+heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
+scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
+              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.5)
+expr <- generate_expression(100, graph,  state = state, cor = 0.8, mean = 0,
+comm = FALSE, dist =TRUE, absolute = FALSE)
+#plot simulated expression data
+heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
+              sub = "(e) Simulated expressinon data (log scale)", cex.sub = 1.5)
+#plot simulated correlations
+heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+```
 
@@
 reference="fig:simulation_inhibiting:fourth"}).
-
-```r
-state <- c(1, 1, -1, 1, 1, 1, 1, -1, 1)
-plot_directed(graph, state=state, layout = layout.kamada.kawai,
-              cex.node=3, cex.arrow=5, arrow_clip = 0.2)
-
-#adjacency matrix
-adj_mat <- make_adjmatrix_graph(graph)
-#relationship matrix
-dist_mat <- make_distance_graph(graph, absolute = FALSE)
-#state matrix
-state_mat <- make_state_matrix(graph, state)
-#sigma matrix directly from graph
-sigma_mat <- make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE)
-#show shortest paths of graph
-shortest_paths <- shortest.paths(graph)
-#generate expression data directly from graph
-expr <- generate_expression(100, graph, cor = 0.8, mean = 0, comm = FALSE,
-                            dist = TRUE, absolute = FALSE, state = state)
-#plot adjacency matrix
-heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
-          col = colorpanel(3, "grey75", "white", "blue"),
-          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-#plot state matrix
-heatmap.2(make_state_matrix(graph, state),
-          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
-          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-#plot relationship matrix
-heatmap.2(make_distance_graph(graph, absolute = FALSE),
-          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
-          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-#plot sigma matrix
-heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
-          scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
-          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0,
-                            comm = FALSE, dist =TRUE, absolute = FALSE)
-#plot simulated expression data
-heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
-          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-#plot simulated correlations
-heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
-          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-```
 

### 2020-04-30 paper/paper.Rmd
Rmarkdown figure 4 #13

@@
 
+```{r simulation_smad, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a biological pathway graph structure}.
+The graph structure (a) of a known biological pathway, the TGF-$\\beta$ receptor signaling activates SMADs (R-HSA-2173789), was used to derive a relationship matrix (b), $\\Sigma$ matrix (c) and correlation structure (d) from the relative distances between the nodes.
+These values are coloured blue to red from $-1$ to $1$.
+This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
+Here modules of genes with correlated expression can be clearly discerned.'}
+
+#import graph from data
+graph <- identity(TGFBeta_Smad_graph)
+
+#generate parameters for inhibitions
+state <- rep(1, length(E(graph)))
+pathway <- get.edgelist(graph)
+state[pathway[,1] %in% c("SMAD6", "SMAD7", "BAMBI", "SMURF1", "SMURF2", "UCHL5", "USP15", "UBB", "UBC", "PMEPA1", "PPP1CA", "PPP1CB", "PPP1CC", "PPP1R15A")] <- 2
+state[is.na(state)] <- 1
+
+plot_directed(graph, state=state, layout = layout.kamada.kawai,
+              cex.node=3, cex.arrow=5, arrow_clip = 0.2,
+              sub = expression(paste("(a) TFG-", Beta, " activates SMADs")), cex.sub = 1.5)
+
+#adjacency matrix
+adj_mat <- make_adjmatrix_graph(graph)
+#relationship matrix
+dist_mat <- make_distance_graph(graph, absolute = FALSE)
+#sigma matrix directly from graph
+sigma_mat <- make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE)
+#show shortest paths of graph
+shortest_paths <- shortest.paths(graph)
+#generate expression data directly from graph
+expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0, comm = FALSE,
+                            dist = TRUE, absolute = FALSE)
+#plot adjacency matrix
+heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
+          col = colorpanel(3, "grey75", "white", "blue"),
+          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
+              sub = "Adjacency matrix", cex.sub = 1.5)
+# #plot relationship matrix
+heatmap.2(make_distance_graph(graph, absolute = FALSE),
+          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
+              sub = "(b) relationship matrix", cex.sub = 1.5)
+# #plot sigma matrix
+heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
+scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
+              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.5)
+expr <- generate_expression(100, graph,  state = state, cor = 0.8, mean = 0,
+comm = FALSE, dist =TRUE, absolute = FALSE)
+#plot simulated expression data
+heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
+              sub = "(e) Simulated expressinon data (log scale)", cex.sub = 1.5)
+#plot simulated correlations
+heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+```
+
 The simulation procedure

### 2020-05-01 paper/paper.Rmd
correc figure formatting for graph structures

@@
 plot_directed(graph, state ="activating", layout = layout.kamada.kawai,
-              cex.node = 2, cex.arrow = 4, arrow_clip = 0.2,
-              sub = "(a) Activating pathway structure", cex.sub = 1.5)
+              cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
+mtext(text = "(a) Activating pathway structure", side=1, line=3, at=0.05, adj=0.5, cex=1.75)
+box()
 
@@
 plot_directed(graph, state=state, layout = layout.kamada.kawai,
-              cex.node = 2, cex.arrow = 4, arrow_clip = 0.2,
-              sub = "(b) Inhibiting pathway structure", cex.sub = 1.5)
+              cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
+mtext(text = "(b) Inhibiting pathway structure", side=1, line=2, at=0.075, adj=0.5, cex=1.75)
+box()
 ```
@@
 plot_directed(graph, state=state, layout = layout.kamada.kawai,
-              cex.node=3, cex.arrow=5, arrow_clip = 0.2,
-              sub = "(a) Activating pathway structure", cex.sub = 1.5)
+              cex.node=2, cex.arrow=4, arrow_clip = 0.2,
+              sub = "(a) Activating pathway structure", cex.sub = 1.75)
+box()
 
@@
 adj_mat <- make_adjmatrix_graph(graph)
+
 #relationship matrix
 dist_mat <- make_distance_graph(graph, absolute = FALSE)
+
 #sigma matrix directly from graph
 sigma_mat <- make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE)
+
 #show shortest paths of graph
 shortest_paths <- shortest.paths(graph)
+
 #generate expression data directly from graph
@@
                             dist = TRUE, absolute = FALSE, state = state)
+
 #plot adjacency matrix
@@
           colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = "Adjacency matrix", cex.sub = 1.5)
+              sub = "Adjacency matrix", cex.sub = 1.75)
+
 #plot relationship matrix
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = "(b) relationship matrix", cex.sub = 1.5)
+              sub = "(b) relationship matrix", cex.sub = 1.75)
+
 #plot sigma matrix
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.5)
+              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.75)
+
+#plot simulated expression data
 expr <- generate_expression(100, graph, cor = 0.8, mean = 0,
 comm = FALSE, dist =TRUE, absolute = FALSE, state = state)
-#plot simulated expression data
 heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = "(e) Simulated expression data (log scale)", cex.sub = 1.5)
+              sub = "(e) Simulated expression data (log scale)", cex.sub = 1.75)
+
 #plot simulated correlations
@@
 plot_directed(graph, state=state, layout = layout.kamada.kawai,
-              cex.node=3, cex.arrow=5, arrow_clip = 0.2,
-              sub = "(a) Inhitibing pathway structure", cex.sub = 1.5)
+              cex.node=2, cex.arrow=4, arrow_clip = 0.2,
+              sub = "(a) Inhitibing pathway structure", cex.sub = 1.75)
+box()
 
@@
 adj_mat <- make_adjmatrix_graph(graph)
+
 #relationship matrix
 dist_mat <- make_distance_graph(graph, absolute = FALSE)
+
 #sigma matrix directly from graph
 sigma_mat <- make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE)
+
 #show shortest paths of graph
 shortest_paths <- shortest.paths(graph)
+
 #generate expression data directly from graph
@@
                             dist = TRUE, absolute = FALSE)
+
 #plot adjacency matrix
@@
           colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = "Adjacency matrix", cex.sub = 1.5)
+              sub = "Adjacency matrix", cex.sub = 1.75)
+
 # #plot relationship matrix
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = "(b) relationship matrix", cex.sub = 1.5)
+              sub = "(b) relationship matrix", cex.sub = 1.75)
+
 # #plot sigma matrix
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.5)
+              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.75)
+
+#plot simulated expression data
 expr <- generate_expression(100, graph,  state = state, cor = 0.8, mean = 0,
 comm = FALSE, dist =TRUE, absolute = FALSE)
-#plot simulated expression data
 heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = "(e) Simulated expressinon data (log scale)", cex.sub = 1.5)
+              sub = "(e) Simulated expression data (log scale)", cex.sub = 1.75)
+
 #plot simulated correlations
@@
 
-plot_directed(graph, state=state, layout = layout.kamada.kawai,
-              cex.node=3, cex.arrow=5, arrow_clip = 0.2,
-              sub = expression(paste("(a) TFG-", Beta, " activates SMADs")), cex.sub = 1.5)
+plot_directed(graph, state = state, layout = layout.kamada.kawai,
+              border.node=scales::alpha("black", 0.75), fill.node="lightblue",
+              col.arrow = c(scales::alpha("navyblue", 0.25), scales::alpha("red", 0.25))[state], 
+              cex.node = 1.5, cex.label = 0.8, cex.arrow = 2, 
+              sub = expression(paste("(a) TFG-", Beta, " activates SMADs")), cex.sub = 1.75)
+box()
 

### 2020-05-01 R/plot_directed.R

@@ -99,7 +99,7 @@ plot_directed <- function(graph, state = NULL, labels = NULL, layout = layout.fr
   Ne <- length(es[1]$V1)
   Xn <- L[,1]
   Yn <- L[,2]
-  plot(Xn, Yn, xaxt="n", yaxt="n", xlab=xlab, ylab=ylab, frame.plot=frame.plot, cex = 2 * cex.node, pch=1, col=par()$bg, main=main, sub=sub, cex.main=cex.main, cex.sub=cex.sub)
+  plot(Xn, Yn, xaxt="n", yaxt="n", xlab=xlab, ylab=ylab, xlim = mean(Xn)+c(-1,1 )*1.2*(max(Xn)-min(Xn))/2, ylim = mean(Yn)+c(-1,1 )*1.2*(max(Yn)-min(Yn))/2, frame.plot=frame.plot, cex = 2 * cex.node, pch=1, col=par()$bg, main=main, sub=sub, cex.main=cex.main, cex.sub=cex.sub)
   if(!is.null(get.edge.attribute(graph, "state"))){
     state <- get.edge.attribute(graph, "state")
   } else {

### 2020-05-01 paper/paper.Rmd
correct figure 2-4 formatting for graph structures

@@
               cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
-mtext(text = "(b) Inhibiting pathway structure", side=1, line=2, at=0.075, adj=0.5, cex=1.75)
+mtext(text = "(b) Inhibiting pathway structure", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
 box()
@@
 plot_directed(graph, state=state, layout = layout.kamada.kawai,
-              cex.node=2, cex.arrow=4, arrow_clip = 0.2,
-              sub = "(a) Activating pathway structure", cex.sub = 1.75)
+              cex.node=2, cex.arrow=4, arrow_clip = 0.2)
+mtext(text = "(a) Activating pathway structure", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
 box()
@@
 
-#plot adjacency matrix
-heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
-          col = colorpanel(3, "grey75", "white", "blue"),
-          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = "Adjacency matrix", cex.sub = 1.75)
-
 #plot relationship matrix
@@
           scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = "(b) relationship matrix", cex.sub = 1.75)
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+mtext(text = "(b) relationship matrix", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
 
@@
 scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.75)
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3, at=0.075, adj=0.5, cex=1.75)
 
-#plot simulated expression data
+#simulated data
 expr <- generate_expression(100, graph, cor = 0.8, mean = 0,
 comm = FALSE, dist =TRUE, absolute = FALSE, state = state)
-heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = "(e) Simulated expression data (log scale)", cex.sub = 1.75)
 
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+mtext(text = "(d) Simulated correlation", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+
+#plot simulated expression data
+heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
+              sub = "(e) Simulated expression data (log scale)", cex.sub = 1.75)
 ```
+
+The simulation procedure
+(Figure [2](#fig:simulation_activating){reference-type="ref"
+reference="fig:simulation_activating"}) can similarly be used for
+pathways containing inhibitory links
+(Figure [3](#fig:simulation_inhibiting){reference-type="ref"
+reference="fig:simulation_inhibiting"}) with several refinements. With
+the inhibitory links
+(Figure [3a](#fig:simulation_inhibiting:first){reference-type="ref"
+reference="fig:simulation_inhibiting:first"}), distances are calculated
+in the same manner as before
+(Figure [3b](#fig:simulation_inhibiting:second){reference-type="ref"
+reference="fig:simulation_inhibiting:second"}) with inhibitions
+accounted for by iteratively multiplying downstream nodes by $-1$ to
+form modules with negative correlations between them
+(Figures [3c
+](#fig:simulation_inhibiting:third){reference-type="ref"
+reference="fig:simulation_inhibiting:third"}
+and [3d](#fig:simulation_inhibiting:fifth){reference-type="ref"
+reference="fig:simulation_inhibiting:fifth"}). A multivariate normal
+distribution with these negative correlations can be sampled to generate
+simulated data
+(Figure [3e](#fig:simulation_inhibiting:fourth){reference-type="ref"
+reference="fig:simulation_inhibiting:fourth"}).
 
@@
 plot_directed(graph, state=state, layout = layout.kamada.kawai,
-              cex.node=2, cex.arrow=4, arrow_clip = 0.2,
-              sub = "(a) Inhitibing pathway structure", cex.sub = 1.75)
+              cex.node=2, cex.arrow=4, arrow_clip = 0.2)
+mtext(text = "(a) Inhibiting pathway structure", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
 box()
@@
 
-#plot adjacency matrix
-heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
-          col = colorpanel(3, "grey75", "white", "blue"),
-          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = "Adjacency matrix", cex.sub = 1.75)
-
 # #plot relationship matrix
@@
 scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.75)
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3, at=0.075, adj=0.5, cex=1.75)
 
-#plot simulated expression data
+#simulated data
 expr <- generate_expression(100, graph,  state = state, cor = 0.8, mean = 0,
 comm = FALSE, dist =TRUE, absolute = FALSE)
-heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = "(e) Simulated expression data (log scale)", cex.sub = 1.75)
 
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+mtext(text = "(d) Simulated correlation", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+
+#plot simulated expression data
+heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
 ```
 
-![\small \textbf{\textbf{Simulating expression from a biological pathway graph structure.}}
-Simulating expression from graph structure with inhibitions.}} The graph structure (a) of a
-known biological pathway, the TGF-$\beta$ receptor signaling activates SMADs (R-HSA-2173789),
-was used to derive a relationship matrix (b), $\Sigma$ matrix (c) and correlation structure
-(d) from the relative distances between the nodes. These values are coloured blue to red from
-$-1$ to $1$. This has been used to generate a simulated expression dataset of 100 samples
-(coloured blue to red from low to high) via sampling from the multivariate normal distribution.
-Here modules of genes with correlated expression can be clearly discerned.](figure4.pdf){ width=95% #fig:simulation_smad label="fig:simulation_smad"}
+The simulation procedure is also demonstrated here
+(Figure [4](#fig:simulation_smad){reference-type="ref"
+reference="fig:simulation_smad"}) on a pathway structure for a known
+biological pathway (from reactome R-HSA-2173789) of TGF-$\beta$ receptor
+signaling activates SMADs
+(Figure [4a](#fig:simulation_smad:first){reference-type="ref"
+reference="fig:simulation_smad:first"}) derived from the Reactome
+database version 52 `[@Reactome]`. Distances are calculated in the same
+manner as before
+(Figure [4b](#fig:simulation_smad:second){reference-type="ref"
+reference="fig:simulation_smad:second"}) producing blocks of correlated
+genes
+(Figures [4c](#fig:simulation_smad:third){reference-type="ref"
+reference="fig:simulation_smad:third"}
+and [4d](#fig:simulation_smad:fifth){reference-type="ref"
+reference="fig:simulation_smad:fifth"}). This shows that
+multivariate normal distribution can be sampled to generate simulated
+data to represent expression with the complexity of a biological pathway
+(Figure [4e](#fig:simulation_smad:fourth){reference-type="ref"
+reference="fig:simulation_smad:fourth"}). Here *SMAD7* exhibits
+negative correlations with the other SMADs consistent with it's
+functions as as an "inhibitor SMAD" with competitively inhibits *SMAD4*.
 
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+mtext(text = "(d) Simulated correlation", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+
+#plot simulated expression data
+heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
 ```
-
-The simulation procedure
-(Figure [2](#fig:simulation_activating){reference-type="ref"
-reference="fig:simulation_activating"}) can similarly be used for
-pathways containing inhibitory links
-(Figure [3](#fig:simulation_inhibiting){reference-type="ref"
-reference="fig:simulation_inhibiting"}) with several refinements. With
-the inhibitory links
-(Figure [3a](#fig:simulation_inhibiting:first){reference-type="ref"
-reference="fig:simulation_inhibiting:first"}), distances are calculated
-in the same manner as before
-(Figure [3b](#fig:simulation_inhibiting:second){reference-type="ref"
-reference="fig:simulation_inhibiting:second"}) with inhibitions
-accounted for by iteratively multiplying downstream nodes by $-1$ to
-form modules with negative correlations between them
-(Figures [3c
-](#fig:simulation_inhibiting:third){reference-type="ref"
-reference="fig:simulation_inhibiting:third"}
-and [3d](#fig:simulation_inhibiting:fifth){reference-type="ref"
-reference="fig:simulation_inhibiting:fifth"}). A multivariate normal
-distribution with these negative correlations can be sampled to generate
-simulated data
-(Figure [3e](#fig:simulation_inhibiting:fourth){reference-type="ref"
-reference="fig:simulation_inhibiting:fourth"}).
-
-The simulation procedure is also demonstrated here
-(Figure [4](#fig:simulation_smad){reference-type="ref"
-reference="fig:simulation_smad"}) on a pathway structure for a known
-biological pathway (from reactome R-HSA-2173789) of TGF-$\beta$ receptor
-signaling activates SMADs
-(Figure [4a](#fig:simulation_smad:first){reference-type="ref"
-reference="fig:simulation_smad:first"}) derived from the Reactome
-database version 52 `[@Reactome]`. Distances are calculated in the same
-manner as before
-(Figure [4b](#fig:simulation_smad:second){reference-type="ref"
-reference="fig:simulation_smad:second"}) producing blocks of correlated
-genes
-(Figures [4c](#fig:simulation_smad:third){reference-type="ref"
-reference="fig:simulation_smad:third"}
-and [4d](#fig:simulation_smad:fifth){reference-type="ref"
-reference="fig:simulation_smad:fifth"}). This shows that
-multivariate normal distribution can be sampled to generate simulated
-data to represent expression with the complexity of a biological pathway
-(Figure [4e](#fig:simulation_smad:fourth){reference-type="ref"
-reference="fig:simulation_smad:fourth"}). Here *SMAD7* exhibits
-negative correlations with the other SMADs consistent with it's
-functions as as an "inhibitor SMAD" with competitively inhibits *SMAD4*.
 

### 2020-05-01 paper/paper.Rmd
correct figure 2-4 formatting for heatmaps

@@
 
+```{r, include=FALSE, eval=FALSE}
+graphics::layout(matrix(c(1:5, 5), nrow = 3, byrow = TRUE))
+```
+
 ```{r simulation_activating, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}.
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = "(e) Simulated expression data (log scale)", cex.sub = 1.75)
+sub = "(e) Simulated expression data (log scale)", cex.sub = 1.75,
+xlim = c(-0.5, 1.5))
 ```

### 2020-05-08 paper/paper.Rmd
correct figure widths (fig2-4, pane E)

@@
     keep_md: TRUE
+    keep_tex: TRUE
     fig_caption: yes
@@
 
+op <- par(mfrow=c(2,2), mar=c(4,4,4,1), oma=c(0.5,0.5,0.5,0))
 #plot relationship matrix
@@
 
+op <- par(mfrow=c(1,1), mar=c(4,4,4,1), oma=c(0.5,0.5,0.5,0))
 #plot simulated expression data
 heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-sub = "(e) Simulated expression data (log scale)", cex.sub = 1.75,
-xlim = c(-0.5, 1.5))
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+par(op)
 ```

### 2020-05-08 paper/paper.tex
correct figure widths (fig2-4, pane E)

@@
-\documentclass[article, shortnames]{jss}
+\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provides euro and other symbols
+\else % if luatex or xelatex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage[]{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\usepackage{hyperref}
+\hypersetup{
+            pdftitle={graphsim: An R package for simulating gene expression data from graph structures of biological pathways},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage[margin=1in]{geometry}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-%% -- LaTeX packages and custom commands --------------
+% set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 
-%% recommended packages
-\usepackage{thumbpdf,lmodern}
+\usepackage{caption}
 
-%% another package (only for this demo article)
-\usepackage{framed}
-\usepackage{subcaption} %continuefloat
-\usepackage{xfrac}
-\usepackage{mathtools}
-\usepackage{natbib}
+\title{graphsim: An R package for simulating gene expression data from graph
+structures of biological pathways}
+\author{}
+\date{\vspace{-2.5em}30 April 2020}
 
-\usepackage[acronym, xindy, toc, acronym, style=super,  nogroupskip, nonumberlist, nopostdot]{glossaries}
-\setlength{\glsdescwidth}{0.75\textwidth}
-\include{glossary}
-\makeglossaries
+\begin{document}
+\maketitle
 
-\usepackage{placeins}
+\hypertarget{summary}{%
+\subsubsection{Summary}\label{summary}}
 
-%\usepackage[colorlinks,
-%           linkcolor={black},
-%           citecolor={black},
-%           urlcolor={blue!80!white}]{hyperref}
-           %allbordercolors={0 0 0},
-           %pdfborderstyle={/S/U/W 1}
+Transcriptomic analysis is used to capture the molecular state of a cell
+or sample in many biological and medical applications. In addition to
+identifying alterations in activity at the level of individual genes,
+understanding changes in the gene networks that regulate fundamental
+biological mechanisms is also an important objective of molecular
+analysis. As a result, databases that describe biological pathways are
+increasingly relied on to assist with the interpretation of results from
+large-scale genomics studies. Incorporating information from biological
+pathways and gene regulatory networks into a genomic data analysis is a
+popular strategy, and there are many methods that provide this
+functionality for gene expression data. When developing or comparing
+such methods, it is important to gain an accurate assessment of their
+performance, with simulation-based validation studies a popular choice.
+This necessitates the use of simulated data that correctly accounts for
+pathway relationships and correlations. Here we present a versatile
+statistical framework to simulate correlated gene expression data from
+biological pathways, by sampling from a multivariate normal distribution
+derived from a graph structure. This procedure has been released as the
+\texttt{graphsim} R package on CRAN and GitHub
+(\url{https://github.com/TomKellyGenetics/graphsim}) and is compatible
+with any graph structure that can be described using the \texttt{igraph}
+package. This package allows the simulation of biological pathways from
+a graph structure based on a statistical model of gene expression.
 
-%% new custom commands
-\newcommand{\class}[1]{`\code{#1}'}
-\newcommand{\fct}[1]{\code{#1()}}
+\hypertarget{sec:intro}{%
+\section{Introduction: inference and modelling of biological
+networks}\label{sec:intro}}
 
-%% - Article metainformation (author, title, ...) ----------
+Network analysis of molecular biological pathways has the potential to
+lead to new insights into biology and medical genetics (Barabási and
+Oltvai 2004; Hu, Thomas, and Brunak 2016). Since gene expression
+profiles capture a consistent signature of the regulatory state of a
+cell (Perou et al. 2000; Ozsolak and Milos 2011; Svensson, Vento-Tormo,
+and Teichmann 2018), they can be used to analyse complex molecular
+states with genome-scale data. However, biological pathways are often
+analysed in a reductionist paradigm as amorphous sets of genes involved
+in particular functions, despite the fact that the relationships defined
+by pathway structure could further inform gene expression analyses. In
+many cases, the pathway relationships are well-defined,
+experimentally-validated, and are available in public databases (Croft
+et al. 2014). As a result, network analysis techniques could play an
+important role in furthering our understanding of biological pathways
+and aiding in the interpretation of genomics studies.
 
-%% - \author{} with primary affiliation
-%% - \Plainauthor{} without affiliations
-%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
-%% - \AND starts a new line, \And does not.
-\author{S. Thomas Kelly\\University of Otago
-   \And Michael A. Black\\University of Otago}
-\Plainauthor{Thomas Kelly, Michael Black}
+Gene networks provide insights into how cells are regulated, by mapping
+regulatory interactions between target genes and transcription factors,
+enhancers, and sites of epigenetic marks or chromatin structures
+(Barabási and Oltvai 2004; Yamaguchi et al. 2007). Inference of these
+regulatory interactions for genomics investigations has the potential to
+radically expand the range of candidate biological pathways to be
+further explored, or to improve the accuracy of bioinformatics and
+functional genomic analysis. A number of methods have already been
+developed to utilise timecourse gene expression data (Arner et al. 2015;
+Yamaguchi et al. 2007) using gene regulatory modules in state-space
+models and recursive vector autoregressive models (Hirose et al. 2008;
+Shimamura et al. 2009). Various approaches to gene regulation and
+networks at the genome-wide scale have led to novel biological insights
+(Arner et al. 2015; Komatsu et al. 2013). However, inference of
+regulatory networks has thus far relied on experimental validation or
+resampling-based approaches to estimate the likelihood of specific
+network modules being predicted (Markowetz and Spang 2007; Hawe, Theis,
+and Heinig 2019).
 
-%% - \title{} in title case
-%% - \Plaintitle{} without LaTeX markup (if any)
-%% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{\pkg{graphsim}: An \proglang{R} package for simulating gene expression data from graph structures of biological pathways}
-\Plaintitle{graphsim: An R package for simulating gene expression data from graph structures of biological pathways}
-\Shorttitle{Simulating gene expression from pathways in R}
+There is a need, therefore, for a systematic framework for statistical
+modelling and simulation of gene expression data derived from
+hypothetical, inferred or known gene networks. Here we present a package
+to achieve this, where samples from a multivariate normal distribution
+are used to generate normally-distributed log-expression data, with
+correlations between genes derived from the structure of the underlying
+pathway or gene regulatory network. This methodology enables simulation
+of expression profiles that approximate the log-transformed and
+normalised data from microarray and bulk or single-cell RNA-Seq
+experiments. This procedure has been released as the package to enable
+the generation of simulated gene expression datasets containing pathway
+relationships from a known underlying network. These simulated datasets
+can be used to evaluate various bioinformatics methodologies, including
+statistical and network inference procedures.
 
-%% - \Abstract{} almost as usual
-\Abstract{
-Transcriptomic analysis is used to capture the molecular state of a cell or sample in many biological and medical applications.
-In addition to identifying alterations in activity at the level of individual genes, understanding changes in the gene networks that regulate fundamental biological mechanisms is also an important objective of molecular analysis.
-As a result, databases that describe biological pathways are increasingly relied on to assist with the interpretation of results from large-scale genomics studies.
-Incorporating information from biological pathways and gene regulatory networks into a genomic data analysis is a popular strategy, and there are many methods that provide this functionality for gene expression data.
-When developing or comparing such methods, it is important to gain an accurate assessment of their performance, with simulation-based validation studies a popular choice.
-This necessitates the use of simulated data that correctly accounts for pathway relationships and correlations.
-Here we present a versatile statistical framework to simulate correlated gene expression data from biological pathways, by sampling from a multivariate normal distribution derived from a graph structure.
-This procedure has been released as the \pkg{graphsim} \proglang{R} package (\url{https://github.com/TomKellyGenetics/graphsim}) and is compatible with any graph structure that can be described using the \pkg{igraph} package.
+\hypertarget{sec:methods}{%
+\section{Methodology and software}\label{sec:methods}}
+
+Here we present a procedure to simulate gene expression data with
+correlation structure derived from a known graph structure. This
+procedure assumes that transcriptomic data have been generated and
+follow a log-normal distribution (i.e.,
+\(log(X_{ij}) \sim MVN({\bf\mu}, \Sigma)\), where \({\bf\mu}\) and
+\(\Sigma\) are the mean vector and variance-covariance matrix
+respectively, for gene expression data derived from a biological
+pathway) after appropriate normalisation (Law et al. 2014; Li et al.
+2015). Log-normality of gene expression matches the assumptions of the
+popular \texttt{limma} package (Ritchie et al. 2015), which is often
+used for the analysis of intensity-based data from gene expression
+microarray studies and count-based data from RNA-Seq experiments. This
+approach has also been applied for modelling UMI-based count data from
+single-cell RNA-Seq experiments in the package (Wang et al. 2018).
+
+In order to simulate transcriptomic data, a pathway is first constructed
+as a graph structure, using the \texttt{igraph} R package (Csardi and
+Nepusz 2006), with the status of the edge relationships defined (i.e,
+whether they activate or inhibit downstream pathway members). {This
+procedure uses} a graph structure such as that presented in
+Figure~\protect\hyperlink{fig:simple_graph:first}{1a}. The graph can be
+defined by an adjacency matrix, \textbf{\(A\)} (with elements
+\(A_{ij}\)), where \[A_{ij} = 
+\begin{cases}
+   1                         & \mbox{if genes } i \mbox{ and } j \mbox{ are adjacent} \\
+   0                         & \mbox{otherwise}
+\end{cases}\]
+
+A matrix, \textbf{\(R\)}, with elements {\(R_{ij}\)}, is calculated
+based on distance (i.e., number of edges contained in the shortest path)
+between nodes, such that closer nodes are given more weight than more
+distant nodes, to define inter-node relationships. A
+geometrically-decreasing (relative) distance weighting is used to
+achieve this:
+
+{\[R_{ij} = 
+\begin{cases}
+   1                            & \mbox{if genes } i \mbox{ and } j \mbox{ are adjacent} \\
+   (\frac{1}{2})^{d_{ij}}      & \mbox{if a path can be found between genes } i \mbox{ and } j \\
+   0                            & \mbox{if no path exists between genes } i \mbox{ and } j 
+\end{cases}\]} where \(d_{ij}\) is the length of the shortest path
+(i.e., minimum number of edges traversed) between genes (nodes) \(i\)
+and \(j\) in graph \(G\). Each more distant node is thus related by
+\(\frac{1}{2}\) compared to the next nearest, as shown in
+Figure~\protect\hyperlink{fig:simulation_activating:second}{2b}. An
+arithmetically-decreasing (absolute) distance weighting is also
+supported in the \texttt{graphsim} R package which implements this
+procedure: {\[R_{ij} = 
+\begin{cases}
+   1                            & \mbox{if genes } i \mbox{ and } j \mbox{ are adjacent} \\
+   1-\frac{d_{ij}}{diam(G)}     & \mbox{if a path can be found between genes } i \mbox{ and } j \\
+   0                            & \mbox{if no path exists between genes } i \mbox{ and } j 
+\end{cases}\]}
+
+Assuming a unit variance for each gene, these values can be used to
+derive a \(\Sigma\) matrix: \[\Sigma_{ij} = 
+\begin{cases}
+   1                            & \mbox{if } i=j \\
+   \rho R_{ij}  & \mbox{otherwise}
+\end{cases}\] where \(\rho\) is the correlation between adjacent nodes.
+Thus covariances between adjacent nodes are assigned by a correlation
+parameter (\(\rho\)) and the remaining off-diagonal values in the matrix
+are based on scaling these correlations by the geometrically weighted
+relationship matrix (or the nearest positive definite matrix for
+\(\Sigma\) with negative correlations).\\
+Computing the nearest positive definite matrix is necessary to ensure
+that the variance-covariance matrix could be inverted when used as a
+parameter in multivariate normal simulations, particularly when negative
+correlations are included for inhibitions (as shown below). Matrices
+that could not be inverted occurred rarely with biologically plausible
+graph structures but this approach allows for the computation of a
+plausible correlation matrix when the given graph structure is
+incomplete or contains loops. When required, the nearest positive
+definite matrix is computed using the \texttt{nearPD} function of the
+\texttt{Matrix} R package (Bates and Maechler 2016) to perform Higham's
+algorithm (Higham 2002) on variance-covariance matrices. The
+\texttt{graphsim} package gives a warning when this occurs.
+
+\hypertarget{sec:illustrations}{%
+\section{Illustrations}\label{sec:illustrations}}
+
+\hypertarget{sec:plot_graph}{%
+\subsection{Generating a Graph Structure}\label{sec:plot_graph}}
+
+The graph structure in
+Figure~\protect\hyperlink{fig:simple_graph:first}{1a} was used to
+simulate correlated gene expression data by sampling from a multivariate
+normal distribution using the R package (Genz and Bretz 2009; Genz et
+al. 2016). The graph structure visualisation in
+Figure~\protect\hyperlink{fig:simple_graph}{1} was specifically
+developed for (directed) \texttt{igraph} objects in and is available in
+the and packages. The \textbackslash{}texttt\{plot\_directed\} function
+enables customisation of plot parameters for each node or edge, and
+mixed (directed) edge types for indicating activation or inhibition.
+These inhibition links (which occur frequently in biological pathways)
+are demonstrated in
+Figure~\protect\hyperlink{fig:simple_graph:second}{1b}.
+
+A graph structure can be generated and plotted using the following
+commands in R:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{#install packages required (once per machine)}
+
+\KeywordTok{install.packages}\NormalTok{(}\StringTok{"graphsim"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{#load required packages (once per R instance)}
+
+\KeywordTok{library}\NormalTok{(}\StringTok{"igraph"}\NormalTok{)}
+\CommentTok{#> }
+\CommentTok{#> Attaching package: 'igraph'}
+\CommentTok{#> The following objects are masked from 'package:stats':}
+\CommentTok{#> }
+\CommentTok{#>     decompose, spectrum}
+\CommentTok{#> The following object is masked from 'package:base':}
+\CommentTok{#> }
+\CommentTok{#>     union}
+\KeywordTok{library}\NormalTok{(}\StringTok{"graphsim"}\NormalTok{)}
+\KeywordTok{library}\NormalTok{(}\StringTok{"gplots"}\NormalTok{)}
+\CommentTok{#> }
+\CommentTok{#> Attaching package: 'gplots'}
+\CommentTok{#> The following object is masked from 'package:stats':}
+\CommentTok{#> }
+\CommentTok{#>     lowess}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{#generate graph structure}
+
+\NormalTok{graph_edges <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\StringTok{"C"}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\StringTok{"B"}\NormalTok{, }\StringTok{"C"}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\StringTok{"C"}\NormalTok{, }\StringTok{"D"}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\StringTok{"D"}\NormalTok{, }\StringTok{"E"}\NormalTok{),}
+                     \KeywordTok{c}\NormalTok{(}\StringTok{"D"}\NormalTok{, }\StringTok{"F"}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\StringTok{"F"}\NormalTok{, }\StringTok{"G"}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\StringTok{"F"}\NormalTok{, }\StringTok{"I"}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\StringTok{"H"}\NormalTok{, }\StringTok{"I"}\NormalTok{))}
+\NormalTok{graph <-}\StringTok{ }\KeywordTok{graph.edgelist}\NormalTok{(graph_edges, }\DataTypeTok{directed =} \OtherTok{TRUE}\NormalTok{)}
+
+\CommentTok{#plot graph structure (Figure 1a)}
+
+\KeywordTok{plot_directed}\NormalTok{(graph, }\DataTypeTok{state =}\StringTok{"activating"}\NormalTok{, }\DataTypeTok{layout =}\NormalTok{ layout.kamada.kawai,}
+              \DataTypeTok{cex.node =} \DecValTok{2}\NormalTok{, }\DataTypeTok{cex.arrow =} \DecValTok{4}\NormalTok{, }\DataTypeTok{arrow_clip =} \FloatTok{0.2}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(a) Activating pathway structure"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.05}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{box}\NormalTok{()}
+
+\CommentTok{#generate parameters for inhibitions}
+
+\NormalTok{state <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{-1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{-1}\NormalTok{, }\DecValTok{1}\NormalTok{)}
+
+\CommentTok{#plot graph structure with inhibitions (Figure 1b)}
+
+\KeywordTok{plot_directed}\NormalTok{(graph, }\DataTypeTok{state=}\NormalTok{state, }\DataTypeTok{layout =}\NormalTok{ layout.kamada.kawai,}
+              \DataTypeTok{cex.node =} \DecValTok{2}\NormalTok{, }\DataTypeTok{cex.arrow =} \DecValTok{4}\NormalTok{, }\DataTypeTok{arrow_clip =} \FloatTok{0.2}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(b) Inhibiting pathway structure"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{box}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+
+{\centering \includegraphics[width=.415\linewidth,height=.415\linewidth]{Plotsimple_graph-1} \includegraphics[width=.415\linewidth,height=.415\linewidth]{Plotsimple_graph-2}
+
 }
 
-%% - \Keywords{} with LaTeX markup, at least one required
-%% - \Plainkeywords{} without LaTeX markup (if necessary)
-%% - Should be comma-separated and in sentence case.
-\Keywords{gene-expression, simulation, genomics, pathway, network, \proglang{R}}
-\Plainkeywords{gene-expression, simulation, genomics, pathway, network, R}
+\caption{\textbf{Simulated graph structures}.
+A constructed graph structure used as an example to demonstrate the simulation procedure in Figures 2 and 3.
+Activating links are denoted by black arrows and inhibiting links by red edges.
+Inhibiting edges have been highlighted in red.}\label{fig:simple_graph}
+\end{figure}
 
-%% - \Address{} of at least one author
-%% - May contain multiple affiliations for each author
-%%   (in extra lines, separated by \emph{and}\\).
-%% - May contain multiple authors for the same affiliation
-%%   (in the same first line, separated by comma).
-\Address{
-  S. Thomas Kelly\\
-  Division of Genomic Medicine
-  RIKEN Center for Integrative Medical Sciences\\
-  Yokohama, Japan \\
-  E-mail: \email{tom.kelly@riken.jp}\\
-  \emph{and}\\
-  Department of Biochemistry\\
-  University of Otago\\
-  Dunedin, New Zealand\\
-  %E-mail: \email{tom.kelly@postgrad.otago.ac.nz}%\\
-  %URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
+\hypertarget{sec:graphsim_demo}{%
+\subsection{Generating a Simulated Expression
+Dataset}\label{sec:graphsim_demo}}
 
-  Michael Black\\
-  Department of Biochemistry\\
-  University of Otago\\
-  PO Box 56\\
-  Dunedin, New Zealand\\
-  E-mail: \email{mik.black@otago.ac.nz}%\\
-  %URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
+The correlation parameter of \(\rho = 0.8\) is used to demonstrate the
+inter-correlated datasets using a geometrically-generated relationship
+matrix (as used for the example in
+Figure~\protect\hyperlink{fig:simulation_activating:third}{2c}). This
+\(\Sigma\) matrix was then used to sample from a multivariate normal
+distribution such that each gene had a mean of \(0\), standard deviation
+\(1\), and covariance within the range \([0,1]\) so that the
+off-diagonal elements of \(\Sigma\) represent correlations. This
+procedure generated a simulated (continuous normally-distributed)
+log-expression profile for each node
+(Figure~\protect\hyperlink{fig:simulation_activating:fourth}{2e}) with a
+corresponding correlation structure
+(Figure~\protect\hyperlink{fig:simulation_activating:fifth}{2d}). The
+simulated correlation structure closely resembled the expected
+correlation structure (\(\Sigma\) in
+Figure~\protect\hyperlink{fig:simulation_activating:third}{2c}) even for
+the relatively modest sample size (\(N=100\)) illustrated in
+Figure~\protect\hyperlink{fig:simulation_activating}{2}. Once a gene
+expression dataset comprising multiple pathways has been generated (as
+in Figure~\protect\hyperlink{fig:simulation_activating:fourth}{2e}), it
+can then be used to test procedures designed for analysis of empirical
+gene expression data (such as those generated by microarrays or RNA-Seq)
+that have been normalised on a log-scale.
+
+The simulated dataset can be generated using the following code:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# activating graph}
+\NormalTok{state <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DecValTok{1}\NormalTok{, }\KeywordTok{length}\NormalTok{(}\KeywordTok{E}\NormalTok{(graph)))}
+\KeywordTok{plot_directed}\NormalTok{(graph, }\DataTypeTok{state=}\NormalTok{state, }\DataTypeTok{layout =}\NormalTok{ layout.kamada.kawai,}
+              \DataTypeTok{cex.node=}\DecValTok{2}\NormalTok{, }\DataTypeTok{cex.arrow=}\DecValTok{4}\NormalTok{, }\DataTypeTok{arrow_clip =} \FloatTok{0.2}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(a) Activating pathway structure"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{box}\NormalTok{()}
+
+\CommentTok{#adjacency matrix}
+\NormalTok{adj_mat <-}\StringTok{ }\KeywordTok{make_adjmatrix_graph}\NormalTok{(graph)}
+
+\CommentTok{#relationship matrix}
+\NormalTok{dist_mat <-}\StringTok{ }\KeywordTok{make_distance_graph}\NormalTok{(graph, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{)}
+
+\CommentTok{#sigma matrix directly from graph}
+\NormalTok{sigma_mat <-}\StringTok{ }\KeywordTok{make_sigma_mat_dist_graph}\NormalTok{(graph, }\FloatTok{0.8}\NormalTok{, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{)}
+
+\CommentTok{#show shortest paths of graph}
+\NormalTok{shortest_paths <-}\StringTok{ }\KeywordTok{shortest.paths}\NormalTok{(graph)}
+
+\CommentTok{#generate expression data directly from graph}
+\NormalTok{expr <-}\StringTok{ }\KeywordTok{generate_expression}\NormalTok{(}\DecValTok{100}\NormalTok{, graph, }\DataTypeTok{cor =} \FloatTok{0.8}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{comm =} \OtherTok{FALSE}\NormalTok{,}
+                            \DataTypeTok{dist =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{state =}\NormalTok{ state)}
+\CommentTok{#> Warning in generate_expression(100, graph, cor = 0.8, mean = 0,}
+\CommentTok{#> comm = FALSE, : sigma matrix was not positive definite, nearest}
+\CommentTok{#> approximation used.}
+
+\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow=}\KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{), }\DataTypeTok{mar=}\KeywordTok{c}\NormalTok{(}\DecValTok{4}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{1}\NormalTok{), }\DataTypeTok{oma=}\KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{,}\FloatTok{0.5}\NormalTok{,}\FloatTok{0.5}\NormalTok{,}\DecValTok{0}\NormalTok{))}
+\CommentTok{#plot relationship matrix}
+\KeywordTok{heatmap.2}\NormalTok{(}\KeywordTok{make_distance_graph}\NormalTok{(graph, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{),}
+          \DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{colorpanel}\NormalTok{(}\DecValTok{50}\NormalTok{, }\StringTok{"white"}\NormalTok{, }\StringTok{"red"}\NormalTok{),}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(b) relationship matrix"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+
+\CommentTok{#plot sigma matrix}
+\KeywordTok{heatmap.2}\NormalTok{(}\KeywordTok{make_sigma_mat_dist_graph}\NormalTok{(graph, }\FloatTok{0.8}\NormalTok{, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{),}
+\DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{colorpanel}\NormalTok{(}\DecValTok{50}\NormalTok{, }\StringTok{"white"}\NormalTok{, }\StringTok{"red"}\NormalTok{),}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"(c) "}\NormalTok{, Sigma, }\StringTok{" matrix"}\NormalTok{)), }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+
+\CommentTok{#simulated data}
+\NormalTok{expr <-}\StringTok{ }\KeywordTok{generate_expression}\NormalTok{(}\DecValTok{100}\NormalTok{, graph, }\DataTypeTok{cor =} \FloatTok{0.8}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{,}
+\DataTypeTok{comm =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{dist =}\OtherTok{TRUE}\NormalTok{, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{state =}\NormalTok{ state)}
+\CommentTok{#> Warning in generate_expression(100, graph, cor = 0.8, mean = 0,}
+\CommentTok{#> comm = FALSE, : sigma matrix was not positive definite, nearest}
+\CommentTok{#> approximation used.}
+
+\CommentTok{#plot simulated correlations}
+\KeywordTok{heatmap.2}\NormalTok{(}\KeywordTok{cor}\NormalTok{(}\KeywordTok{t}\NormalTok{(expr)), }\DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{colorpanel}\NormalTok{(}\DecValTok{50}\NormalTok{, }\StringTok{"white"}\NormalTok{, }\StringTok{"red"}\NormalTok{),}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(d) Simulated correlation"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+
+\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow=}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{), }\DataTypeTok{mar=}\KeywordTok{c}\NormalTok{(}\DecValTok{4}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{1}\NormalTok{), }\DataTypeTok{oma=}\KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{,}\FloatTok{0.5}\NormalTok{,}\FloatTok{0.5}\NormalTok{,}\DecValTok{0}\NormalTok{))}
+\CommentTok{#plot simulated expression data}
+\KeywordTok{heatmap.2}\NormalTok{(expr, }\DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{bluered}\NormalTok{(}\DecValTok{50}\NormalTok{),}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(e) Simulated expression data (log scale)"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{par}\NormalTok{(op)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+
+{\centering \includegraphics[width=.415\linewidth,height=.415\linewidth]{Plotsimulation_activating-1} \includegraphics[width=.415\linewidth,height=.415\linewidth]{Plotsimulation_activating-2} \includegraphics[width=.415\linewidth,height=.415\linewidth]{Plotsimulation_activating-3} \includegraphics[width=.415\linewidth,height=.415\linewidth]{Plotsimulation_activating-4} \includegraphics[width=.830\linewidth,height=.415\linewidth]{Plotsimulation_activating-5}
+
 }
 
-\begin{document}
+\caption{\textbf{Simulating expression from a graph structure}.
+An example of a graph structure (a) that has been used to derive a relationship matrix (b), $\Sigma$  matrix (c) and correlation structure (d) from the relative distances between the nodes.
+Non-negative values are coloured white to red from $0$ to $1$.
+This $\Sigma$ matrix has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
+Here genes with closer relationships in the pathway structure show a higher correlation between simulated values.}\label{fig:simulation_activating}
+\end{figure}
 
-%% - Introduction ---------------------
+The simulation procedure
+(Figure~\protect\hyperlink{fig:simulation_activating}{2}) can similarly
+be used for pathways containing inhibitory links
+(Figure~\protect\hyperlink{fig:simulation_inhibiting}{3}) with several
+refinements. With the inhibitory links
+(Figure~\protect\hyperlink{fig:simulation_inhibiting:first}{3a}),
+distances are calculated in the same manner as before
+(Figure~\protect\hyperlink{fig:simulation_inhibiting:second}{3b}) with
+inhibitions accounted for by iteratively multiplying downstream nodes by
+\(-1\) to form modules with negative correlations between them
+(Figures~\protect\hyperlink{fig:simulation_inhibiting:third}{3c}
+and~\protect\hyperlink{fig:simulation_inhibiting:fifth}{3d}). A
+multivariate normal distribution with these negative correlations can be
+sampled to generate simulated data
+(Figure~\protect\hyperlink{fig:simulation_inhibiting:fourth}{3e}).
 
-%% - In principle "as usual".
-%% - But should typically have some discussion of both _software_ and _methods_.
-%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
-%% - If such markup is in (sub)section titles, a plain text version has to be
-%%   added as well.
-%% - All software mentioned should be properly \cite-d.
-%% - All abbreviations should be introduced.
-%% - Unless the expansions of abbreviations are proper names (like "Journal
-%%   of Statistical Software" above) they should be in sentence case (like
-%%   "generalized linear models" below).
+\begin{Shaded}
+\begin{Highlighting}[]
 
-\section[Introduction: inference and modelling of biological networks]{Introduction: inference and modelling of biological networks} \label{sec:intro}
+\CommentTok{#generate parameters for inhibitions}
+\NormalTok{state <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{-1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{-1}\NormalTok{)}
+\KeywordTok{plot_directed}\NormalTok{(graph, }\DataTypeTok{state=}\NormalTok{state, }\DataTypeTok{layout =}\NormalTok{ layout.kamada.kawai,}
+              \DataTypeTok{cex.node=}\DecValTok{2}\NormalTok{, }\DataTypeTok{cex.arrow=}\DecValTok{4}\NormalTok{, }\DataTypeTok{arrow_clip =} \FloatTok{0.2}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(a) Inhibiting pathway structure"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{box}\NormalTok{()}
 
-\shortcites{Perou2000, Arner2015, Reactome, Komatsu2013, Yamaguchi2007, Shimamura2009, Hirose2008, Li2015, Genz2016}
+\CommentTok{#adjacency matrix}
+\NormalTok{adj_mat <-}\StringTok{ }\KeywordTok{make_adjmatrix_graph}\NormalTok{(graph)}
 
-Network analysis of molecular biological pathways has the potential to lead to new insights into biology and medical genetics \citep{Barabasi2004, Hu2016}.
-Since gene expression profiles capture a consistent signature of the regulatory state of a cell \citep{Perou2000, Ozsolak2011, Svensson2018}, they can be used to analyse complex molecular states with genome-scale data.
-However, biological pathways are often analysed in a reductionist paradigm as amorphous sets of genes involved in particular functions, despite the fact that the relationships defined by pathway structure could further inform gene expression analyses.
-In many cases, the pathway relationships are well-defined, experimentally-validated, and are available in public databases \citep{Reactome}.
-%While pathway analysis over-representation analysis using the hypergeometric test is an effective approach to identify enriched functional groups \citep{Mooney2015, Werner2008}, it is limited by the accuracy and coverage of the biological databases used.
-Furthermore, pathway enrichment does not utilise the pathway structure information that has been gathered for many pathways and is supported by databases such as Reactome \citep{Reactome}.
-As a result, network analysis techniques could play an important role in furthering our understanding of biological pathways and aiding in the interpretation of genomics studies.
-%Furthermore with modelling and simulation procedures to evaluate them are needed.
+\CommentTok{#relationship matrix}
+\NormalTok{dist_mat <-}\StringTok{ }\KeywordTok{make_distance_graph}\NormalTok{(graph, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{)}
 
-Gene networks provide insights into how cells are regulated, by mapping regulatory interactions between target genes and transcription factors, enhancers, and sites of epigenetic marks or chromatin structures \citep{Barabasi2004, Yamaguchi2007}.
-Inference of these regulatory interactions for genomics investigations has the potential to radically expand the range of candidate biological pathways to be further explored, or to improve the accuracy of bioinformatics and functional genomic analysis.
-%which relies on these pathways being curated.
-A number of methods have already been developed to utilise timecourse gene expression data \citep{Arner2015, Yamaguchi2007} using gene regulatory modules in state-space models and recursive vector autoregressive models \citep{Hirose2008, Shimamura2009}.
-Various approaches to gene regulation and networks at the genome-wide scale have lead to novel biological insights \citep{Arner2015, Komatsu2013}.
-However, inference of regulatory networks has thus far relied on experimental validation or resampling-based approaches to estimate the likelihood of specific network modules being predicted \citep{Markowetz2007,Hawe2019}.
+\CommentTok{#sigma matrix directly from graph}
+\NormalTok{sigma_mat <-}\StringTok{ }\KeywordTok{make_sigma_mat_dist_graph}\NormalTok{(graph, }\DataTypeTok{state =}\NormalTok{ state, }\FloatTok{0.8}\NormalTok{, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{)}
 
-There is a need, therefore, for a systematic framework for statistical modelling and simulation of gene expression data derived from hypothetical, inferred or known gene networks.
-Here we present an \proglang{R} package to achieve this, where samples from a multivariate normal distribution are used to generate normally-distributed log-expression data, with correlations between genes derived from the structure of the underlying pathway or gene regulatory network.
-This methodology enables simulation of expression profiles that approximate the log-transformed and normalised data from microarray and bulk or single-cell RNA-Seq experiments.
-This procedure has been released as the \pkg{graphsim} \proglang{R} package to enable the generation of simulated gene expression datasets containing pathway relationships from a known underlying network.
-These simulated datasets can be used to evaluate various bioinformatics methodologies, including statistical and network inference procedures.
+\CommentTok{#show shortest paths of graph}
+\NormalTok{shortest_paths <-}\StringTok{ }\KeywordTok{shortest.paths}\NormalTok{(graph)}
 
-%This procedure enables modelling of genes within a biological pathway and the investigation of the impact of pathway structure on identification of biomarkers from gene expression analysis techniques.
-It was developed for biological pathway members with correlated gene expression (higher than the background of genes in other pathways) but it may also be applicable to modelling protein levels (e.g, in a kinase regulation cascade) or substrates and products (e.g., in a metabolic pathway).
-The simulation procedure was designed to use graph structures (such as those shown in Figure~\ref{fig:simple_graph}) to inform development of a ``Sigma'' variance-covariance matrix ($\Sigma$) for sampling from a multivariate normal distribution.
-% (using the \pkg{mvtnorm} \proglang{R} package).
+\CommentTok{#generate expression data directly from graph}
+\NormalTok{expr <-}\StringTok{ }\KeywordTok{generate_expression}\NormalTok{(}\DecValTok{100}\NormalTok{, graph, }\DataTypeTok{state =}\NormalTok{ state, }\DataTypeTok{cor =} \FloatTok{0.8}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{comm =} \OtherTok{FALSE}\NormalTok{,}
+                            \DataTypeTok{dist =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{)}
+\CommentTok{#> Warning in generate_expression(100, graph, state = state, cor =}
+\CommentTok{#> 0.8, mean = 0, : sigma matrix was not positive definite, nearest}
+\CommentTok{#> approximation used.}
 
-\iffalse
-\begin{leftbar}
-The introduction is in principle ``as usual''. However, it should usually embed
-both the implemented \emph{methods} and the \emph{software} into the respective
-relevant literature. For the latter both competing and complementary software
-should be discussed (within the same software environment and beyond), bringing
-out relative (dis)advantages. All software mentioned should be properly
-\verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
-\textsc{Bib}{\TeX}.)
+\CommentTok{# #plot relationship matrix}
+\KeywordTok{heatmap.2}\NormalTok{(}\KeywordTok{make_distance_graph}\NormalTok{(graph, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{),}
+          \DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{colorpanel}\NormalTok{(}\DecValTok{50}\NormalTok{, }\StringTok{"white"}\NormalTok{, }\StringTok{"red"}\NormalTok{),}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(b) relationship matrix"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 
-For writing about software JSS requires authors to use the markup
-\verb|\proglang{}| (programming languages and large programmable systems),
-\verb|\pkg{}| (software packages), \verb|\code{}| (functions, commands,
-arguments, etc.). If there is such markup in (sub)section titles (as above), a
-plain text version has to be provided in the {\LaTeX} command as well. Below we
-also illustrate how abbreviations should be introduced and citation commands can
-be employed. See the {\LaTeX} code for more details.
-\end{leftbar}
+\CommentTok{# #plot sigma matrix}
+\KeywordTok{heatmap.2}\NormalTok{(}\KeywordTok{make_sigma_mat_dist_graph}\NormalTok{(graph, }\DataTypeTok{state =}\NormalTok{ state, }\FloatTok{0.8}\NormalTok{, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{),}
+\DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{colorpanel}\NormalTok{(}\DecValTok{50}\NormalTok{, }\StringTok{"blue"}\NormalTok{, }\StringTok{"white"}\NormalTok{, }\StringTok{"red"}\NormalTok{),}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"(c) "}\NormalTok{, Sigma, }\StringTok{" matrix"}\NormalTok{)), }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 
-Modeling count variables is a common task in economics and the social sciences.
-The classical Poisson regression model for count data is often of limited use in
-these disciplines because empirical count data sets typically exhibit
-overdispersion and/or an excess number of zeros. The former issue can be
-addressed by extending  the plain Poisson regression model in various
-directions: e.g., using sandwich covariances or estimating an additional
-dispersion parameter (in a so-called quasi-Poisson model). Another more formal
-way is to use a negative binomial (NB) regression. All of these models belong to
-the family of generalized linear models (GLMs). However, although these models
-typically can capture overdispersion rather well, they are in many applications
-not sufficient for  modeling excess zeros. Since \citep{Mullahy:1986} there is
-increased interest in zero-augmented models that address this issue by a second
-model component capturing zero counts. An overview of count data models in
-econometrics, including  hurdle and zero-inflated models, is provided in
-\citep{Cameron+Trivedi:2013}.
+\CommentTok{#simulated data}
+\NormalTok{expr <-}\StringTok{ }\KeywordTok{generate_expression}\NormalTok{(}\DecValTok{100}\NormalTok{, graph,  }\DataTypeTok{state =}\NormalTok{ state, }\DataTypeTok{cor =} \FloatTok{0.8}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{,}
+\DataTypeTok{comm =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{dist =}\OtherTok{TRUE}\NormalTok{, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{)}
+\CommentTok{#> Warning in generate_expression(100, graph, state = state, cor =}
+\CommentTok{#> 0.8, mean = 0, : sigma matrix was not positive definite, nearest}
+\CommentTok{#> approximation used.}
 
-In \proglang{R} \citep{R}, GLMs are provided by the model fitting functions
-\fct{glm} in the \pkg{stats} package and \fct{glm.nb} in the \pkg{MASS} package
-\citep[][Chapter~7.4]{Venables+Ripley:2002} along with associated methods for
-diagnostics and inference. The manuscript that this document is based on
-\citep{Zeileis+Kleiber+Jackman:2008} then introduced hurdle and zero-inflated
-count models in the functions \fct{hurdle} and \fct{zeroinfl} in the \pkg{pscl}
-package \citep{Jackman:2015}. Of course, much more software could be discussed
-here, including (but not limited to) generalized additive models for count data
-as available in the \proglang{R} packages \pkg{mgcv} \citep{Wood:2006}, 
-\pkg{gamlss} \citep{Stasinopoulos+Rigby:2007}, or \pkg{VGAM} \citep{Yee:2009}.
-\fi
+\CommentTok{#plot simulated correlations}
+\KeywordTok{heatmap.2}\NormalTok{(}\KeywordTok{cor}\NormalTok{(}\KeywordTok{t}\NormalTok{(expr)), }\DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{colorpanel}\NormalTok{(}\DecValTok{50}\NormalTok{, }\StringTok{"blue"}\NormalTok{, }\StringTok{"white"}\NormalTok{, }\StringTok{"red"}\NormalTok{),}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(d) Simulated correlation"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 
-%% - Manuscript ---------------------
+\CommentTok{#plot simulated expression data}
+\KeywordTok{heatmap.2}\NormalTok{(expr, }\DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{bluered}\NormalTok{(}\DecValTok{50}\NormalTok{),}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(e) Simulated expression data (log scale)"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
-%% - In principle "as usual" again.
-%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
-%%   avoid empty lines before and after the equation (which would signal a new
-%%   paragraph.
-%% - When describing longer chunks of code that are _not_ meant for execution
-%%   (e.g., a function synopsis or list of arguments), the environment {Code}
-%%   is recommended. Alternatively, a plain {verbatim} can also be used.
-%%   (For executed code see the next section.)
+\begin{figure}
 
-\section{Methodology and software} \label{sec:methods}
+{\centering \includegraphics[width=.415\linewidth,height=.415\linewidth]{Plotsimulation_inhibiting-1} \includegraphics[width=.415\linewidth,height=.415\linewidth]{Plotsimulation_inhibiting-2} \includegraphics[width=.415\linewidth,height=.415\linewidth]{Plotsimulation_inhibiting-3} \includegraphics[width=.415\linewidth,height=.415\linewidth]{Plotsimulation_inhibiting-4} \includegraphics[width=.830\linewidth,height=.415\linewidth]{Plotsimulation_inhibiting-5}
 
-\begin{figure*}[!htb]
-%\begin{mdframed}
-%  \resizebox{\textwidth}{!}{
-         \begin{center}
-%
-        \subcaptionbox{Activating pathway structure  \label{fig:simple_graph:first}}{%
-            \fbox{
-            \includegraphics[width=0.45\textwidth]{{simple_graph.png}}
-            }
-        }%
-        \subcaptionbox{Pathway structure with inhibitions \label{fig:simple_graph:second}}{%
-            \fbox{
-           \includegraphics[width=0.45\textwidth]{{simple_graph_inhibiting.png}}
-           }
-        }%
-%
-    \end{center}
-   \caption[Simulated graph structures]{\small \textbf{\textbf{Simulated graph structures.}} A constructed graph structure used as an example to demonstrate the simulation procedure in Figures~\ref{fig:simulation_activating} and~\ref{fig:simulation_inhibiting}.
-Activating links are denoted by black arrows and inhibiting links by red edges.
-Inhibiting edges have been highlighted in red.}
-%}
-\label{fig:simple_graph}
-%\end{mdframed}
-\end{figure*}
-
-Here we present a procedure to simulate gene expression data with correlation structure derived from a known graph structure.
-This procedure assumes that transcriptomic data have been generated and follow a log-normal distribution (i.e., $log(X_{ij}) \sim MVN({\bf\mu}, \Sigma)$, where ${\bf\mu}$ and $\Sigma$ are the mean vector  and variance-covariance matrix respectively, for gene expression data derived from a biological pathway) after appropriate normalisation \citep{Law2014, Li2015}.
-Log-normality of gene expression matches the assumptions of the popular \pkg{limma} package, which is often used for the analysis of intensity-based data from gene expression microarray studies and count-based data from RNA-Seq experiments.
-This approach has also been applied for modelling UMI-based count data from single-cell RNA-Seq experiments in the \pkg{DESCEND} package \citep{Wang2018}.
-%rather than correlated blocks for $\Sigma$ in \pkg{mvtnorm} \citet{mvtnorm}
-%This procedure produces a ``Sigma'' variance-covariance matrix ($\Sigma$) for sampling from a multivariate normal distribution (using the \pkg{mvtnorm} \proglang{R} package).
-
-In order to simulate transcriptomic data, a pathway is first constructed as a graph structure, using the \pkg{igraph} \proglang{R} package \citep{igraph}, with the status of the edge relationships defined (i.e, whether they activate or inhibit downstream pathway members).
-%
-\textcolor{black}{This procedure uses} a graph structure such as that presented in Figure~\ref{fig:simple_graph:first}.
-The graph can be defined by an adjacency matrix, \textbf{$A$} (with elements \textcolor{black}{$A_{ij}$}), where  \textcolor{black}{
-%
-\[
-A_{ij} = 
-\begin{dcases*}
-   1                         & if genes $i$ and $j$ are adjacent \\
-   0                         & otherwise
-\end{dcases*}
-\]}
-%
-\noindent
-A matrix, \textbf{$R$}, with elements \textcolor{black}{$R_{ij}$}, is calculated based on distance (i.e., number of edges contained in the shortest path) between nodes, such that closer nodes are given more weight than more distant nodes, to define inter-node relationships.
-A geometrically-decreasing (relative) distance weighting is used to achieve this: \textcolor{black}{
-%
-\[
-R_{ij} = 
-\begin{dcases*}
-   1                            & if genes $i$ and $j$ are adjacent \\
-   (\sfrac{1}{2})^{d_{ij}}      & if a path can be found  between genes $i$ and $j$ \\
-   0                            & if no path exists between genes $i$ and $j$ 
-\end{dcases*}
-\]}
-%
-\noindent
-where $d_{ij}$ is the length of the shortest path (i.e., minimum number of edges traversed) between genes (nodes) $i$ and $j$ in graph $G$.
-Each more distant node is thus related by $\sfrac{1}{2}$ compared to the next nearest, as shown in Figure~\ref{fig:simulation_activating:second}.
-An arithmetically-decreasing (absolute) distance weighting is also supported in the \pkg{graphsim} \proglang{R} package which implements this procedure: \textcolor{black}{
-%
-\[
-R_{ij} = 
-\begin{dcases*}
-   1                            & if genes $i$ and $j$ are adjacent \\
-   1-\frac{d_{ij}}{diam(G)}     & if a path can be found  between genes $i$ and $j$ \\
-   0                            & if no path exists between genes $i$ and $j$ 
-\end{dcases*}
-\]
-%
 }
 
-Assuming a unit variance for each gene, these values can be used to derive a $\Sigma$ matrix:
-%
-\[
-\Sigma_{ij} = 
-\begin{dcases*}
-   1                            & if $i=j$ \\
-   \rho R_{ij}  & otherwise
-\end{dcases*}
-\]
-%
-\noindent
-where $\rho$ is the correlation between adjacent nodes. 
-%The $\Sigma$ matrix (with a diagonal of $1$) would thus have a standard deviation of 1 for each node. 
-Thus covariances between adjacent nodes are assigned by a correlation parameter ($\rho$) and the remaining off-diagonal values in the matrix are based on scaling these correlations by the geometrically weighted relationship matrix (or the nearest positive definite matrix for $\Sigma$ with negative correlations).\\
-
-Computing the nearest positive definite matrix is necessary to ensure that the variance-covariance matrix could be inverted when used as a parameter in multivariate normal simulations, particularly when negative correlations are included for inhibitions (as shown below).
-Matrices that could not be inverted occurred rarely with biologically plausible graph structures but this approach allows for the computation of a plausible correlation matrix when the graph structure given is incomplete or contains loops.
-When required, the nearest positive definite matrix is computed using the \texttt{nearPD} function of the \pkg{Matrix} \proglang{R} package \citep{Matrix} to perform Higham's algorithm \citep{Higham2002} on variance-covariance matrices.
-The \pkg{graphsim} package gives a warning when this occurs.
-
-\section{Illustrations} \label{sec:illustrations}
-\subsection{Generating a Graph Structure} \label{sec:plot_graph}
-
-The graph structure in Figure~\ref{fig:simple_graph:first} was used to simulate correlated gene expression data by sampling from a multivariate normal distribution using the \pkg{mvtnorm} \proglang{R} package \citep{Genz2009, mvtnorm}.
-%This simulation procedure will be demonstrated with the relatively simple constructed graph structure shown in Figure~\ref{fig:simple_graph}. 
-The graph structure visualisation in Figure~\ref{fig:simple_graph} was specifically developed for (directed) iGraph objects in \proglang{R} and is available in the \pkg{plot.igraph} and \pkg{igraph.extensions} packages.
-The \texttt{plot\_directed} function enables customisation of plot parameters for each node or edge, and mixed (directed) edge types for indicating activation or inhibition.
-These inhibition links (which occur frequently in biological pathways) are demonstrated in Figure~\ref{fig:simple_graph:second}.
-
-A graph structure can be generated and plotted using the following commands in R:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-#install packages required (once per machine)
-install.packages("igraph")
-install.packages("devtools")
-library("devtools")
-devtools::install\_github("TomKellyGenetics/graphsim")
-#install custom plotting package
-devtools::install\_github("TomKellyGenetics/plot.igraph")
-\end{CodeInput}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-#load required packages (once per R instance)
-library("igraph")
-library("graphsim")
-library("plot.igraph")
-\end{CodeInput}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-#generate graph structure
-graph\_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"), c("D", "E"), 
-    c("D", "F"), c("F", "G"), c("F", "I"), c("H", "I"))
-graph <- graph.edgelist(graph\_edges, directed = T)
-
-#plot graph structure (Figure 1)
-plot\_directed(graph, state = "activating", layout = layout.kamada.kawai, 
-    cex.node=3, cex.arrow=5, arrow\_clip = 0.2)
-\end{CodeInput}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-#generate parameters for inhibitions
-state <-  c(1, 1, -1, 1, 1, 1, 1, -1, 1)
-
-#plot graph structure with inhibitions (Figure 2)
-plot\_directed(graph, state=state,  layout = layout.kamada.kawai, 
-    cex.node=3, cex.arrow=5, arrow\_clip = 0.2)
-\end{CodeInput}
-\end{CodeChunk}
-
-\subsection{Generating a Simulated Expression Dataset} \label{sec:graphsim_demo}
-
-\begin{figure*}[!hp]
-%\begin{mdframed}
-%  \resizebox{\textwidth}{!}{
-         \begin{center}
-%
-        \subcaptionbox{Activating pathway structure \label{fig:simulation_activating:first}}{%
-            \includegraphics[width=0.35\textwidth]{{simple_graph.png}}
-        }%
-        \subcaptionbox{Relationship matrix \label{fig:simulation_activating:second}}{%
-            \includegraphics[width=0.35\textwidth]{{dist_mat.png}}
-        }%
-
-        \subcaptionbox{$\Sigma$ (covariance matrix) \label{fig:simulation_activating:third}}{%
-           \includegraphics[width=0.35\textwidth]{{sigma_mat.png}}
-        }%
-	\subcaptionbox{Simulated correlation\label{fig:simulation_activating:fifth}}{%
-           \includegraphics[width=0.35\textwidth]{{expr_cor_mat.png}}
-        }%
-
-	\subcaptionbox{Simulated expression data (log scale) \label{fig:simulation_activating:fourth}}{%
-            \includegraphics[width=1\textwidth]{{expr_mat_wide.png}}
-        }%
-        %\subcaptionbox{Simulated gene function calls \label{fig:simulation_activating:sixth}}{%
-        %   \includegraphics[width=0.35\textwidth]{{expr_disc_mat.png}}
-        %}%
-    \end{center}
-   \caption[Simulating expression from a graph structure]{\small \textbf{\textbf{Simulating expression from a graph structure.}} An example of a graph structure (a) that has been used to derive a relationship matrix (b), $\Sigma$ matrix (c) and correlation structure (d) from the relative distances between the nodes.
-Non-negative values are coloured white to red from $0$ to $1$.
-This $\Sigma$ matrix has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
-Here genes with closer relationships in the pathway structure show higher correlation between simulated values.}
-%}
-\label{fig:simulation_activating}
-%\end{mdframed}
-\end{figure*}
-
-The correlation parameter of $\rho = 0.8$ is used to demonstrate the inter-correlated datasets using a geometrically-generated relationship matrix (as used for the example in Figure~\ref{fig:simulation_activating:third}).
-This $\Sigma$ matrix was then used to sample from a multivariate normal distribution such that each gene had a mean of $0$, standard deviation $1$, and covariance within the range $[0,1]$ so that the off-diagonal elements of $\Sigma$ represent correlations.
-This procedure generated a simulated (continuous normally-distributed) log-expression profile for each node (Figure~\ref{fig:simulation_activating:fourth}) with a corresponding correlation structure (Figure~\ref{fig:simulation_activating:fifth}).
-The simulated correlation structure closely resembled the expected correlation structure ($\Sigma$ in Figure~\ref{fig:simulation_activating:third}) even for the relatively modest sample size ($N=100$) illustrated in Figure~\ref{fig:simulation_activating}.
-Once a gene expression dataset comprising multiple pathways has been generated (as in Figure~\ref{fig:simulation_activating:fourth}),
-%then a discrete matrix of gene function was constructed with a functional threshold quantile to simulate functional relationships of synthetic lethality (Figure~\ref{fig:SL_Model_Expression}).
-%Throughout this thesis, this threshold is the 0.3 quantile (as discussed in Section~\ref{methods:SL_Model}) which generates functional discrete matrices such as those used for synthetic lethal simulation in Section~\ref{methods:simulating_SL} (as shown Figure~\ref{fig:simulation_activating:sixth}).
-it can then be used to test procedures designed for analysis of empirical gene expression data (such as those generated by microarrays or RNA-Seq) that have been normalised on a log-scale.
-
-The simulated dataset can be generated using the following \proglang{R} code:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-#adjacency matrix
-adj\_mat <- make\_adjmatrix\_graph(graph)
-
-#relationship matrix
-dist\_mat <- make\_distance\_graph(graph\_test4, absolute = F)
-
-#sigma matrix directly from graph
-sigma\_mat <- make\_sigma\_mat\_dist\_graph(graph, 0.8, absolute = F)
-
-#show shortest paths of graph
-shortest\_paths <- shortest.paths(graph)
-
-#generate expression data directly from graph
-expr <- generate\_expression(100, graph, cor = 0.8, mean = 0, comm = F,
-                            dist = T, absolute = F, state = state)
-\end{CodeInput}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-##plot steps
-
-#plot adjacency matrix
-heatmap.2(make\_adjmatrix\_graph(graph), scale = "none", trace = "none", 
-          col = colorpanel(3, "grey75", "white", "blue"),
-          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-
-#plot relationship matrix
-heatmap.2(make\_distance\_graph(graph\_test4, absolute = F), scale = "none",
-          trace = "none", col = bluered(50), colsep = 1:length(V(graph)), 
-        rowsep = 1:length(V(graph)))
-
-#plot sigma matrix
-heatmap.2(make\_sigma\_mat\_dist\_graph(graph, 0.8, absolute = F), scale = "none", 
-          trace = "none", col = bluered(50), colsep = 1:length(V(graph)), 
-          rowsep = 1:length(V(graph)))
-expr <- generate\_expression(100, graph, cor = 0.8, mean = 0, comm = F, dist = T,
-                            absolute = F, state = state)
-
-#plot simulated expression data
-heatmap.2(expr, scale = "none", trace = "none", col = bluered(50), 
-          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-
-#plot simulated correlations
-heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = bluered(50), 
-          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-\end{CodeInput}
-\end{CodeChunk}
-
-\begin{figure*}[!hp]
-%\begin{mdframed}
-%  \resizebox{\textwidth}{!}{
-         \begin{center}
-%
-        \subcaptionbox{Inhibiting pathway structure\label{fig:simulation_inhibiting:first}}{%
-            \includegraphics[width=0.35\textwidth]{{simple_graph_inhibiting.png}}
-        }%
-        \subcaptionbox{Relationship matrix \label{fig:simulation_inhibiting:second}}{%
-            \includegraphics[width=0.35\textwidth]{{dist_mat.png}}
-        }%
-
-        \subcaptionbox{$\Sigma$ (covariance matrix) \label{fig:simulation_inhibiting:third}}{%
-           \includegraphics[width=0.35\textwidth]{{sigma_mat_inhibiting.png}}
-        }%
-	\subcaptionbox{Simulated correlation\label{fig:simulation_inhibiting:fifth}}{%
-           \includegraphics[width=0.35\textwidth]{{expr_cor_mat_inhibiting.png}}
-        }%
-
-        \subcaptionbox{Simulated expression data (log scale) \label{fig:simulation_inhibiting:fourth}}{%
-
-            \includegraphics[width=1\textwidth]{{expr_mat_inhibiting_wide.png}}
-        }%
-        %\subcaptionbox{Simulated gene function calls \label{fig:simulation_inhibiting:sixth}}{%
-        %   \includegraphics[width=0.35\textwidth]{{expr_inhib_disc_mat.png}}
-        %}%
-    \end{center}
-    \caption[Simulating expression from graph structure with inhibitions]{\small \textbf{\textbf{Simulating expression from graph structure with inhibitions.}} An example of a graph structure (a), that has been used to derive a relationship matrix (b), $\Sigma$ matrix (c), and correlation structure (d), from the relative distances between the nodes.
+\caption{\textbf{Simulating expression from graph structure with inhibitions}.
+An example of a graph structure (a), that has been used to derive a relationship matrix (b), $\Sigma$ matrix (c), and correlation structure (d), from the relative distances between the nodes.
 These values are coloured blue to red from $-1$ to $1$.
 This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
-Here the inhibitory relationships between genes are reflected in negatively correlated simulated  values.}
-%}
-\label{fig:simulation_inhibiting}
-%\end{mdframed}
-\end{figure*}
+Here the inhibitory relationships between genes are reflected in negatively correlated simulated  values.}\label{fig:simulation_inhibiting}
+\end{figure}
 
-The simulation procedure (Figure~\ref{fig:simulation_activating}) can similarly be used for pathways containing inhibitory links (Figure~\ref{fig:simulation_inhibiting}) with several refinements.
-With the inhibitory links (Figure~\ref{fig:simulation_inhibiting:first}), distances are calculated in the same manner as before (Figure~\ref{fig:simulation_inhibiting:second}) with inhibitions accounted for by iteratively multiplying downstream nodes by $-1$ to form modules with negative correlations between them (Figures~\ref{fig:simulation_inhibiting:third} and~\ref{fig:simulation_inhibiting:fifth}).
-A multivariate normal distribution with these negative correlations can be sampled to generate simulated data (Figure~\ref{fig:simulation_inhibiting:fourth}).% and~\ref{fig:simulation_inhibiting:sixth}).
+The simulation procedure is also demonstrated here
+(Figure~\protect\hyperlink{fig:simulation_smad}{4}) on a pathway
+structure for a known biological pathway (from reactome R-HSA-2173789)
+of TGF-\(\beta\) receptor signaling activates SMADs
+(Figure~\protect\hyperlink{fig:simulation_smad:first}{4a}) derived from
+the Reactome database version 52 \texttt{{[}@Reactome{]}}. Distances are
+calculated in the same manner as before
+(Figure~\protect\hyperlink{fig:simulation_smad:second}{4b}) producing
+blocks of correlated genes
+(Figures~\protect\hyperlink{fig:simulation_smad:third}{4c}
+and~\protect\hyperlink{fig:simulation_smad:fifth}{4d}). This shows that
+multivariate normal distribution can be sampled to generate simulated
+data to represent expression with the complexity of a biological pathway
+(Figure~\protect\hyperlink{fig:simulation_smad:fourth}{4e}). Here
+\emph{SMAD7} exhibits negative correlations with the other SMADs
+consistent with it's functions as as an ``inhibitor SMAD'' with
+competitively inhibits \emph{SMAD4}.
 
-\begin{figure*}[!hp]
-%\begin{mdframed}
-%  \resizebox{\textwidth}{!}{
-         \begin{center}
-%
-        \subcaptionbox{TGF-$\beta$ activates SMADs\label{fig:simulation_smad:first}}{%
-            \includegraphics[width=0.35\textwidth]{{graph_smad_simple.pdf}}
-        }%
-        \subcaptionbox{Relationship matrix \label{fig:simulation_smad:second}}{%
-            \includegraphics[width=0.35\textwidth]{{dist_mat_smad_simple.png}}
-        }%
+\begin{Shaded}
+\begin{Highlighting}[]
 
-        \subcaptionbox{$\Sigma$ (covariance matrix) \label{fig:simulation_smad:third}}{%
-           \includegraphics[width=0.35\textwidth]{{sigma_mat_inhibiting_smad_simple.png}}
-        }%
-	\subcaptionbox{Simulated correlation\label{fig:simulation_smad:fifth}}{%
-           \includegraphics[width=0.35\textwidth]{{expr_cor_mat_inhibiting_smad_simple.png}}
-        }%
+\CommentTok{#import graph from data}
+\NormalTok{graph <-}\StringTok{ }\KeywordTok{identity}\NormalTok{(TGFBeta_Smad_graph)}
 
-        \subcaptionbox{Simulated expression data (log scale) \label{fig:simulation_smad:fourth}}{%
+\CommentTok{#generate parameters for inhibitions}
+\NormalTok{state <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DecValTok{1}\NormalTok{, }\KeywordTok{length}\NormalTok{(}\KeywordTok{E}\NormalTok{(graph)))}
+\NormalTok{pathway <-}\StringTok{ }\KeywordTok{get.edgelist}\NormalTok{(graph)}
+\NormalTok{state[pathway[,}\DecValTok{1}\NormalTok{] }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"SMAD6"}\NormalTok{, }\StringTok{"SMAD7"}\NormalTok{, }\StringTok{"BAMBI"}\NormalTok{, }\StringTok{"SMURF1"}\NormalTok{, }\StringTok{"SMURF2"}\NormalTok{, }\StringTok{"UCHL5"}\NormalTok{, }\StringTok{"USP15"}\NormalTok{, }\StringTok{"UBB"}\NormalTok{, }\StringTok{"UBC"}\NormalTok{, }\StringTok{"PMEPA1"}\NormalTok{, }\StringTok{"PPP1CA"}\NormalTok{, }\StringTok{"PPP1CB"}\NormalTok{, }\StringTok{"PPP1CC"}\NormalTok{, }\StringTok{"PPP1R15A"}\NormalTok{)] <-}\StringTok{ }\DecValTok{2}
+\NormalTok{state[}\KeywordTok{is.na}\NormalTok{(state)] <-}\StringTok{ }\DecValTok{1}
 
-            \includegraphics[width=0.8\textwidth]{{expr_mat_inhibiting_wide_smad_simple.png}}
-        }%
-        %\subcaptionbox{Simulated gene function calls \label{fig:simulation_smad:sixth}}{%
-        %   \includegraphics[width=0.35\textwidth]{{expr_inhib_disc_mat.png}}
-        %}%
-    \end{center}
-   \caption[Simulating expression from a biological pathway graph structure]{\small \textbf{\textbf{Simulating expression from graph structure with inhibitions.}} The graph structure (a) of a known biological pathway, the TGF-$\beta$ receptor signaling activates SMADs (R-HSA-2173789), was used to derive a relationship matrix (b), $\Sigma$ matrix (c) and correlation structure (d) from the relative distances between the nodes.
+\KeywordTok{plot_directed}\NormalTok{(graph, }\DataTypeTok{state =}\NormalTok{ state, }\DataTypeTok{layout =}\NormalTok{ layout.kamada.kawai,}
+              \DataTypeTok{border.node=}\NormalTok{scales}\OperatorTok{::}\KeywordTok{alpha}\NormalTok{(}\StringTok{"black"}\NormalTok{, }\FloatTok{0.75}\NormalTok{), }\DataTypeTok{fill.node=}\StringTok{"lightblue"}\NormalTok{,}
+              \DataTypeTok{col.arrow =} \KeywordTok{c}\NormalTok{(scales}\OperatorTok{::}\KeywordTok{alpha}\NormalTok{(}\StringTok{"navyblue"}\NormalTok{, }\FloatTok{0.25}\NormalTok{), scales}\OperatorTok{::}\KeywordTok{alpha}\NormalTok{(}\StringTok{"red"}\NormalTok{, }\FloatTok{0.25}\NormalTok{))[state], }
+              \DataTypeTok{cex.node =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.label =} \FloatTok{0.8}\NormalTok{, }\DataTypeTok{cex.arrow =} \DecValTok{2}\NormalTok{, }
+              \DataTypeTok{sub =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"(a) TFG-"}\NormalTok{, Beta, }\StringTok{" activates SMADs"}\NormalTok{)), }\DataTypeTok{cex.sub =} \FloatTok{1.75}\NormalTok{)}
+\KeywordTok{box}\NormalTok{()}
+
+\CommentTok{#adjacency matrix}
+\NormalTok{adj_mat <-}\StringTok{ }\KeywordTok{make_adjmatrix_graph}\NormalTok{(graph)}
+
+\CommentTok{#relationship matrix}
+\NormalTok{dist_mat <-}\StringTok{ }\KeywordTok{make_distance_graph}\NormalTok{(graph, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{)}
+
+\CommentTok{#sigma matrix directly from graph}
+\NormalTok{sigma_mat <-}\StringTok{ }\KeywordTok{make_sigma_mat_dist_graph}\NormalTok{(graph, }\DataTypeTok{state =}\NormalTok{ state, }\FloatTok{0.8}\NormalTok{, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{)}
+\CommentTok{#> Warning in eattrs[[name]][index] <- value: number of items to}
+\CommentTok{#> replace is not a multiple of replacement length}
+
+\CommentTok{#show shortest paths of graph}
+\NormalTok{shortest_paths <-}\StringTok{ }\KeywordTok{shortest.paths}\NormalTok{(graph)}
+
+\CommentTok{#generate expression data directly from graph}
+\NormalTok{expr <-}\StringTok{ }\KeywordTok{generate_expression}\NormalTok{(}\DecValTok{100}\NormalTok{, graph, }\DataTypeTok{state =}\NormalTok{ state, }\DataTypeTok{cor =} \FloatTok{0.8}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{comm =} \OtherTok{FALSE}\NormalTok{,}
+                            \DataTypeTok{dist =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{)}
+\CommentTok{#> Warning in eattrs[[name]][index] <- value: number of items to}
+\CommentTok{#> replace is not a multiple of replacement length}
+\CommentTok{#> Warning in state_path[jj] <- state[kk]: number of items to replace}
+\CommentTok{#> is not a multiple of replacement length}
+
+\CommentTok{#> Warning in state_path[jj] <- state[kk]: number of items to replace}
+\CommentTok{#> is not a multiple of replacement length}
+
+\CommentTok{#> Warning in state_path[jj] <- state[kk]: number of items to replace}
+\CommentTok{#> is not a multiple of replacement length}
+\CommentTok{#> Warning in generate_expression(100, graph, state = state, cor =}
+\CommentTok{#> 0.8, mean = 0, : sigma matrix was not positive definite, nearest}
+\CommentTok{#> approximation used.}
+
+\CommentTok{# #plot relationship matrix}
+\KeywordTok{heatmap.2}\NormalTok{(}\KeywordTok{make_distance_graph}\NormalTok{(graph, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{),}
+          \DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{colorpanel}\NormalTok{(}\DecValTok{50}\NormalTok{, }\StringTok{"white"}\NormalTok{, }\StringTok{"red"}\NormalTok{),}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)),}
+              \DataTypeTok{sub =} \StringTok{"(b) relationship matrix"}\NormalTok{, }\DataTypeTok{cex.sub =} \FloatTok{1.75}\NormalTok{)}
+
+\CommentTok{# #plot sigma matrix}
+\KeywordTok{heatmap.2}\NormalTok{(}\KeywordTok{make_sigma_mat_dist_graph}\NormalTok{(graph, }\DataTypeTok{state =}\NormalTok{ state, }\FloatTok{0.8}\NormalTok{, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{),}
+\DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{colorpanel}\NormalTok{(}\DecValTok{50}\NormalTok{, }\StringTok{"blue"}\NormalTok{, }\StringTok{"white"}\NormalTok{, }\StringTok{"red"}\NormalTok{),}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
+\CommentTok{#> Warning in eattrs[[name]][index] <- value: number of items to}
+\CommentTok{#> replace is not a multiple of replacement length}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"(c) "}\NormalTok{, Sigma, }\StringTok{" matrix"}\NormalTok{)), }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+
+\CommentTok{#simulated data}
+\NormalTok{expr <-}\StringTok{ }\KeywordTok{generate_expression}\NormalTok{(}\DecValTok{100}\NormalTok{, graph,  }\DataTypeTok{state =}\NormalTok{ state, }\DataTypeTok{cor =} \FloatTok{0.8}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{,}
+\DataTypeTok{comm =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{dist =}\OtherTok{TRUE}\NormalTok{, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{)}
+\CommentTok{#> Warning in eattrs[[name]][index] <- value: number of items to}
+\CommentTok{#> replace is not a multiple of replacement length}
+\CommentTok{#> Warning in state_path[jj] <- state[kk]: number of items to replace}
+\CommentTok{#> is not a multiple of replacement length}
+
+\CommentTok{#> Warning in state_path[jj] <- state[kk]: number of items to replace}
+\CommentTok{#> is not a multiple of replacement length}
+
+\CommentTok{#> Warning in state_path[jj] <- state[kk]: number of items to replace}
+\CommentTok{#> is not a multiple of replacement length}
+\CommentTok{#> Warning in generate_expression(100, graph, state = state, cor =}
+\CommentTok{#> 0.8, mean = 0, : sigma matrix was not positive definite, nearest}
+\CommentTok{#> approximation used.}
+
+\CommentTok{#plot simulated correlations}
+\KeywordTok{heatmap.2}\NormalTok{(}\KeywordTok{cor}\NormalTok{(}\KeywordTok{t}\NormalTok{(expr)), }\DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{colorpanel}\NormalTok{(}\DecValTok{50}\NormalTok{, }\StringTok{"blue"}\NormalTok{, }\StringTok{"white"}\NormalTok{, }\StringTok{"red"}\NormalTok{),}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(d) Simulated correlation"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+
+\CommentTok{#plot simulated expression data}
+\KeywordTok{heatmap.2}\NormalTok{(expr, }\DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{bluered}\NormalTok{(}\DecValTok{50}\NormalTok{),}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(e) Simulated expression data (log scale)"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+
+{\centering \includegraphics[width=.415\linewidth,height=.415\linewidth]{Plotsimulation_smad-1} \includegraphics[width=.415\linewidth,height=.415\linewidth]{Plotsimulation_smad-2} \includegraphics[width=.415\linewidth,height=.415\linewidth]{Plotsimulation_smad-3} \includegraphics[width=.415\linewidth,height=.415\linewidth]{Plotsimulation_smad-4} \includegraphics[width=.830\linewidth,height=.415\linewidth]{Plotsimulation_smad-5}
+
+}
+
+\caption{\textbf{Simulating expression from a biological pathway graph structure}.
+The graph structure (a) of a known biological pathway, the TGF-$\beta$ receptor signaling activates SMADs (R-HSA-2173789), was used to derive a relationship matrix (b), $\Sigma$ matrix (c) and correlation structure (d) from the relative distances between the nodes.
 These values are coloured blue to red from $-1$ to $1$.
 This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
-Here modules of genes with correlated expression can be clearly discerned.}
-%}
-\label{fig:simulation_smad}
-%\end{mdframed}
-\end{figure*}
-
-The simulation procedure is also demonstrated here (Figure~\ref{fig:simulation_smad}) on a pathway structure for a known biological pathway (from reactome R-HSA-2173789) of TGF-$\beta$ receptor signaling activates SMADs (Figure~\ref{fig:simulation_smad:first}) derived from the Reactome database version 52 \citep{Reactome}.
-Distances are calculated in the same manner as before (Figure~\ref{fig:simulation_smad:second}) producing blocks of correlated genes (Figures~\ref{fig:simulation_inhibiting:third} and~\ref{fig:simulation_inhibiting:fifth}).
-This shows that multivariate normal distribution can be sampled to generate simulated data to represent expression with the complexity of a biological pathway (Figure~\ref{fig:simulation_inhibiting:fourth}).
-Here \textit{SMAD7} exhibits negative correlations with the other SMADs consistent with it's functions as as an ``inhibitor SMAD'' with competitively inhibits \textit{SMAD4}.
-% and~\ref{fig:simulation_inhibiting:sixth}).
-
-\iffalse
-These simulated datasets could then be used for simulating synthetic lethal partners of a query gene within a graph network.
-The query gene was assumed to be separate from the graph network pathway and was added to the dataset using the procedure in Section~\ref{methods:simulating_SL}.
-Thus I can simulate known synthetic lethal partner genes within a synthetic lethal partner pathway structure.
-\fi
-
-\iffalse
-
-	\includegraphics{{dist_mat.png}}
-	\includegraphics{{sigma_mat.png}}
-		\includegraphics{{expr_mat.png}}
-			\includegraphics{{expr_cor_mat.png}}
-		\includegraphics{{expr_disc_mat.png}}
-
-	\includegraphics{{state_matrix_inhibiting.png}}
-	\includegraphics{{dist_mat.png}}
-		\includegraphics{{sigma_mat_inhibiting.png}}
-		\includegraphics{{expr_inhib_mat.png}}
-			\includegraphics{{expr_inhib_cor_mat.png}}
-		\includegraphics{{expr_inhib_disc_mat.png}}
-
-\fi
-
-%\FloatBarrier
-
-%% - Illustrations --------------------
-
-%% - Virtually all JSS manuscripts list source code along with the generated
-%%   output. The style files provide dedicated environments for this.
-%% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
-%%   or knitr using the render_sweave() hook - are used (without the need to
-%%   load Sweave.sty).
-%% - Equivalently, {CodeInput} and {CodeOutput} can be used.
-%% - The code input should use "the usual" command prompt in the respective
-%%   software system.
-%% - For \proglang{R} code, the prompt "R> " should be used with "+  " as the
-%%   continuation prompt.
-%% - Comments within the code chunks should be avoided - these should be made
-%%   within the regular LaTeX text.
-
-\iffalse
-
-\section{Illustrations} \label{sec:illustrations}
-
-For a simple illustration of basic Poisson and NB count regression the
-\code{quine} data from the \pkg{MASS} package is used. This provides the number
-of \code{Days} that children were absent from school in Australia in a
-particular year, along with several covariates that can be employed as regressors.
-The data can be loaded by
-%
-\begin{CodeChunk}
-\begin{CodeInput}
-R> data("quine", package = "MASS")
-\end{CodeInput}
-\end{CodeChunk}
-%
-and a basic frequency distribution of the response variable is displayed in
-Figure~\ref{fig:quine}.
-
-\begin{leftbar}
-For code input and output, the style files provide dedicated environments.
-Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
-or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
-produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
-hook. Please make sure that all code is properly spaced, e.g., using
-\code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
-use ``the usual'' command prompt in the respective software system. For
-\proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
-the continuation prompt. Generally, comments within the code chunks should be
-avoided - and made in the regular {\LaTeX} text instead. Finally, empty lines
-before and after code input/output should be avoided (see above).
-\end{leftbar}
-
-\begin{figure}[t!]
-\centering
-\includegraphics{article-visualization}
-\caption{\label{fig:quine} Frequency distribution for number of days absent
-from school.}
+Here modules of genes with correlated expression can be clearly discerned.}\label{fig:simulation_smad}
 \end{figure}
 
-As a first model for the \code{quine} data, we fit the basic Poisson regression
-model. (Note that JSS prefers when the second line of code is indented by two
-spaces.)
-%
-\begin{CodeChunk}
-\begin{CodeInput}
-R> m\_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
-+    family = poisson)
-\end{CodeInput}
-\end{CodeChunk}
-%
-To account for potential overdispersion we also consider a negative binomial
-GLM.
-%
-\begin{CodeChunk}
-\begin{CodeInput}
-R> library("MASS")
-R> m\_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
-\end{CodeInput}
-\end{CodeChunk}
-%
-In a comparison with the BIC the latter model is clearly preferred.
-%
-\begin{CodeChunk}
-\begin{CodeInput}
-R> BIC(m\_pois, m\_nbin)
-\end{CodeInput}
-\begin{CodeOutput}
-       df      BIC
-m\_pois 18 2046.851
-m\_nbin 19 1157.235
-\end{CodeOutput}
-\end{CodeChunk}
-%
-Hence, the full summary of that model is shown below.
-%
-\begin{CodeChunk}
-\begin{CodeInput}
-R> summary(m\_nbin)
-\end{CodeInput}
-\begin{CodeOutput}
-Call:
-glm.nb(formula = Days ~ (Eth + Sex + Age + Lrn)^2, data = quine, 
-    init.theta = 1.60364105, link = log)
+These simulated datasets can also be used for simulating gene expression
+data within a graph network to test genomic analysis techniques.
+Correlation structure can be included into datasets generated when
+testing whether true positive genes or samples can be detected in a
+sample with the background of complex pathway structure.
 
-Deviance Residuals: 
-    Min       1Q   Median       3Q      Max  
--3.0857  -0.8306  -0.2620   0.4282   2.0898
+\hypertarget{sec:summary}{%
+\section{Summary and discussion}\label{sec:summary}}
 
-Coefficients: (1 not defined because of singularities)
-            Estimate Std. Error z value Pr(>|z|)    
-(Intercept)  3.00155    0.33709   8.904  < 2e-16 ***
-EthN        -0.24591    0.39135  -0.628  0.52977    
-SexM        -0.77181    0.38021  -2.030  0.04236 *  
-AgeF1       -0.02546    0.41615  -0.061  0.95121    
-AgeF2       -0.54884    0.54393  -1.009  0.31296    
-AgeF3       -0.25735    0.40558  -0.635  0.52574    
-LrnSL        0.38919    0.48421   0.804  0.42153    
-EthN:SexM    0.36240    0.29430   1.231  0.21818    
-EthN:AgeF1  -0.70000    0.43646  -1.604  0.10876    
-EthN:AgeF2  -1.23283    0.42962  -2.870  0.00411 ** 
-EthN:AgeF3   0.04721    0.44883   0.105  0.91622    
-EthN:LrnSL   0.06847    0.34040   0.201  0.84059    
-SexM:AgeF1   0.02257    0.47360   0.048  0.96198    
-SexM:AgeF2   1.55330    0.51325   3.026  0.00247 ** 
-SexM:AgeF3   1.25227    0.45539   2.750  0.00596 ** 
-SexM:LrnSL   0.07187    0.40805   0.176  0.86019    
-AgeF1:LrnSL -0.43101    0.47948  -0.899  0.36870    
-AgeF2:LrnSL  0.52074    0.48567   1.072  0.28363    
-AgeF3:LrnSL       NA         NA      NA       NA    
--
-Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+Biological pathways are of fundamental importance to understanding
+molecular biology. In order to translate findings from genomics studies
+into real-world applications such as improved healthcare, the roles of
+genes must be studied in the context of molecular pathways. Here we
+present a statistical framework to simulate gene expression from
+biological pathways, and provide the \texttt{graphsim} package in R to
+generate these simulated datasets. This approach is versatile and can be
+fine-tuned for modelling existing biological pathways or for testing
+whether constructed pathways can be detected by other means. In
+particular, methods to infer biological pathways and gene regulatory
+networks from gene expression data can be tested on simulated datasets
+using this framework. The package also enables simulation of complex
+gene expression datasets to test how these pathways impact on
+statistical analysis of gene expression data using existing methods or
+novel statistical methods being developed for gene expression data
+analysis.
 
-(Dispersion parameter for Negative Binomial(1.6036) family taken to be 1)
+\hypertarget{computational-details}{%
+\section*{Computational details}\label{computational-details}}
+\addcontentsline{toc}{section}{Computational details}
 
-    Null deviance: 235.23  on 145  degrees of freedom
-Residual deviance: 167.53  on 128  degrees of freedom
-AIC: 1100.5
+The results in this paper were obtained using R 3.6.1 with the
+\texttt{igraph} 1.2.4.1 \texttt{Matrix} 1.2-17, \texttt{matrixcalc}
+1.0-3, and \texttt{mvtnorm} 1.0-11 packages. R itself and all dependent
+packages used are available from the Comprehensive Archive Network
+(CRAN) at \url{https://CRAN.R-project.org}. The \texttt{graphsim}
+package presented can be installed from CRAN and the issues can be
+reported to the development version on GitHub
+(\url{https://github.com/TomKellyGenetics/graphsim}). This package is
+included in the library on GitHub
+(\url{https://github.com/TomKellyGenetics/igraph.extensions}) which
+installs various tools for \texttt{igraph} analysis. This software is
+cross-platform and compatible with installations on Windows, Mac, and
+Linux operating systems. The package GitHub repository also contains
+vignettes with more information and examples on running functions
+released in the package. The package (\texttt{graphsim} 0.1.2) has been
+released on CRAN and will be updated.
 
-Number of Fisher Scoring iterations: 1
+\hypertarget{acknowledgements}{%
+\section*{Acknowledgements}\label{acknowledgements}}
+\addcontentsline{toc}{section}{Acknowledgements}
 
-              Theta:  1.604 
-          Std. Err.:  0.214
+This package was developed as part of a PhD research project funded by
+the Postgraduate Tassell Scholarship in Cancer Research Scholarship
+awarded to STK. We thank members of the Laboratory of Professor Satoru
+Miyano at the University of Tokyo, Institute for Medical Science,
+Professor Seiya Imoto, Associate Professor Rui Yamaguchi, and Dr Paul
+Sheridan (Assistant Professor at Hirosaki University,CSO at Tupac Bio)
+for helpful discussions in this field. We also thank Professor Parry
+Guilford at the University of Otago, Professor Cristin Print at the
+University of Auckland, and Dr Erik Arner at the RIKEN Center for
+Integrative Medical Sciences for their excellent advice during this
+project.
 
- 2 x log-likelihood:  -1062.546 
-\end{CodeOutput}
-\end{CodeChunk}
-\fi
+\hypertarget{contributions}{%
+\section*{Author Contributions}\label{contributions}}
+\addcontentsline{toc}{section}{Author Contributions}
 
-%% - Summary/conclusions/discussion ---------------
+S.T.K. and M.A.B. conceived of the presented methodology. S.T.K.
+developed the theory and performed the computations. M.A.B. provided
+guidance throughout the project and gave feedback on the package. All
+authors discussed the package and contributed to the final manuscript.
 
-\FloatBarrier
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
 
-\section{Summary and discussion} \label{sec:summary}
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Arner2015}{}%
+Arner, E., C. O. Daub, K. Vitting-Seerup, R. Andersson, B. Lilje, F.
+Drabløs, A. Lennartsson, et al. 2015. ``Transcribed enhancers lead waves
+of coordinated transcription in transitioning mammalian cells.''
+\emph{Science} 347 (6225): 1010--4.
 
-\begin{leftbar}
-Biological pathways are of fundamental importance to understanding molecular biology.
-In order to translate findings from genomics studies into real-world applications such as improved healthcare, the roles of genes must be studied in the context of molecular pathways.
-Here we present a statistical framework to simulate gene expression from biological pathways, and provide the \pkg{graphsim} package in \proglang{R} to generate these simulated datasets.
-This approach is versatile and can be fine-tuned for modelling existing biological pathways or for testing whether constructed pathways can be detected by other means.
-In particular, methods to infer biological pathways and gene regulatory networks from gene expression data can be tested on simulated datasets using this framework.
-The package also enables simulation of complex gene expression datasets to test how these pathways impact on statistical analysis of gene expression data using existing methods or novel statistical methods being developed for gene expression data analysis.
-\end{leftbar}
+\leavevmode\hypertarget{ref-Barabasi2004}{}%
+Barabási, A. L., and Z. N. Oltvai. 2004. ``Network Biology:
+Understanding the Cell's Functional Organization.'' \emph{Nat Rev Genet}
+5 (2): 101--13.
 
-%% - Optional special unnumbered sections -------------
+\leavevmode\hypertarget{ref-Matrix}{}%
+Bates, Douglas, and Martin Maechler. 2016. \emph{Matrix: Sparse and
+Dense Matrix Classes and Methods}.
+\url{https://CRAN.R-project.org/package=Matrix}.
 
-\section*{Computational details}
+\leavevmode\hypertarget{ref-Reactome}{}%
+Croft, D, A F Mundo, R Haw, M Milacic, J Weiser, G Wu, M Caudy, et al.
+2014. ``The Reactome pathway knowledgebase.'' Journal Article.
+\emph{Nucleic Acids Res} 42 (database issue): D472--D477.
+\url{https://doi.org/10.1093/nar/gkt1102}.
 
-\iffalse
-\begin{leftbar}
-If necessary or useful, information about certain computational details
-such as version numbers, operating systems, or compilers could be included
-in an unnumbered section. Also, auxiliary packages (say, for visualizations,
-maps, tables, \dots) that are not cited in the main text can be credited here.
-\end{leftbar}
-\fi
+\leavevmode\hypertarget{ref-igraph}{}%
+Csardi, Gabor, and Tamas Nepusz. 2006. ``The Igraph Software Package for
+Complex Network Research.'' \emph{InterJournal} Complex Systems: 1695.
+\url{http://igraph.org}.
 
-The results in this paper were obtained using
-\proglang{R}~3.6.1 with the
-\pkg{igraph}~1.2.4.1 \pkg{Matrix}~1.2-17, \pkg{matrixcalc}~1.0-3, and \pkg{mvtnorm}~1.0-11 packages. \proglang{R} itself
-and all dependent packages used are available from the Comprehensive
-\proglang{R} Archive Network (CRAN) at
-\url{https://CRAN.R-project.org/}. 
-The \pkg{graphsim} and \pkg{plot.igraph} packages presented can be installed from \url{https://github.com/TomKellyGenetics/graphsim} 
-and \url{https://github.com/TomKellyGenetics/plot.igraph} respectively.
-These functions can also be installed using the \pkg{igraph.extensions} library at \url{https://github.com/TomKellyGenetics/igraph.extensions} which includes other plotting functions used.
-This software is cross-platform and compatible with \proglang{R} installations on Windows, Mac, and Linux operating systems.
-The package GitHub repository also contains Vignettes with more information and examples on running functions released in the \proglang{R} package.
-The package (\pkg{graphsim}~0.1.0) meets CRAN submission criteria and will be released.
+\leavevmode\hypertarget{ref-Genz2009}{}%
+Genz, Alan, and Frank Bretz. 2009. ``Computation of Multivariate Normal
+and t Probabilities.'' In \emph{Lecture Notes in Statistics}. Vol. 195.
+Heidelberg: Springer-Verlag.
 
-\section*{Acknowledgements} %change to British spelling from American: ``Acknowledgments''
+\leavevmode\hypertarget{ref-mvtnorm}{}%
+Genz, Alan, Frank Bretz, Tetsuhisa Miwa, Xuefei Mi, Friedrich Leisch,
+Fabian Scheipl, and Torsten Hothorn. 2016. \emph{Mvtnorm: Multivariate
+Normal and T Distributions}.
+\url{http://CRAN.R-project.org/package=mvtnorm}.
 
-\begin{leftbar}
-This package was developed as part of a PhD research project funded by the Postgraduate Tassell Scholarship in Cancer Research Scholarship awarded to STK.
-We thank members of the Laboratory of Professor Satoru Miyano at the University of Tokyo, Institute for Medical Science, Professor Seiya Imoto, Associate Professor Rui Yamaguchi, and Dr Paul Sheridan (Assistant Professor at Hirosaki University,CSO at Tupac Bio) for helpful discussions in this field.
-We also thank Professor Parry Guilford at the University of Otago, Professor Cristin Print at the University of Auckland, and Dr Erik Arner at the RIKEN Center for Integrative Medical Sciences for their excellent advice during this project.
-\end{leftbar}
+\leavevmode\hypertarget{ref-Hawe2019}{}%
+Hawe, J. S., F. J. Theis, and M. Heinig. 2019. ``Inferring Interaction
+Networks From Multi-Omics Data.'' \emph{Front Genet} 10: 535.
 
-%% - Bibliography ---------------------
-%% - References need to be provided in a .bib BibTeX database.
-%% - All references should be made with \cite, \citet, \cite, \citealp etc.
-%%   (and never hard-coded). See the FAQ for details.
-%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
-%% - Titles in the .bib should be in title case.
-%% - DOIs should be included where available.
+\leavevmode\hypertarget{ref-Higham2002}{}%
+Higham, N. J. 2002. ``Computing the Nearest Correlation Matrix--a
+Problem from Finance.'' \emph{IMA Journal of Numerical Analysis} 22 (3).
+Oxford University Press: 329--43.
+\url{https://doi.org/10.1093/imanum/22.3.329}.
 
-\bibliography{refs}
+\leavevmode\hypertarget{ref-Hirose2008}{}%
+Hirose, Osamu, Ryo Yoshida, Seiya Imoto, Rui Yamaguchi, Tomoyuki
+Higuchi, D. Stephen Charnock-Jones, Cristin Print, and Satoru Miyano.
+2008. ``Statistical Inference of Transcriptional Module-Based Gene
+Networks from Time Course Gene Expression Profiles by Using State Space
+Models.'' \emph{Bioinformatics} 24 (7): 932--42.
+\url{https://doi.org/10.1093/bioinformatics/btm639}.
 
-%% - Appendix (if any) -------------------
-%% - After the bibliography with page break.
-%% - With proper section titles and _not_ just "Appendix".
+\leavevmode\hypertarget{ref-Hu2016}{}%
+Hu, J. X., C. E. Thomas, and S. Brunak. 2016. ``Network biology concepts
+in complex disease comorbidities.'' \emph{Nat. Rev. Genet.} 17 (10):
+615--29.
 
-\newpage
+\leavevmode\hypertarget{ref-Komatsu2013}{}%
+Komatsu, M., T. Yoshimaru, T. Matsuo, K. Kiyotani, Y. Miyoshi, T.
+Tanahashi, K. Rokutan, et al. 2013. ``Molecular features of triple
+negative breast cancer cells by genome-wide gene expression profiling
+analysis.'' \emph{Int. J. Oncol.} 42 (2): 478--506.
 
-%\begin{appendix}
+\leavevmode\hypertarget{ref-Law2014}{}%
+Law, C. W., Y. Chen, W. Shi, and G. K. Smyth. 2014. ``voom: Precision
+weights unlock linear model analysis tools for RNA-seq read counts.''
+\emph{Genome Biol.} 15 (2). Springer Nature: R29.
+\url{https://doi.org/10.1186/gb-2014-15-2-r29}.
 
-%\section{More technical details} \label{app:technical}
+\leavevmode\hypertarget{ref-Li2015}{}%
+Li, P., Y. Piao, H. S. Shon, and K. H. Ryu. 2015. ``Comparing the
+normalization methods for the differential analysis of Illumina
+high-throughput RNA-Seq data.'' \emph{BMC Bioinformatics} 16 (October):
+347.
 
-\begin{leftbar}
-%Would it be appropriate to include the Vignette for the package here?
-\iffalse
-Appendices can be included after the bibliography (with a page break). Each
-section within the appendix should have a proper section title (rather than
-just \emph{Appendix}).
+\leavevmode\hypertarget{ref-Markowetz2007}{}%
+Markowetz, F., and R. Spang. 2007. ``Inferring cellular networks--a
+review.'' \emph{BMC Bioinformatics} 8 Suppl 6 (September): S5.
 
-For more technical style details, please check out JSS's style FAQ at
-\url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
-which includes the following topics:
-\begin{itemize}
-  \item Title vs.\ sentence case.
-  \item Graphics formatting.
-  \item Naming conventions.
-  \item Turning JSS manuscripts into \proglang{R} package vignettes.
-  \item Trouble shooting.
-  \item Many other potentially helpful details\dots
-\end{itemize}
-\fi
-\end{leftbar}
+\leavevmode\hypertarget{ref-Ozsolak2011}{}%
+Ozsolak, F., and P. M. Milos. 2011. ``RNA sequencing: advances,
+challenges and opportunities.'' \emph{Nat. Rev. Genet.} 12 (2): 87--98.
 
-\iffalse
-\section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
+\leavevmode\hypertarget{ref-Perou2000}{}%
+Perou, C. M., T. Sørlie, M. B. Eisen, M. van de Rijn, S. S. Jeffrey, C.
+A. Rees, J. R. Pollack, et al. 2000. ``Molecular portraits of human
+breast tumours.'' \emph{Nature} 406 (6797): 747--52.
 
-\begin{leftbar}
-References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
-references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
-\verb|\citealp| etc.\ (and never hard-coded). This commands yield different
-formats of author-year citations and allow to include additional details (e.g.,
-pages, chapters, \dots) in brackets. In case you are not familiar with these
-commands see the JSS style FAQ for details.
+\leavevmode\hypertarget{ref-limma}{}%
+Ritchie, Matthew E, Belinda Phipson, Di Wu, Yifang Hu, Charity W Law,
+Wei Shi, and Gordon K Smyth. 2015. ``limma Powers Differential
+Expression Analyses for RNA-Sequencing and Microarray Studies.''
+\emph{Nucleic Acids Research} 43 (7): e47.
 
-Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task - especially
-when acquiring the entries automatically from mixed online sources. However,
-it is important that informations are complete and presented in a consistent
-style to avoid confusions. JSS requires the following format.
-\begin{itemize}
-  \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
-    be used in the references.
-  \item Titles should be in title case.
-  \item Journal titles should not be abbreviated and in title case.
-  \item DOIs should be included where available.
-  \item Software should be properly cited as well. For \proglang{R} packages
-    \code{citation("pkgname")} typically provides a good starting point.
-\end{itemize}
-\end{leftbar}
-\fi
+\leavevmode\hypertarget{ref-Shimamura2009}{}%
+Shimamura, Teppei, Seiya Imoto, Rui Yamaguchi, André Fujita, Masao
+Nagasaki, and Satoru Miyano. 2009. ``Recursive Regularization for
+Inferring Gene Networks from Time-Course Gene Expression Profiles.''
+\emph{BMC Systems Biology} 3 (1): 41.
+\url{https://doi.org/10.1186/1752-0509-3-41}.
 
-%\end{appendix}
+\leavevmode\hypertarget{ref-Svensson2018}{}%
+Svensson, V., R. Vento-Tormo, and S. A. Teichmann. 2018. ``Exponential
+scaling of single-cell RNA-seq in the past decade.'' \emph{Nat Protoc}
+13 (4): 599--604.
 
-%% --------------------------
+\leavevmode\hypertarget{ref-Wang2018}{}%
+Wang, J., M. Huang, E. Torre, H. Dueck, S. Shaffer, J. Murray, A. Raj,
+M. Li, and N. R. Zhang. 2018. ``Gene expression distribution
+deconvolution in single-cell RNA sequencing.'' \emph{Proc. Natl. Acad.
+Sci. U.S.A.} 115 (28): E6437--E6446.
+
+\leavevmode\hypertarget{ref-Yamaguchi2007}{}%
+Yamaguchi, Rui, Ryo Yoshida, Seiya Imoto, Tomoyuki Higuchi, and Satoru
+Miyano. 2007. ``Finding module-based gene networks with state-space
+models - Mining high-dimensional and short time-course gene expression
+data.'' \emph{IEEE Signal Processing Magazine} 24 (1): 37--46.
 

### 2020-05-09 paper/paper.Rmd
correct figure formatting and labels  (fig2-4, pane B-D)

@@
 options(width = 68, cli.unicode = FALSE, cli.width = 68)
-par(mai=c(1.02, 0.82, 0.82, 0.82)-0.82)
+#par(mai=c(2.82, 2.82, 0.82, 0.82)-0.82)
+par(mar=c(7, 10, 4, 2) + 0.1)
 par(bty="o")
@@
               cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
-mtext(text = "(a) Activating pathway structure", side=1, line=3, at=0.05, adj=0.5, cex=1.75)
+mtext(text = "(a) Activating pathway structure", side=1, line=3.5, at=0.05, adj=0.5, cex=1.75)
 box()
@@
               cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
-mtext(text = "(b) Inhibiting pathway structure", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+mtext(text = "(b) Inhibiting pathway structure", side=1, line=3.5, at=0.075, adj=0.5, cex=1.75)
 box()
@@
 
+```{r, include = FALSE}
+set.seed(9000)
+```
+
 ```{r simulation_activating, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}.
@@
               cex.node=2, cex.arrow=4, arrow_clip = 0.2)
-mtext(text = "(a) Activating pathway structure", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+mtext(text = "(a) Activating pathway structure", side=1, line=3.5, at=0.075, adj=0.5, cex=1.75)
 box()
@@
 
-op <- par(mfrow=c(2,2), mar=c(4,4,4,1), oma=c(0.5,0.5,0.5,0))
 #plot relationship matrix
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(b) relationship matrix", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+mtext(text = "(b) relationship matrix", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(d) Simulated correlation", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+mtext(text = "(d) Simulated correlation", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 
-op <- par(mfrow=c(1,1), mar=c(4,4,4,1), oma=c(0.5,0.5,0.5,0))
 #plot simulated expression data
 heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
-par(op)
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+mtext(text = "samples", side=1, line=1.5, at=0.2, adj=0.5, cex=1.5)
+mtext(text = "genes", side=4, line=1, at=-0.4, adj=0.5, cex=1.5)
+mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
+```
+
+```{r, include=FALSE}
+pdf("Plotsimulation_activating-5.pdf", width = 13.44, height = 6.72)
+#plot simulated expression data
+heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "", mar = c(6, 6), keysize = 1)
+mtext(text = "samples", side=1, line=1.5, at=0.55, adj=0.5, cex=1.5)
+mtext(text = "genes", side=4, line=1, at=0.45, adj=0.5, cex=1.5)
+mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0.5, adj=0.5, cex=1.75)
+dev.off()
+system("sed -i.bak '/Plotsimulation_activating-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.md && rm paper.md.bak")
+system("sed -i.bak '/Plotsimulation_activating-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.tex && rm paper.tex.bak")
 ```
@@
 
+```{r, include = FALSE}
+set.seed(9000)
+```
+
 ```{r simulation_inhibiting, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from graph structure with inhibitions}.
@@
               cex.node=2, cex.arrow=4, arrow_clip = 0.2)
-mtext(text = "(a) Inhibiting pathway structure", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+mtext(text = "(a) Inhibiting pathway structure", side=1, line=3.5, at=0.075, adj=0.5, cex=1.75)
 box()
@@
 
-# #plot relationship matrix
+#plot relationship matrix
 heatmap.2(make_distance_graph(graph, absolute = FALSE),
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(d) Simulated correlation", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+mtext(text = "(d) Simulated correlation", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 
@@
 heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+mtext(text = "samples", side=1, line=1.5, at=0.2, adj=0.5, cex=1.5)
+mtext(text = "genes", side=4, line=1, at=-0.4, adj=0.5, cex=1.5)
+mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
+```
+
+```{r, include=FALSE}
+pdf("Plotsimulation_inhibiting-5.pdf", width = 13.44, height = 6.72)
+#plot simulated expression data
+heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "", mar = c(6, 6), keysize = 1)
+mtext(text = "samples", side=1, line=1.5, at=0.55, adj=0.5, cex=1.5)
+mtext(text = "genes", side=4, line=1, at=0.45, adj=0.5, cex=1.5)
+mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0.5, adj=0.5, cex=1.75)
+dev.off()
+system("sed -i.bak '/Plotsimulation_inhibiting-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.md && rm paper.md.bak")
+system("sed -i.bak '/Plotsimulation_inhibiting-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.tex && rm paper.tex.bak")
 ```
@@
 functions as as an "inhibitor SMAD" with competitively inhibits *SMAD4*.
+
+```{r, include = FALSE}
+set.seed(9000)
+```
 
@@
           scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = "(b) relationship matrix", cex.sub = 1.75)
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+mtext(text = "(b) relationship matrix", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 
@@
 scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 
@@
 heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(d) Simulated correlation", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+mtext(text = "(d) Simulated correlation", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 
@@
 heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+mtext(text = "samples", side=1, line=1.5, at=0.2, adj=0.5, cex=1.5)
+mtext(text = "genes", side=4, line=1, at=-0.4, adj=0.5, cex=1.5)
+mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
+```
+
+```{r, include=FALSE}
+pdf("Plotsimulation_smad-5.pdf", width = 13.44, height = 6.72)
+#plot simulated expression data
+heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "", mar = c(6, 6), keysize = 1)
+mtext(text = "samples", side=1, line=1.5, at=0.55, adj=0.5, cex=1.5)
+mtext(text = "genes", side=4, line=1, at=0.45, adj=0.5, cex=1.5)
+mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0.5, adj=0.5, cex=1.75)
+dev.off()
+system("sed -i.bak '/Plotsimulation_smad-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.md && rm paper.md.bak")
+system("sed -i.bak '/Plotsimulation_smad-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.tex && rm paper.tex.bak")
 ```

### 2020-05-09 paper/paper.tex
correct figure formatting and labels  (fig2-4, pane B-D)

@@
               \DataTypeTok{cex.node =} \DecValTok{2}\NormalTok{, }\DataTypeTok{cex.arrow =} \DecValTok{4}\NormalTok{, }\DataTypeTok{arrow_clip =} \FloatTok{0.2}\NormalTok{)}
-\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(a) Activating pathway structure"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.05}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(a) Activating pathway structure"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.05}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 \KeywordTok{box}\NormalTok{()}
@@
               \DataTypeTok{cex.node =} \DecValTok{2}\NormalTok{, }\DataTypeTok{cex.arrow =} \DecValTok{4}\NormalTok{, }\DataTypeTok{arrow_clip =} \FloatTok{0.2}\NormalTok{)}
-\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(b) Inhibiting pathway structure"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(b) Inhibiting pathway structure"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 \KeywordTok{box}\NormalTok{()}
@@
               \DataTypeTok{cex.node=}\DecValTok{2}\NormalTok{, }\DataTypeTok{cex.arrow=}\DecValTok{4}\NormalTok{, }\DataTypeTok{arrow_clip =} \FloatTok{0.2}\NormalTok{)}
-\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(a) Activating pathway structure"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(a) Activating pathway structure"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 \KeywordTok{box}\NormalTok{()}
@@
 
-\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow=}\KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{), }\DataTypeTok{mar=}\KeywordTok{c}\NormalTok{(}\DecValTok{4}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{1}\NormalTok{), }\DataTypeTok{oma=}\KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{,}\FloatTok{0.5}\NormalTok{,}\FloatTok{0.5}\NormalTok{,}\DecValTok{0}\NormalTok{))}
 \CommentTok{#plot relationship matrix}
@@
 \DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
-\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(b) relationship matrix"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(b) relationship matrix"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\DecValTok{0}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 
@@
 \DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
-\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"(c) "}\NormalTok{, Sigma, }\StringTok{" matrix"}\NormalTok{)), }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"(c) "}\NormalTok{, Sigma, }\StringTok{" matrix"}\NormalTok{)), }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\DecValTok{0}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 
@@
 \DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
-\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(d) Simulated correlation"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(d) Simulated correlation"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\DecValTok{0}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 
-\NormalTok{op <-}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow=}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{), }\DataTypeTok{mar=}\KeywordTok{c}\NormalTok{(}\DecValTok{4}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{1}\NormalTok{), }\DataTypeTok{oma=}\KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{,}\FloatTok{0.5}\NormalTok{,}\FloatTok{0.5}\NormalTok{,}\DecValTok{0}\NormalTok{))}
 \CommentTok{#plot simulated expression data}
 \KeywordTok{heatmap.2}\NormalTok{(expr, }\DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{bluered}\NormalTok{(}\DecValTok{50}\NormalTok{),}
-\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
-\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(e) Simulated expression data (log scale)"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
-\KeywordTok{par}\NormalTok{(op)}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{labCol =} \StringTok{""}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"samples"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{1.5}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.2}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.5}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"genes"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{4}\NormalTok{, }\DataTypeTok{line=}\DecValTok{1}\NormalTok{, }\DataTypeTok{at=}\OperatorTok{-}\FloatTok{0.4}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.5}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(e) Simulated expression data (log scale)"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\DecValTok{0}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 \end{Highlighting}
@@
               \DataTypeTok{cex.node=}\DecValTok{2}\NormalTok{, }\DataTypeTok{cex.arrow=}\DecValTok{4}\NormalTok{, }\DataTypeTok{arrow_clip =} \FloatTok{0.2}\NormalTok{)}
-\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(a) Inhibiting pathway structure"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(a) Inhibiting pathway structure"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 \KeywordTok{box}\NormalTok{()}
@@
 
-\CommentTok{# #plot relationship matrix}
+\CommentTok{#plot relationship matrix}
 \KeywordTok{heatmap.2}\NormalTok{(}\KeywordTok{make_distance_graph}\NormalTok{(graph, }\DataTypeTok{absolute =} \OtherTok{FALSE}\NormalTok{),}
@@
 \DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
-\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(d) Simulated correlation"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(d) Simulated correlation"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\DecValTok{0}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 
@@
 \KeywordTok{heatmap.2}\NormalTok{(expr, }\DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{bluered}\NormalTok{(}\DecValTok{50}\NormalTok{),}
-\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
-\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(e) Simulated expression data (log scale)"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{labCol =} \StringTok{""}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"samples"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{1.5}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.2}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.5}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"genes"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{4}\NormalTok{, }\DataTypeTok{line=}\DecValTok{1}\NormalTok{, }\DataTypeTok{at=}\OperatorTok{-}\FloatTok{0.4}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.5}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(e) Simulated expression data (log scale)"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\DecValTok{0}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 \end{Highlighting}
@@
           \DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{colorpanel}\NormalTok{(}\DecValTok{50}\NormalTok{, }\StringTok{"white"}\NormalTok{, }\StringTok{"red"}\NormalTok{),}
-\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)),}
-              \DataTypeTok{sub =} \StringTok{"(b) relationship matrix"}\NormalTok{, }\DataTypeTok{cex.sub =} \FloatTok{1.75}\NormalTok{)}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{labCol =} \StringTok{""}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(b) relationship matrix"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\DecValTok{0}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 
@@
 \DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{colorpanel}\NormalTok{(}\DecValTok{50}\NormalTok{, }\StringTok{"blue"}\NormalTok{, }\StringTok{"white"}\NormalTok{, }\StringTok{"red"}\NormalTok{),}
-\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{labCol =} \StringTok{""}\NormalTok{)}
 \CommentTok{#> Warning in eattrs[[name]][index] <- value: number of items to}
 \CommentTok{#> replace is not a multiple of replacement length}
-\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"(c) "}\NormalTok{, Sigma, }\StringTok{" matrix"}\NormalTok{)), }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"(c) "}\NormalTok{, Sigma, }\StringTok{" matrix"}\NormalTok{)), }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\DecValTok{0}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 
@@
 \KeywordTok{heatmap.2}\NormalTok{(}\KeywordTok{cor}\NormalTok{(}\KeywordTok{t}\NormalTok{(expr)), }\DataTypeTok{scale =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{trace =} \StringTok{"none"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{colorpanel}\NormalTok{(}\DecValTok{50}\NormalTok{, }\StringTok{"blue"}\NormalTok{, }\StringTok{"white"}\NormalTok{, }\StringTok{"red"}\NormalTok{),}
-\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
-\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(d) Simulated correlation"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\DecValTok{3}\NormalTok{, }\DataTypeTok{at=}\FloatTok{0.075}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{labCol =} \StringTok{""}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(d) Simulated correlation"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\DecValTok{0}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 

### 2020-05-09 paper/paper.Rmd
correct figure subcaptions

@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(b) relationship matrix", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
+mtext(text = "(b) Relationship matrix", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 
@@
 dev.off()
-system("sed -i.bak '/Plotsimulation_activating-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.md && rm paper.md.bak")
-system("sed -i.bak '/Plotsimulation_activating-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.tex && rm paper.tex.bak")
+#system("sed -i.bak '/Plotsimulation_activating-5/s/\\(.*\\)width=[.]415/\1width=.830/g' paper.md && rm paper.md.bak")
+#system("sed -i.bak '/Plotsimulation_activating-5/s/\\(.*\\)width=[.]415/\1width=.830/g' paper.tex && rm paper.tex.bak")
+#sed -i '/Plot.*-5/s/\(.*\)width=[.]415/\1width=.830/g' paper.md paper.tex
 ```
@@
 dev.off()
-system("sed -i.bak '/Plotsimulation_inhibiting-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.md && rm paper.md.bak")
-system("sed -i.bak '/Plotsimulation_inhibiting-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.tex && rm paper.tex.bak")
+#("sed -i.bak '/Plotsimulation_inhibiting-5/s/\\(.*\\)width=[.]415/\1width=.830/g' paper.md && rm paper.md.bak")
+#system("sed -i.bak '/Plotsimulation_inhibiting-5/s/\\(.*\\)width=[.]415/\1width=.830/g' paper.tex && rm paper.tex.bak")
+#sed -i '/Plot.*-5/s/\(.*\)width=[.]415/\1width=.830/g' paper.md paper.tex
 ```
@@
               cex.node = 1.5, cex.label = 0.8, cex.arrow = 2, 
-              sub = expression(paste("(a) TFG-", Beta, " activates SMADs")), cex.sub = 1.75)
+              sub = expression(paste("(a) TFG-", beta, " activates SMADs")), cex.sub = 1.75)
 box()
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
-mtext(text = "(b) relationship matrix", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
+mtext(text = "(b) Relationship matrix", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 
@@
 dev.off()
-system("sed -i.bak '/Plotsimulation_smad-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.md && rm paper.md.bak")
-system("sed -i.bak '/Plotsimulation_smad-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.tex && rm paper.tex.bak")
+#system("sed -i.bak '/Plotsimulation_smad-5/s/\\(.*\\)width=[.]415/\1width=.830/g' paper.md && rm paper.md.bak")
+#system("sed -i.bak '/Plotsimulation_smad-5/s/\\(.*\\)width=[.]415/\1width=.830/g' paper.tex && rm paper.tex.bak")
+#sed -i '/Plot.*-5/s/\(.*\)width=[.]415/\1width=.830/g' paper.md paper.tex
 ```

### 2020-05-09 paper/paper.tex
correct figure subcaptions

@@
 \DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)))}
-\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(b) relationship matrix"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\DecValTok{0}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(b) Relationship matrix"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\DecValTok{0}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 
@@
               \DataTypeTok{cex.node =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.label =} \FloatTok{0.8}\NormalTok{, }\DataTypeTok{cex.arrow =} \DecValTok{2}\NormalTok{, }
-              \DataTypeTok{sub =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"(a) TFG-"}\NormalTok{, Beta, }\StringTok{" activates SMADs"}\NormalTok{)), }\DataTypeTok{cex.sub =} \FloatTok{1.75}\NormalTok{)}
+              \DataTypeTok{sub =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"(a) TFG-"}\NormalTok{, beta, }\StringTok{" activates SMADs"}\NormalTok{)), }\DataTypeTok{cex.sub =} \FloatTok{1.75}\NormalTok{)}
 \KeywordTok{box}\NormalTok{()}
@@
 \DataTypeTok{colsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{rowsep =} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(}\KeywordTok{V}\NormalTok{(graph)), }\DataTypeTok{labCol =} \StringTok{""}\NormalTok{)}
-\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(b) relationship matrix"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\DecValTok{0}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
+\KeywordTok{mtext}\NormalTok{(}\DataTypeTok{text =} \StringTok{"(b) Relationship matrix"}\NormalTok{, }\DataTypeTok{side=}\DecValTok{1}\NormalTok{, }\DataTypeTok{line=}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{at=}\DecValTok{0}\NormalTok{, }\DataTypeTok{adj=}\FloatTok{0.5}\NormalTok{, }\DataTypeTok{cex=}\FloatTok{1.75}\NormalTok{)}
 

### 2020-05-13 R/distance.R
correct passing state for inhibiting graph attributes (migrate to state computing in sigma rather than generate)

@@ -30,9 +30,13 @@ make_distance_graph <- function(graph, directed = FALSE, absolute = FALSE){
   if(directed == FALSE) graph <- as.undirected(graph)
   diam <- diameter(graph)
   if (absolute){
-    mat <- (diam-shortest.paths(graph))/diam
+    paths <- shortest.paths(graph)
+    diam <- max(diam, max(paths))
+    mat <- (diam-paths)/diam
   } else {
-    mat <- 1^-diam/(diam*shortest.paths(graph))
+    paths <- shortest.paths(graph)
+    diam <- max(diam, max(paths))
+    mat <- 1^-diam/(diam*paths)
     diag(mat) <- 1
   }
   rownames(mat) <- colnames(mat) <- names(V(graph))
@@ -47,9 +51,13 @@ make_distance_adjmat <- function(mat, directed = FALSE, absolute = FALSE){
   graph <- graph_from_adjacency_matrix(mat, weighted = NULL, mode = ifelse(directed, "directed", "undirected"))
   diam <- diameter(graph)
   if (absolute){
-    mat <- (diam-shortest.paths(graph))/diam
+    paths <- shortest.paths(graph)
+    diam <- max(diam, max(paths))
+    mat <- (diam-paths)/diam
   } else {
-    mat <- 1^-diam/(diam*shortest.paths(graph))
+    paths <- shortest.paths(graph)
+    diam <- max(diam, max(paths))
+    mat <- 1^-diam/(diam*paths)
     diag(mat) <- 1
   }
   rownames(mat) <- colnames(mat) <- names(V(graph))
@@ -65,9 +73,13 @@ make_distance_laplacian <- function(mat, directed = FALSE, absolute = FALSE){
   graph <- graph_from_adjacency_matrix(adj_mat, weighted = TRUE, mode = ifelse(directed, "directed", "undirected"))
   diam <- diameter(graph)
   if (absolute){
-    mat <- (diam-shortest.paths(graph))/diam
+    paths <- shortest.paths(graph)
+    diam <- max(diam, max(paths))
+    mat <- (diam-paths)/diam
   } else {
-    mat <- 1^-diam/(diam*shortest.paths(graph))
+    paths <- shortest.paths(graph)
+    diam <- max(diam, max(paths))
+    mat <- 1^-diam/(diam*paths)
     diag(mat) <- 1
   }
   rownames(mat) <- colnames(mat) <- names(V(graph))

### 2020-05-13 R/generate.R

@@ -36,6 +36,14 @@
 ##' 
 ##' @export
 generate_expression <- function(n, graph, state = NULL, cor = 0.8, mean = 0, comm = FALSE, dist = FALSE, absolute = FALSE, laplacian = FALSE){
+  if(!is.null(get.edge.attribute(graph, "state"))){
+    state <- get.edge.attribute(graph, "state")
+  } else {
+    # add default state if not specified
+    if(is.null(state)){
+      state <- rep(1, length(E(graph)))
+    }
+  }
   if(!is.integer(n)){
     if(is.numeric(n)){
       if(floor(n) == n){
@@ -50,15 +58,19 @@ generate_expression <- function(n, graph, state = NULL, cor = 0.8, mean = 0, com
   }
   if(length(n) > 1) stop("sample size n must be an integer of length 1")
   if(!is_igraph(graph)) stop("graph must be an igraph class")
-  if(is.vector(state) || length(state) == 1) state <- make_state_matrix(graph, state)
+  if(length(state) == 1) state <- rep(state, length(E(graph)))
   if(!(is.vector(mean)) || length(mean) == 1 ) mean <- rep(mean,length(V(graph)))
   if(dist){
-    sig <- make_sigma_mat_dist_graph(graph, cor, absolute = absolute)
+    sig <- make_sigma_mat_dist_graph(graph, state = state, cor, absolute = absolute)
   } else {
-    sig <- make_sigma_mat_graph(graph, cor, comm = comm, laplacian = laplacian)
+    sig <- make_sigma_mat_graph(graph, state = state, cor, comm = comm, laplacian = laplacian)
   }
   ## migrate state to calling sigma ##
-  if(!(is.null(state))) sig <- state * sig
+  if(!(is.null(state))){
+    if(all(sig >= 0)){
+      sig <- state * sig
+    }
+  }
   if(is.symmetric.matrix(sig) == FALSE) {
     warning("sigma matrix was not positive definite, nearest approximation used.")
     sig <- as.matrix(nearPD(sig, corr=T, keepDiag = TRUE)$mat) #postive definite correction
@@ -77,6 +89,9 @@ generate_expression <- function(n, graph, state = NULL, cor = 0.8, mean = 0, com
 ##' @rdname generate_expression
 ##' @export
[…]

### 2020-05-13 R/sigma.R

@@ -30,9 +30,14 @@
 ##' 
 ##' @return a numeric covariance matrix of values in the range [-1, 1]
 ##' @export
-make_sigma_mat_adjmat <- function(mat, cor = 0.8){
+make_sigma_mat_adjmat <- function(mat, state = NULL, cor = 0.8){
   sig <- ifelse(mat > 0, cor, 0)
   diag(sig) <- 1
+  if(!is.null(state)){
+    #pass state parameters sign of sigma
+    state_mat <- make_state_matrix(graph, state)
+    sig <- sig * sign(state_mat)
+  }
   rownames(sig) <- rownames(mat)
   colnames(sig) <- colnames(mat)
   return(sig)
@@ -40,12 +45,17 @@ make_sigma_mat_adjmat <- function(mat, cor = 0.8){
 
 ##' @rdname make_sigma
 ##' @export
-make_sigma_mat_comm <- function(mat, cor = 0.8){
+make_sigma_mat_comm <- function(mat, state = NULL, cor = 0.8){
   mat <- abs(mat)
   mat <- apply(mat, 1, function(x) x/max(x))
   mat <- apply(mat, 2, function(x) x/max(x))
   sig <- ifelse(mat>0, cor*mat/max(mat), 0)
   diag(sig) <- 1
+  if(!is.null(state)){
+    #pass state parameters sign of sigma
+    state_mat <- make_state_matrix(graph, state)
+    sig <- sig * sign(state_mat)
+  }
   rownames(sig) <- rownames(mat)
   colnames(sig) <- colnames(mat)
   return(sig)
@@ -53,13 +63,18 @@ make_sigma_mat_comm <- function(mat, cor = 0.8){
 
 ##' @rdname make_sigma
 ##' @export
-make_sigma_mat_laplacian <- function(mat, cor = 0.8){
+make_sigma_mat_laplacian <- function(mat, state = NULL, cor = 0.8){
   mat <- abs(mat)
   diag(mat)[diag(mat) == 0] <- 1
   mat <- apply(mat, 1, function(x) x/max(x))
   mat <- apply(mat, 2, function(x) x/max(x))
   sig <- ifelse(mat>0, cor*mat/max(mat), 0)
   diag(sig) <- 1
+  if(!is.null(state)){
+    #pass state parameters sign of sigma
+    state_mat <- make_state_matrix(graph, state)
+    sig <- sig * sign(state_mat)
+  }
   rownames(sig) <- rownames(mat)
   colnames(sig) <- colnames(mat)
[…]

### 2020-05-25 paper/images.sh
handling paper formatting

@@ -1,0 +1,1 @@
+sed -i '/Plot.*-5/s/\(.*\)width=[.]415/\1width=.830/g' paper.md paper.tex

### 2020-05-25 paper/paper.Rmd
format code in paper
reformat code examples

--- 2020-05-08 paper/paper.Rmd
@@
 #install packages required (once per machine)
-
 install.packages("graphsim")
 ```
-
-```{r}
+```{r, warning=FALSE, results='hide'}
 #load required packages (once per R instance)
-
-library("igraph")
-library("graphsim")
-library("gplots")
+library("graphsim"); library("igraph"); library("gplots")
 ```
@@
 #generate graph structure
-
 graph_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"), c("D", "E"),
@@
 #plot graph structure (Figure 1a)
-
 plot_directed(graph, state ="activating", layout = layout.kamada.kawai,
               cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
-mtext(text = "(a) Activating pathway structure", side=1, line=3, at=0.05, adj=0.5, cex=1.75)
+mtext(text = "(a) Activating pathway structure", side=1, line=3.5, at=0.05, adj=0.5, cex=1.75)
 box()
@@
 #generate parameters for inhibitions
-
 state <- c(1, 1, -1, 1, 1, 1, 1, -1, 1)
-
 #plot graph structure with inhibitions (Figure 1b)
-
 plot_directed(graph, state=state, layout = layout.kamada.kawai,
               cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
-mtext(text = "(b) Inhibiting pathway structure", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+mtext(text = "(b) Inhibiting pathway structure", side=1, line=3.5, at=0.075, adj=0.5, cex=1.75)
 box()
@@
 
+```{r, include = FALSE}
+set.seed(9000)
+```
+
+```{r simulation_activating_setup, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, warning=FALSE}
+# activating graph
+state <- rep(1, length(E(graph)))
+plot_directed(graph, state=state, layout = layout.kamada.kawai,
+              cex.node=2, cex.arrow=4, arrow_clip = 0.2)
+mtext(text = "(a) Activating pathway structure", side=1, line=3.5, at=0.075, adj=0.5, cex=1.75)
+box()
+#adjacency matrix
+adj_mat <- make_adjmatrix_graph(graph)
+#relationship matrix
+dist_mat <- make_distance_graph(graph, absolute = FALSE)
+#sigma matrix directly from graph
+sigma_mat <- make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE)
+#show shortest paths of graph
+shortest_paths <- shortest.paths(graph)
+#generate expression data directly from graph
+expr <- generate_expression(100, graph, cor = 0.8, mean = 0, comm = FALSE,
+                            dist = TRUE, absolute = FALSE, state = state)
+```
+
 ```{r simulation_activating, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}.
@@
 This $\\Sigma$ matrix has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
-Here genes with closer relationships in the pathway structure show a higher correlation between simulated values.'}
-# activating graph
-state <- rep(1, length(E(graph)))
-plot_directed(graph, state=state, layout = layout.kamada.kawai,
-              cex.node=2, cex.arrow=4, arrow_clip = 0.2)
-mtext(text = "(a) Activating pathway structure", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
-box()
+Here genes with closer relationships in the pathway structure show a higher correlation between simulated values.', warning=FALSE}
 
-#adjacency matrix
-adj_mat <- make_adjmatrix_graph(graph)
-
-#relationship matrix
-dist_mat <- make_distance_graph(graph, absolute = FALSE)
-
-#sigma matrix directly from graph
-sigma_mat <- make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE)
-
-#show shortest paths of graph
-shortest_paths <- shortest.paths(graph)
-
-#generate expression data directly from graph
-expr <- generate_expression(100, graph, cor = 0.8, mean = 0, comm = FALSE,
-                            dist = TRUE, absolute = FALSE, state = state)
-
-op <- par(mfrow=c(2,2), mar=c(4,4,4,1), oma=c(0.5,0.5,0.5,0))
 #plot relationship matrix
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(b) relationship matrix", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
-
+mtext(text = "(b) Relationship matrix", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 #plot sigma matrix
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3, at=0.075, adj=0.5, cex=1.75)
-
+mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 #simulated data
@@
 comm = FALSE, dist =TRUE, absolute = FALSE, state = state)
-
 #plot simulated correlations
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(d) Simulated correlation", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
-
-op <- par(mfrow=c(1,1), mar=c(4,4,4,1), oma=c(0.5,0.5,0.5,0))
+mtext(text = "(d) Simulated correlation", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 #plot simulated expression data
 heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
-par(op)
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+mtext(text = "samples", side=1, line=1.5, at=0.2, adj=0.5, cex=1.5)
+mtext(text = "genes", side=4, line=1, at=-0.4, adj=0.5, cex=1.5)
+mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
+```
+
+```{r, include=FALSE, warning=FALSE}
+pdf("Plotsimulation_activating-5.pdf", width = 13.44, height = 6.72)
+#plot simulated expression data
+heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "", mar = c(6, 6), keysize = 1)
+mtext(text = "samples", side=1, line=1.5, at=0.55, adj=0.5, cex=1.5)
+mtext(text = "genes", side=4, line=1, at=0.45, adj=0.5, cex=1.5)
+mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0.5, adj=0.5, cex=1.75)
+dev.off()
+#system("sed -i.bak '/Plotsimulation_activating-5/s/\\(.*\\)width=[.]415/\1width=.830/g' paper.md && rm paper.md.bak")
+#system("sed -i.bak '/Plotsimulation_activating-5/s/\\(.*\\)width=[.]415/\1width=.830/g' paper.tex && rm paper.tex.bak")
+#sed -i '/Plot.*-5/s/\(.*\)width=[.]415/\1width=.830/g' paper.md paper.tex
 ```
@@
 
+```{r, include = FALSE}
+set.seed(9000)
+```
+
+```{r simulation_inhibiting_setup, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, warning=FALSE}
+#generate parameters for inhibitions
+state <- c(1, 1, -1, 1, 1, 1, 1, -1)
+plot_directed(graph, state=state, layout = layout.kamada.kawai,
+              cex.node=2, cex.arrow=4, arrow_clip = 0.2)
+mtext(text = "(a) Inhibiting pathway structure", side=1, line=3.5, at=0.075, adj=0.5, cex=1.75)
+box()
+#adjacency matrix
+adj_mat <- make_adjmatrix_graph(graph)
+#relationship matrix
+dist_mat <- make_distance_graph(graph, absolute = FALSE)
+#sigma matrix directly from graph
+sigma_mat <- make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE)
+#show shortest paths of graph
+shortest_paths <- shortest.paths(graph)
+#generate expression data directly from graph
+expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0, comm = FALSE,
+                            dist = TRUE, absolute = FALSE)
+```
 ```{r simulation_inhibiting, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from graph structure with inhibitions}.
@@
 This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
-Here the inhibitory relationships between genes are reflected in negatively correlated simulated  values.'}
-
-#generate parameters for inhibitions
-state <- c(1, 1, -1, 1, 1, 1, 1, -1)
-plot_directed(graph, state=state, layout = layout.kamada.kawai,
-              cex.node=2, cex.arrow=4, arrow_clip = 0.2)
-mtext(text = "(a) Inhibiting pathway structure", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
-box()
-
-#adjacency matrix
-adj_mat <- make_adjmatrix_graph(graph)
-
-#relationship matrix
-dist_mat <- make_distance_graph(graph, absolute = FALSE)
-
-#sigma matrix directly from graph
-sigma_mat <- make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE)
-
-#show shortest paths of graph
-shortest_paths <- shortest.paths(graph)
-
-#generate expression data directly from graph
-expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0, comm = FALSE,
-                            dist = TRUE, absolute = FALSE)
-
-# #plot relationship matrix
+Here the inhibitory relationships between genes are reflected in negatively correlated simulated  values.', warning=FALSE}
+#plot relationship matrix
 heatmap.2(make_distance_graph(graph, absolute = FALSE),
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(b) relationship matrix", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
-
+mtext(text = "(b) Relationship matrix", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 # #plot sigma matrix
@@
 comm = FALSE, dist =TRUE, absolute = FALSE)
-
 #plot simulated correlations
@@
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(d) Simulated correlation", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
-
+mtext(text = "(d) Simulated correlation", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 #plot simulated expression data
 heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+mtext(text = "samples", side=1, line=1.5, at=0.2, adj=0.5, cex=1.5)
+mtext(text = "genes", side=4, line=1, at=-0.4, adj=0.5, cex=1.5)
+mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
+```
+
+```{r, include=FALSE, warning=FALSE}
+pdf("Plotsimulation_inhibiting-5.pdf", width = 13.44, height = 6.72)
+#plot simulated expression data
+heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "", mar = c(6, 6), keysize = 1)
+mtext(text = "samples", side=1, line=1.5, at=0.55, adj=0.5, cex=1.5)
+mtext(text = "genes", side=4, line=1, at=0.45, adj=0.5, cex=1.5)
+mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0.5, adj=0.5, cex=1.75)
+dev.off()
+#("sed -i.bak '/Plotsimulation_inhibiting-5/s/\\(.*\\)width=[.]415/\1width=.830/g' paper.md && rm paper.md.bak")
+#system("sed -i.bak '/Plotsimulation_inhibiting-5/s/\\(.*\\)width=[.]415/\1width=.830/g' paper.tex && rm paper.tex.bak")
+#sed -i '/Plot.*-5/s/\(.*\)width=[.]415/\1width=.830/g' paper.md paper.tex
 ```
@@
 
-```{r simulation_smad, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a biological pathway graph structure}.
-The graph structure (a) of a known biological pathway, the TGF-$\\beta$ receptor signaling activates SMADs (R-HSA-2173789), was used to derive a relationship matrix (b), $\\Sigma$ matrix (c) and correlation structure (d) from the relative distances between the nodes.
-These values are coloured blue to red from $-1$ to $1$.
-This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
-Here modules of genes with correlated expression can be clearly discerned.'}
+```{r, include = FALSE}
+set.seed(9000)
+```
 
+```{r simulation_smad_setup, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2,  warning=FALSE}
 #import graph from data
 graph <- identity(TGFBeta_Smad_graph)
-
 #generate parameters for inhibitions
@@
 state[is.na(state)] <- 1
-
 plot_directed(graph, state = state, layout = layout.kamada.kawai,
@@
               cex.node = 1.5, cex.label = 0.8, cex.arrow = 2, 
-              sub = expression(paste("(a) TFG-", Beta, " activates SMADs")), cex.sub = 1.75)
+              sub = expression(paste("(a) TFG-", beta, " activates SMADs")), cex.sub = 1.75)
 box()
-
 #adjacency matrix
 adj_mat <- make_adjmatrix_graph(graph)
-
 #relationship matrix
 dist_mat <- make_distance_graph(graph, absolute = FALSE)
-
 #sigma matrix directly from graph
 sigma_mat <- make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE)
-
 #show shortest paths of graph
 shortest_paths <- shortest.paths(graph)
-
 #generate expression data directly from graph
@@
                             dist = TRUE, absolute = FALSE)
-
+```
+```{r simulation_smad, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a biological pathway graph structure}.
+The graph structure (a) of a known biological pathway, the TGF-$\\beta$ receptor signaling activates SMADs (R-HSA-2173789), was used to derive a relationship matrix (b), $\\Sigma$ matrix (c) and correlation structure (d) from the relative distances between the nodes.
+These values are coloured blue to red from $-1$ to $1$.
+This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
+Here modules of genes with correlated expression can be clearly discerned.', warning=FALSE}
 # #plot relationship matrix
@@
           scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
-              sub = "(b) relationship matrix", cex.sub = 1.75)
-
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+mtext(text = "(b) Relationship matrix", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 # #plot sigma matrix
@@
 scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3, at=0.075, adj=0.5, cex=1.75)
-
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 #simulated data
@@
 comm = FALSE, dist =TRUE, absolute = FALSE)
-
 #plot simulated correlations
 heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(d) Simulated correlation", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
-
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+mtext(text = "(d) Simulated correlation", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 #plot simulated expression data
 heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
-colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+mtext(text = "samples", side=1, line=1.5, at=0.2, adj=0.5, cex=1.5)
+mtext(text = "genes", side=4, line=1, at=-0.4, adj=0.5, cex=1.5)
+mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
+```
+
+```{r, include=FALSE, warning=FALSE}
+pdf("Plotsimulation_smad-5.pdf", width = 13.44, height = 6.72)
+#plot simulated expression data
+heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "", mar = c(6, 6), keysize = 1)
+mtext(text = "samples", side=1, line=1.5, at=0.55, adj=0.5, cex=1.5)
+mtext(text = "genes", side=4, line=1, at=0.45, adj=0.5, cex=1.5)
+mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0.5, adj=0.5, cex=1.75)
+dev.off()
+#system("sed -i.bak '/Plotsimulation_smad-5/s/\\(.*\\)width=[.]415/\1width=.830/g' paper.md && rm paper.md.bak")
+#system("sed -i.bak '/Plotsimulation_smad-5/s/\\(.*\\)width=[.]415/\1width=.830/g' paper.tex && rm paper.tex.bak")
+#sed -i '/Plot.*-5/s/\(.*\)width=[.]415/\1width=.830/g' paper.md paper.tex
 ```

### 2020-05-25 paper/paper.Rmd
update code examples

@@
 
-```{r simulation_activating_setup, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, warning=FALSE}
+```{r simulation_activating, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}.
+An example of a graph structure (a) that has been used to derive a relationship matrix (b), $\\Sigma$  matrix (c) and correlation structure (d) from the relative distances between the nodes.
+Non-negative values are coloured white to red from $0$ to $1$.
+This $\\Sigma$ matrix has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
+Here genes with closer relationships in the pathway structure show a higher correlation between simulated values.', warning=FALSE}
 # activating graph
@@
 box()
-#adjacency matrix
-adj_mat <- make_adjmatrix_graph(graph)
-#relationship matrix
-dist_mat <- make_distance_graph(graph, absolute = FALSE)
-#sigma matrix directly from graph
-sigma_mat <- make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE)
-#show shortest paths of graph
-shortest_paths <- shortest.paths(graph)
-#generate expression data directly from graph
-expr <- generate_expression(100, graph, cor = 0.8, mean = 0, comm = FALSE,
-                            dist = TRUE, absolute = FALSE, state = state)
-```
-
-```{r simulation_activating, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}.
-An example of a graph structure (a) that has been used to derive a relationship matrix (b), $\\Sigma$  matrix (c) and correlation structure (d) from the relative distances between the nodes.
-Non-negative values are coloured white to red from $0$ to $1$.
-This $\\Sigma$ matrix has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
-Here genes with closer relationships in the pathway structure show a higher correlation between simulated values.', warning=FALSE}
-
 #plot relationship matrix
@@
 
-```{r simulation_inhibiting_setup, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, warning=FALSE}
+```{r simulation_inhibiting, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from graph structure with inhibitions}.
+An example of a graph structure (a), that has been used to derive a relationship matrix (b), $\\Sigma$ matrix (c), and correlation structure (d), from the relative distances between the nodes.
+These values are coloured blue to red from $-1$ to $1$.
+This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
+Here the inhibitory relationships between genes are reflected in negatively correlated simulated  values.', warning=FALSE}
 #generate parameters for inhibitions
@@
 box()
-#adjacency matrix
-adj_mat <- make_adjmatrix_graph(graph)
-#relationship matrix
-dist_mat <- make_distance_graph(graph, absolute = FALSE)
-#sigma matrix directly from graph
-sigma_mat <- make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE)
-#show shortest paths of graph
-shortest_paths <- shortest.paths(graph)
-#generate expression data directly from graph
-expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0, comm = FALSE,
-                            dist = TRUE, absolute = FALSE)
-```
-```{r simulation_inhibiting, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from graph structure with inhibitions}.
-An example of a graph structure (a), that has been used to derive a relationship matrix (b), $\\Sigma$ matrix (c), and correlation structure (d), from the relative distances between the nodes.
-These values are coloured blue to red from $-1$ to $1$.
-This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
-Here the inhibitory relationships between genes are reflected in negatively correlated simulated  values.', warning=FALSE}
 #plot relationship matrix
@@
 
-```{r simulation_smad_setup, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2,  warning=FALSE}
+```{r simulation_smad, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a biological pathway graph structure}.
+The graph structure (a) of a known biological pathway, the TGF-$\\beta$ receptor signaling activates SMADs (R-HSA-2173789), was used to derive a relationship matrix (b), $\\Sigma$ matrix (c) and correlation structure (d) from the relative distances between the nodes.
+These values are coloured blue to red from $-1$ to $1$.
+This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
+Here modules of genes with correlated expression can be clearly discerned.', warning=FALSE}
 #import graph from data
@@
 #generate parameters for inhibitions
-state <- rep(1, length(E(graph)))
-pathway <- get.edgelist(graph)
-state[pathway[,1] %in% c("SMAD6", "SMAD7", "BAMBI", "SMURF1", "SMURF2", "UCHL5", "USP15", "UBB", "UBC", "PMEPA1", "PPP1CA", "PPP1CB", "PPP1CC", "PPP1R15A")] <- 2
+state <- rep(1, length(E(graph))); pathway <- get.edgelist(graph)
+state[pathway[,1] %in% c("SMAD6", "SMAD7", "BAMBI", "SMURF1", "SMURF2", "UCHL5",
+  "USP15", "UBB", "UBC", "PMEPA1", "PPP1CA", "PPP1CB", "PPP1CC", "PPP1R15A")] <- 2
 state[is.na(state)] <- 1
@@
 box()
-#adjacency matrix
-adj_mat <- make_adjmatrix_graph(graph)
-#relationship matrix
-dist_mat <- make_distance_graph(graph, absolute = FALSE)
-#sigma matrix directly from graph
-sigma_mat <- make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE)
-#show shortest paths of graph
-shortest_paths <- shortest.paths(graph)
-#generate expression data directly from graph
-expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0, comm = FALSE,
-                            dist = TRUE, absolute = FALSE)
-```
-```{r simulation_smad, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a biological pathway graph structure}.
-The graph structure (a) of a known biological pathway, the TGF-$\\beta$ receptor signaling activates SMADs (R-HSA-2173789), was used to derive a relationship matrix (b), $\\Sigma$ matrix (c) and correlation structure (d) from the relative distances between the nodes.
-These values are coloured blue to red from $-1$ to $1$.
-This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
-Here modules of genes with correlated expression can be clearly discerned.', warning=FALSE}
 # #plot relationship matrix
@@
 #simulated data
-expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0,
-comm = FALSE, dist =TRUE, absolute = FALSE)
+expr <- generate_expression(100, graph, state = state, cor = 0.8,
+  mean = 0,comm = FALSE, dist =TRUE, absolute = FALSE)
 #plot simulated correlations

### 2020-05-28 paper/paper.Rmd
hide code for subcaptions in figures

@@
 install.packages("graphsim")
+
 ```
-```{r, warning=FALSE, results='hide'}
+```{r, warning=FALSE, results='hide', message=FALSE}
 #load required packages (once per R instance)
-library("graphsim"); library("igraph"); library("gplots")
+library("graphsim")
+#load packages for examples
+library("igraph"); library("gplots")
+```
+
+```{r simple_graph_hide, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, eval=FALSE}
+#generate graph structure
+graph_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"), c("D", "E"),
+   c("D", "F"), c("F", "G"), c("F", "I"), c("H", "I"))
+graph <- graph.edgelist(graph_edges, directed = TRUE)
+
+#plot graph structure (Figure 1a)
+plot_directed(graph, state ="activating", layout = layout.kamada.kawai,
+  cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
+
+#generate parameters for inhibitions for each edge in E(graph)
+state <- c(1, 1, -1, 1, 1, 1, 1, -1, 1)
+#plot graph structure with inhibitions (Figure 1b)
+plot_directed(graph, state=state, layout = layout.kamada.kawai,
+  cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
 ```
@@
 Activating links are denoted by black arrows and inhibiting links by red edges.
-Inhibiting edges have been highlighted in red.'}
+Inhibiting edges have been highlighted in red.', echo=FALSE}
 #generate graph structure
@@
 
-#generate parameters for inhibitions
+#generate parameters for inhibitions for each edge in E(graph)
 state <- c(1, 1, -1, 1, 1, 1, 1, -1, 1)
@@
 
+```{r simulation_activating_hide, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, eval=FALSE, warning=FALSE}
+# activating graph
+state <- rep(1, length(E(graph)))
+plot_directed(graph, state=state, layout = layout.kamada.kawai,
+  cex.node=2, cex.arrow=4, arrow_clip = 0.2)
+
+#plot relationship matrix
+heatmap.2(make_distance_graph(graph, absolute = FALSE),
+  scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
+  colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+
+#plot sigma matrix
+heatmap.2(make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE),
+  scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
+  colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+
+#simulated data
+expr <- generate_expression(100, graph, cor = 0.8, mean = 0,
+  comm = FALSE, dist =TRUE, absolute = FALSE, state = state)
+#plot simulated correlations
+heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
+  colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+  mtext(text = "(d) Simulated correlation", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
+#plot simulated expression data
+heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+  colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+```
+
 ```{r simulation_activating, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}.
@@
 This $\\Sigma$ matrix has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
-Here genes with closer relationships in the pathway structure show a higher correlation between simulated values.', warning=FALSE}
+Here genes with closer relationships in the pathway structure show a higher correlation between simulated values.', warning=FALSE, echo=FALSE}
 # activating graph
@@
 
+```{r simulation_inhibiting_hide, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, eval=FALSE, warning=FALSE}
+#generate parameters for inhibitions
+state <- c(1, 1, -1, 1, 1, 1, 1, -1)
+plot_directed(graph, state=state, layout = layout.kamada.kawai,
+  cex.node=2, cex.arrow=4, arrow_clip = 0.2)
+
+#plot relationship matrix
+heatmap.2(make_distance_graph(graph, absolute = FALSE),
+  scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
+  colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+
+# #plot sigma matrix
+heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
+  scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
+  colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+
+#simulated data
+expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0,
+  comm = FALSE, dist =TRUE, absolute = FALSE)
+#plot simulated correlations
+heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
+  colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
+
+#plot simulated expression data
+heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+  colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+```
+
 ```{r simulation_inhibiting, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from graph structure with inhibitions}.
@@
 This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
-Here the inhibitory relationships between genes are reflected in negatively correlated simulated  values.', warning=FALSE}
+Here the inhibitory relationships between genes are reflected in negatively correlated simulated  values.', warning=FALSE, echo=FALSE}
 #generate parameters for inhibitions
@@
 
+```{r simulation_smad_hide, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, eval=FALSE, warning=FALSE}
+#import graph from data
+graph <- identity(TGFBeta_Smad_graph)
+#generate parameters for inhibitions
+state <- E(graph)$state
+
+plot_directed(graph, state = state, layout = layout.kamada.kawai,
+  border.node=scales::alpha("black", 0.75), fill.node="lightblue",
+  col.arrow = c(scales::alpha("navyblue", 0.25), scales::alpha("red", 0.25))[state], 
+  cex.node = 1.5, cex.label = 0.8, cex.arrow = 2)
+
+# #plot relationship matrix
+heatmap.2(make_distance_graph(graph, absolute = FALSE),
+  scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
+  colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+
+# #plot sigma matrix
+heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
+  scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
+  colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+
+#simulated data
+expr <- generate_expression(100, graph, state = state, cor = 0.8,
+  mean = 0,comm = FALSE, dist =TRUE, absolute = FALSE)
+#plot simulated correlations
+heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
+  colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+#plot simulated expression data
+heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
+```
+
 ```{r simulation_smad, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a biological pathway graph structure}.
@@
 This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
-Here modules of genes with correlated expression can be clearly discerned.', warning=FALSE}
+Here modules of genes with correlated expression can be clearly discerned.', warning=FALSE, echo=FALSE}
 #import graph from data

### 2020-05-28 paper/paper.Rmd
reduce size of figures

@@
 
-```{r simple_graph_hide, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, eval=FALSE}
+```{r simple_graph_hide, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.375\\linewidth', out.width='.375\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, eval=FALSE}
 #generate graph structure
@@
 
-```{r simple_graph, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulated graph structures}.
+```{r simple_graph, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.375\\linewidth', out.width='.375\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulated graph structures}.
 A constructed graph structure used as an example to demonstrate the simulation procedure in Figures 2 and 3.
@@
 
-```{r simulation_activating_hide, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, eval=FALSE, warning=FALSE}
+```{r simulation_activating_hide, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.375\\linewidth', out.width='.375\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, eval=FALSE, warning=FALSE}
 # activating graph
@@
 
-```{r simulation_activating, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}.
+```{r simulation_activating, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.375\\linewidth', out.width='.375\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}.
 An example of a graph structure (a) that has been used to derive a relationship matrix (b), $\\Sigma$  matrix (c) and correlation structure (d) from the relative distances between the nodes.
@@
 ```{r, include=FALSE, warning=FALSE}
-pdf("Plotsimulation_activating-5.pdf", width = 13.44, height = 6.72)
+pdf("Plotsimulation_activating-5.pdf", width = 12.14, height = 6.072)
 #plot simulated expression data
@@
 
-```{r simulation_inhibiting_hide, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, eval=FALSE, warning=FALSE}
+```{r simulation_inhibiting_hide, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.375\\linewidth', out.width='.375\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, eval=FALSE, warning=FALSE}
 #generate parameters for inhibitions
@@
 
-```{r simulation_inhibiting, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from graph structure with inhibitions}.
+```{r simulation_inhibiting, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.375\\linewidth', out.width='.375\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from graph structure with inhibitions}.
 An example of a graph structure (a), that has been used to derive a relationship matrix (b), $\\Sigma$ matrix (c), and correlation structure (d), from the relative distances between the nodes.
@@
 ```{r, include=FALSE, warning=FALSE}
-pdf("Plotsimulation_inhibiting-5.pdf", width = 13.44, height = 6.72)
+pdf("Plotsimulation_inhibiting-5.pdf", width = 12.14, height = 6.072)
 #plot simulated expression data
@@
 
-```{r simulation_smad_hide, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, eval=FALSE, warning=FALSE}
+```{r simulation_smad_hide, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.375\\linewidth', out.width='.375\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, eval=FALSE, warning=FALSE}
 #import graph from data
@@
 
-```{r simulation_smad, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a biological pathway graph structure}.
+```{r simulation_smad, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.375\\linewidth', out.width='.375\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a biological pathway graph structure}.
 The graph structure (a) of a known biological pathway, the TGF-$\\beta$ receptor signaling activates SMADs (R-HSA-2173789), was used to derive a relationship matrix (b), $\\Sigma$ matrix (c) and correlation structure (d) from the relative distances between the nodes.
@@
 ```{r, include=FALSE, warning=FALSE}
-pdf("Plotsimulation_smad-5.pdf", width = 13.44, height = 6.72)
+pdf("Plotsimulation_smad-5.pdf", width = 12.14, height = 6.072)
 #plot simulated expression data

### 2020-05-28 paper/images.sh

@@ -1,1 +1,1 @@
-sed -i '/Plot.*-5/s/\(.*\)width=[.]415/\1width=.830/g' paper.md paper.tex
+sed -i '/Plot.*-5/s/\(.*\)width=[.]375/\1width=.750/g' paper.md paper.tex

### 2020-05-28 paper/paper.Rmd
correct code style in paper examples

@@
 #plot sigma matrix
-heatmap.2(make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE),
+heatmap.2(make_sigma_mat_dist_graph(graph, cor = 0.8, absolute = FALSE),
   scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
@@
 #plot sigma matrix
-heatmap.2(make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE),
+heatmap.2(make_sigma_mat_dist_graph(graph, cor = 0.8, absolute = FALSE),
 scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
@@
 
-# #plot sigma matrix
-heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
+#plot sigma matrix
+heatmap.2(make_sigma_mat_dist_graph(graph, state = state, cor = 0.8, absolute = FALSE),
   scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
@@
   colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-
 #plot simulated expression data
@@
 mtext(text = "(b) Relationship matrix", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
-# #plot sigma matrix
-heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
+#plot sigma matrix
+heatmap.2(make_sigma_mat_dist_graph(graph, state = state, cor = 0.8, absolute = FALSE),
 scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
@@
 box()
-# #plot relationship matrix
+#plot relationship matrix
 heatmap.2(make_distance_graph(graph, absolute = FALSE),

### 2020-05-28 paper/paper.Rmd
update code examples in paper

@@
 #load packages for examples
-library("igraph"); library("gplots")
+library("igraph"); library("gplots"); library("scales")
 ```
@@
 #plot sigma matrix
-heatmap.2(make_sigma_mat_dist_graph(graph, state = state, cor = 0.8, absolute = FALSE),
+heatmap.2(make_sigma_mat_dist_graph(graph, state, cor = 0.8, absolute = FALSE),
   scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
@@
 #simulated data
-expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0,
+expr <- generate_expression(100, graph, state, cor = 0.8, mean = 0,
   comm = FALSE, dist =TRUE, absolute = FALSE)
@@
 #plot sigma matrix
-heatmap.2(make_sigma_mat_dist_graph(graph, state = state, cor = 0.8, absolute = FALSE),
+heatmap.2(make_sigma_mat_dist_graph(graph, state, cor = 0.8, absolute = FALSE),
 scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
@@
 #simulated data
-expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0,
+expr <- generate_expression(100, graph, state, cor = 0.8, mean = 0,
 comm = FALSE, dist =TRUE, absolute = FALSE)
@@
 plot_directed(graph, state = state, layout = layout.kamada.kawai,
-  border.node=scales::alpha("black", 0.75), fill.node="lightblue",
-  col.arrow = c(scales::alpha("navyblue", 0.25), scales::alpha("red", 0.25))[state], 
+  border.node=alpha("black", 0.75), fill.node="lightblue",
+  col.arrow = c(alpha("navyblue", 0.25), alpha("red", 0.25))[state], 
   cex.node = 1.5, cex.label = 0.8, cex.arrow = 2)
@@
 #simulated data
-expr <- generate_expression(100, graph, state = state, cor = 0.8,
+expr <- generate_expression(100, graph, state, cor = 0.8,
   mean = 0,comm = FALSE, dist =TRUE, absolute = FALSE)
@@
 plot_directed(graph, state = state, layout = layout.kamada.kawai,
-  border.node=scales::alpha("black", 0.75), fill.node="lightblue",
-  col.arrow = c(scales::alpha("navyblue", 0.25), scales::alpha("red", 0.25))[state], 
+  border.node=alpha("black", 0.75), fill.node="lightblue",
+  col.arrow = c(alpha("navyblue", 0.25), alpha("red", 0.25))[state], 
   cex.node = 1.5, cex.label = 0.8, cex.arrow = 2, 

### 2020-05-28 paper/paper.Rmd
correct code examples for paper margins

@@
 #plot simulated correlations
-heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
+heatmap.2(cor(t(expr)), scale = "none", trace = "none",
+          col = colorpanel(50, "white", "red"),
   colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
-  mtext(text = "(d) Simulated correlation", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
 #plot simulated expression data
@@
 #plot simulated correlations
-heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
+heatmap.2(cor(t(expr)), scale = "none", trace = "none",
+          col = colorpanel(50, "blue", "white", "red"),
   colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
@@
 #plot simulated correlations
-heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
+heatmap.2(cor(t(expr)), scale = "none", trace = "none", 
+          col = colorpanel(50, "blue", "white", "red"),
   colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")

### 2020-06-01 paper/paper.Rmd
restore data for TGFBeta-Smad pathway

@@
 install.packages("graphsim")
-
 ```

### 2020-06-02 R/state.R
update handling of duplicate edges in state parameter

@@ -157,8 +157,29 @@ make_state_matrix <- function(graph, state = NULL){
           edges[,1] == names(paths[[ii]])[jj-1] & edges[,2] == names(paths[[ii]])[jj] |
           edges[,2] == names(paths[[ii]])[jj-1] & edges[,1] == names(paths[[ii]])[jj]
           )
+          #check for multiple edges
+          if(length(kk) > 1){
+            #check for same edges
+            if(all(apply(edges[kk,], 2, function(x) all(x == x[1])))){
+              #check for same state
+              if(all(state[kk] == state[kk][1])){
+                state[kk] <- state[kk][1]
+                kk <- kk[1]
+              } else {
+                Mode <- function(x) unique(x)[which.max(tabulate(match(x, unique(x))))][1]
+                state[kk] <- Mode(state[kk])
+                kk <- kk[1]
+              }
+            }
+          }
           #state for edges
-          state_path[jj] <- state[kk]
+          if(length(state_path[jj]) == length(state[kk])){
+            state_path[jj] <- state[kk]
+          } else {
+            print(ii)
+            print(jj)
+            stop("State not compatible with graph paths")
+          }
         }
         #cumulative product changes state along a path (*-1)
         state_path <- cumprod(state_path)

### 2020-06-02 R/state.R
warn for state conflicts

@@ -166,6 +166,7 @@ make_state_matrix <- function(graph, state = NULL){
                 state[kk] <- state[kk][1]
                 kk <- kk[1]
               } else {
+                warning(paste("Conflicting state for redundant edges", edges[kk[1],1], edges[kk[1],2], "with state", state[kk]))
                 Mode <- function(x) unique(x)[which.max(tabulate(match(x, unique(x))))][1]
                 state[kk] <- Mode(state[kk])
                 kk <- kk[1]

### 2020-06-02 R/state.R
check state of edges in opposite directions

@@ -160,6 +160,11 @@ make_state_matrix <- function(graph, state = NULL){
           #check for multiple edges
           if(length(kk) > 1){
             #check for same edges
+            if(!all(apply(edges[kk,], 2, function(x) all(x == x[1])))){
+              #reverse edge if no match
+              edges[kk,][edges[kk[1],1] != edges[kk,1],] <- edges[kk,][edges[kk[1],1] != edges[kk,1],2:1]
+            }
+              #check is still no match
             if(all(apply(edges[kk,], 2, function(x) all(x == x[1])))){
               #check for same state
               if(all(state[kk] == state[kk][1])){

### 2020-06-02 R/generate.R
correct passing state as edge attributes

@@ -44,6 +44,15 @@ generate_expression <- function(n, graph, state = NULL, cor = 0.8, mean = 0, com
       state <- rep(1, length(E(graph)))
     }
   }
+  if(is.null(get.edge.attribute(graph, "state"))){
+    E(graph)$state <-  state
+  }
+  # this could also be done with igraph::simplify(graph, remove.multiple = TRUE, remove.loops = TRUE, edge.attr.comb = igraph_opt("edge.attr.comb"))
+  ## to do: migrate to this
+  graph <- as.undirected(graph, mode = "collapse", edge.attr.comb = function(x) ifelse(any(x %in% list(-1, 2, "inhibiting", "inhibition")), -1, 1))
+  # remove duplicate edges (and corresponding states)
+  graph <- as.directed(graph, mode = "arbitrary")
+  state <- get.edge.attribute(graph, "state")
   if(!is.integer(n)){
     if(is.numeric(n)){
       if(floor(n) == n){
@@ -68,7 +77,8 @@ generate_expression <- function(n, graph, state = NULL, cor = 0.8, mean = 0, com
   ## migrate state to calling sigma ##
   if(!(is.null(state))){
     if(all(sig >= 0)){
-      sig <- state * sig
+      state_mat <- make_state_matrix(graph, state)
+      sig <- sig * sign(state_mat)
     }
   }
   if(is.symmetric.matrix(sig) == FALSE) {

### 2020-06-02 R/sigma.R

@@ -149,6 +149,16 @@ make_sigma_mat_dist_graph <- function(graph, state = NULL, cor = 0.8, absolute =
     }
   }
   mat <- make_distance_graph(graph, absolute = absolute)
-  sig <- make_sigma_mat_dist_adjmat(mat, state, cor)
+  if(!(all(diag(mat) == 1))) stop("distance matrix must have diagonal of zero")
+  if(!(max(mat[mat != 1]) > 0) || !(max(mat[mat!=1]) <= 1)) stop("distance matrix expected, not adjacency matrix")
+  sig <- mat/max(mat[mat != 1]) * cor
+  sig <- ifelse(sig > 0, sig, 0)
+  diag(sig) <- 1
+  #derive states directly from graph if available
+  state_mat <- make_state_matrix(graph, state)
+  sig <- sig * sign(state_mat)
+  rownames(sig) <- rownames(mat)
+  colnames(sig) <- colnames(mat)
+  return(sig)
   return(sig)
 }

### 2020-06-02 R/state.R

@@ -38,8 +38,12 @@ make_state_matrix <- function(graph, state = NULL){
       state <- "activating"
     }
   }
-  graph <- as.undirected(graph, mode = "collapse")
-  E(graph)$state <-  state
+  if(is.null(get.edge.attribute(graph, "state"))){
+    E(graph)$state <-  state
+  }
+  # this could also be done with igraph::simplify(graph, remove.multiple = TRUE, remove.loops = TRUE, edge.attr.comb = igraph_opt("edge.attr.comb"))
+  ## to do: migrate to this
+  graph <- as.undirected(graph, mode = "collapse", edge.attr.comb = function(x) ifelse(any(x %in% list(-1, 2, "inhibiting", "inhibition")), -1, 1))
   # remove duplicate edges (and corresponding states)
   graph <- as.directed(graph, mode = "arbitrary")
   state <- get.edge.attribute(graph, "state")

### 2020-06-02 paper/paper.Rmd
update paper figures

@@
    index: 2
-date: 30 April 2020
+date: "`r  format(Sys.time(), '%d %B %Y')`"
 bibliography: paper.bib

### 2020-06-04 paper/paper.Rmd
add citations to introduction

@@
 network modules being predicted [@Markowetz2007; @Hawe2019].
+
+GeneNetWeaver @Schaffter2011 - benchmarking of gene regulatory network inference methods
+Splatter @Zappia2017 - simulated single-cell RNA-Seq data (gamma-poisson)
+dyngen @Cannoodt2020 - single-cell multimodal -  Gillespie’s stochastic simulation at single-molecule level
+SERGIO @Dibaeinia2019 - single-cell gene expression (noise with poisson)
 

### 2020-06-10 paper/paper.Rmd
update intro and discussion

--- 2020-06-02 paper/paper.Rmd
@@
 novel biological insights [@Arner2015; @Komatsu2013]. However, inference
-of regulatory networks has thus far relied on experimental validation or
-resampling-based approaches to estimate the likelihood of specific
-network modules being predicted [@Markowetz2007; @Hawe2019].
+of regulatory networks has thus far primarily relied on experimental
+validation or resampling-based approaches to estimate the likelihood
+of specific network modules being predicted [@Markowetz2007; @Hawe2019].
+
+Simulated datasets with pathway structure are of particular interest for
+benchmarking regulatory network inference techniques and methods
+being developed for genomics data with complex biological 
+interactions [@Schaffter2011, @Saelens2019]. Dynamical models using differential equations have been employed,
+such as by GeneNetWeaver [@Schaffter2011], to generate simulated datasets
+specifically for benchmarking gene regulatory network inference techniques.
+There is renewed interest in modelling biological pathways and simulating
+data for benchmarking due to the emergence of single-cell genomics
+technologies and the plethora of bioinformatics techniques developed
+to use this data [@Zappia2017, @Saelens2019].
+
+Splatter [@Zappia2017] was developed to model single-cell data and 
+SERGIO [@Dibaeinia2019] and dyngen [@Cannoodt2020] build on this to
+include gene regulatory networks and multimodality respectively.
+
+These methods have been designed based on known or modelled reaction
+states and added stochasticity to determinisitic relationships.
+However, it is computationally-intensive to model these reactions
+at scale over iterations in benchmarking. In some cases, it is only
+necessary to model the statistical variability of RNA-Seq data,
+
+GeneNetWeaver @Schaffter2011 - benchmarking of gene regulatory network inference methods
+Splatter @Zappia2017 - simulated single-cell RNA-Seq data (gamma-poisson)
+
+dyngen @Cannoodt2020 - single-cell multimodal -  Gillespie’s stochastic simulation at single-molecule level
+
+SERGIO @Dibaeinia2019 - single-cell gene expression (noise with poisson)
 
@@
 Methodology and software {#sec:methods}
-========================
+========================d
 
@@
 present a statistical framework to simulate gene expression from
-biological pathways, and provide the \texttt{graphsim} package in R to generate these
-simulated datasets. This approach is versatile and can be fine-tuned for
-modelling existing biological pathways or for testing whether
-constructed pathways can be detected by other means. In particular,
-methods to infer biological pathways and gene regulatory networks from
-gene expression data can be tested on simulated datasets using this
-framework. The package also enables simulation of complex gene
+biological pathways, and provide the \texttt{graphsim} package in R 
+to generate these simulated datasets. This approach is versatile and
+can be fine-tuned for modelling existing biological pathways or for
+testing whether constructed pathways can be detected by other means.
+In particular, methods to infer biological pathways and gene regulatory
+networks from gene expression data can be tested on simulated datasets
+using this framework. The package also enables simulation of complex gene
 expression datasets to test how these pathways impact on statistical
@@
 statistical methods being developed for gene expression data analysis.
+This approach is intended to be applied to bulk gene expression data
+but could in principle be adapted to modelling single-cell or
+different modalities such as epigenomics data.
 

### 2020-06-10 paper/paper.Rmd
Give details on state of the in the manuscript

--- 2020-06-02 paper/paper.Rmd
@@
 Introduction: inference and modelling of biological networks {#sec:intro}
-============================================================
+===============================================================================
 
@@
 novel biological insights [@Arner2015; @Komatsu2013]. However, inference
-of regulatory networks has thus far relied on experimental validation or
-resampling-based approaches to estimate the likelihood of specific
-network modules being predicted [@Markowetz2007; @Hawe2019].
+of regulatory networks has thus far primarily relied on experimental
+validation or resampling-based approaches to estimate the likelihood
+of specific network modules being predicted [@Markowetz2007; @Hawe2019].
+
+Simulated datasets with pathway structure are of particular interest for
+benchmarking regulatory network inference techniques and methods
+being developed for genomics data with complex biological 
+interactions [@Schaffter2011; @Saelens2019]. Dynamical models using
+differential equations have been employed, such as by GeneNetWeaver
+[@Schaffter2011], to generate simulated datasets
+specifically for benchmarking gene regulatory network inference techniques.
+There is renewed interest in modelling biological pathways and simulating
+data for benchmarking due to the emergence of single-cell genomics
+technologies and the plethora of bioinformatics techniques developed
+to use this data [@Zappia2017; @Saelens2019]. Packages such as Splatter
+[@Zappia2017], which uses the gamma-poisson distribution,
+have been developed to model single-cell data.
+SERGIO [@Dibaeinia2019] and dyngen [@Cannoodt2020] build on
+this to include gene regulatory networks and multimodality
+respectively. These methods have been designed based on known
+or modelled reaction states and added stochasticity to determinisitic
+relationships. However, it is computationally-intensive to model
+these reactions at scale over iterations in benchmarking.
+In some cases, it is only necessary to model the statistical
+variability of RNA-Seq data to evaluate methods with these
+correlation structures.
 
@@
 Methodology and software {#sec:methods}
-========================
+===============================================================================
 
@@
 Illustrations {#sec:illustrations}
-=============
+===============================================================================
 
 Generating a Graph Structure {#sec:plot_graph}
-----------------------------
+-------------------------------------------------------------------------------
 
@@
 Summary and discussion {#sec:summary}
-======================
+===============================================================================
 
@@
 statistical methods being developed for gene expression data analysis.
+This approach is intended to be applied to bulk gene expression data
+but could in principle be adapted to modelling single-cell or
+different modalities such as epigenomics data.
 
 Computational details {#computational-details .unnumbered .unnumbered}
-=====================
+===============================================================================
 
@@
 Acknowledgements {#acknowledgements .unnumbered .unnumbered}
-================
+===============================================================================
 
@@
 Author Contributions  {#contributions .unnumbered .unnumbered}
-================
+===============================================================================
 

### 2020-06-10 paper/paper.Rmd
paper formatting
update documentation with examples of biological networks #6

@@
 technologies and the plethora of bioinformatics techniques developed
-to use this data [@Zappia2017; @Saelens2019]. Packages such as Splatter
+to use this data [@Zappia2017; @Saelens2019]. Packages such as 'splatter'
 [@Zappia2017], which uses the gamma-poisson distribution,
@@
 #generate parameters for inhibitions for each edge in E(graph)
-state <- c(1, 1, -1, 1, 1, 1, 1, -1, 1)
+state <- c(1, 1, -1, 1, 1, 1, 1, -1)
 #plot graph structure with inhibitions (Figure 1b)

### 2020-06-10 R/adjmat.R

@@ -13,10 +13,31 @@
 ##' @import igraph
 ##' @examples 
 ##' 
+##' # construct a synthetic graph module
 ##' library("igraph")
 ##' graph_test_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
 ##' graph_test <- graph.edgelist(graph_test_edges, directed = TRUE)
+##' 
+##' # compute adjacency matrix for toy example
 ##' adjacency_matrix <- make_adjmatrix_graph(graph_test)
+##' adjacency_matrix
+##' 
+##' # construct a synthetic graph network
+##' graph_structure_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"), c("D", "E"),
+##'                                c("D", "F"), c("F", "G"), c("F", "I"), c("H", "I"))
+##' graph_structure <- graph.edgelist(graph_structure_edges, directed = TRUE)
+##' # compute adjacency matrix for toy network
+##' graph_structure_adjacency_matrix <- make_adjmatrix_graph(graph_structure)
+##' graph_structure_adjacency_matrix
+##' 
+##' # import graph from package for reactome pathway
+##' # TGF-β receptor signaling activates SMADs (R-HSA-2173789)
+##' TGFBeta_Smad_graph <- identity(TGFBeta_Smad_graph)
+##' 
+##' # compute adjacency matrix for TGF-β receptor signaling activates SMADs
+##' TGFBeta_Smad_adjacency_matrix <- make_adjmatrix_graph(TGFBeta_Smad_graph)
+##' dim(TGFBeta_Smad_adjacency_matrix)
+##' TGFBeta_Smad_adjacency_matrix[1:12, 1:12]
 ##' 
 ##' @return An adjacency matrix compatible with generating an expression matrix
 ##' 

### 2020-06-10 R/commonlink.R

@@ -12,11 +12,39 @@
 ##' @import igraph
 ##' @examples 
 ##' 
+##' # construct a synthetic graph module
 ##' library("igraph")
 ##' graph_test_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
 ##' graph_test <- graph.edgelist(graph_test_edges, directed = TRUE)
+##' 
+##' # compute adjacency matrix for toy example
 ##' adjacency_matrix <- make_adjmatrix_graph(graph_test)
+##' # compute nodes with shared edges to a 3rd node
 ##' common_link_matrix <- make_commonlink_adjmat(adjacency_matrix)
+##' common_link_matrix
+##' 
+##' # construct a synthetic graph network
+##' graph_structure_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"), c("D", "E"),
+##'                                c("D", "F"), c("F", "G"), c("F", "I"), c("H", "I"))
+##' graph_structure <- graph.edgelist(graph_structure_edges, directed = TRUE)
+##' # compute adjacency matrix for toy network
+##' graph_structure_adjacency_matrix <- make_adjmatrix_graph(graph_structure)
+##' # compute nodes with shared edges to a 3rd node
+##' graph_structure_common_link_matrix <- make_commonlink_adjmat(graph_structure_adjacency_matrix)
+##' graph_structure_common_link_matrix
+##' 
+##' # import graph from package for reactome pathway
+##' # TGF-β receptor signaling activates SMADs (R-HSA-2173789)
+##' TGFBeta_Smad_graph <- identity(TGFBeta_Smad_graph)
+##' # compute nodes with shared edges to a 3rd node 
+##' TGFBeta_Smad_adjacency_matrix <- make_adjmatrix_graph(TGFBeta_Smad_graph)
+##' TGFBeta_Smad_common_link_matrix <- make_commonlink_adjmat(TGFBeta_Smad_adjacency_matrix)
+##' # we show summary statistics as the graph is large
+##' dim(TGFBeta_Smad_common_link_matrix)
+##' TGFBeta_Smad_common_link_matrix[1:12, 1:12]
+##' # visualise matrix
+##' library("gplots")
+##' heatmap.2(TGFBeta_Smad_common_link_matrix, scale = "none", trace = "none", col = colorpanel(50, "white", "red"))
 ##' 
 ##' @return An integer matrix of number of links shared between nodes
 ##' 

### 2020-06-10 R/distance.R

@@ -14,16 +14,61 @@
 ##' @import igraph
 ##' @examples 
 ##' 
+##' # construct a synthetic graph module
 ##' library("igraph")
 ##' graph_test_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
 ##' graph_test <- graph.edgelist(graph_test_edges, directed = TRUE)
+##' 
+##' # compute adjacency matrix for toy example
 ##' adjacency_matrix <- make_adjmatrix_graph(graph_test)
-##' distance_matrix <- make_distance_adjmat(adjacency_matrix)
+##' # compute nodes with relationships between nodes (geometrically decreasing by default)
+##' distance_matrix_geom <- make_distance_adjmat(adjacency_matrix)
+##' distance_matrix_geom
 ##' 
+##' # compute nodes with relationships between nodes (arithmetically decreasing)
+##' distance_matrix_abs <- make_distance_adjmat(adjacency_matrix, absolute = TRUE)
+##' distance_matrix_abs
+##' 
+##' # compute Laplacian matrix
 ##' laplacian_matrix <- make_laplacian_graph(graph_test)
-##' distance_matrix <- make_distance_adjmat(laplacian_matrix)
+##' # compute distances from Laplacian
+##' distance_matrix <- make_distance_laplacian(laplacian_matrix)
 ##' 
-##' @return A numeric matrix of values in the range [0, 1] where lower values are closer
+##' # construct a synthetic graph network
+##' graph_structure_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"), c("D", "E"),
+##'                                c("D", "F"), c("F", "G"), c("F", "I"), c("H", "I"))
+##' graph_structure <- graph.edgelist(graph_structure_edges, directed = TRUE)
+##' # compute adjacency matrix for toy network
+##' graph_structure_adjacency_matrix <- make_adjmatrix_graph(graph_structure)
+##' # compute nodes with relationships between nodes (geometrically decreasing by default)
+##' graph_structure_distance_matrix_geom <- make_distance_adjmat(graph_structure_adjacency_matrix)
+##' graph_structure_distance_matrix_geom
+##' # visualise matrix
+##' library("gplots")
+##' heatmap.2(graph_structure_distance_matrix_geom, scale = "none", trace = "none", col = colorpanel(50, "white", "red"))
[…]

### 2020-06-10 R/generate.R

@@ -23,11 +23,81 @@
 ##' @importFrom gplots heatmap.2
 ##' @examples
 ##' 
+##' # construct a synthetic graph module
+##' library("igraph")
+##' graph_test_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
+##' graph_test <- graph.edgelist(graph_test_edges, directed = TRUE)
+##' 
+##' # compute a simulated dataset for toy example
+##' # n = 100 samples
+##' # cor = 0.8 max correlation between samples
+##' # absolute = FALSE (geometric distance by default)
+##' test_data <- generate_expression(100, graph_test, cor = 0.8)
+##' ##' # visualise matrix
+##' library("gplots")
+##' # expression data
+##' heatmap.2(test_data, scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"))
+##' # correlations
+##' heatmap.2(cor(t(test_data)), scale = "none", trace = "none", col = colorpanel(50, "white", "red"))
+##' # expected correlations (Σ)
+##' sigma_matrix <- make_sigma_mat_graph(graph_test, cor = 0.8)
+##' heatmap.2(make_sigma_mat_graph(graph_test, cor = 0.8), scale = "none", trace = "none", col = colorpanel(50, "white", "red"))
+##' 
+##' # compute a simulated dataset for toy example
+##' # n = 100 samples
+##' # cor = 0.8 max correlation between samples
+##' # absolute = TRUE (arithmetic distance)
+##' test_data <- generate_expression(100, graph_test, cor = 0.8, absolute = TRUE)
+##' ##' # visualise matrix
+##' library("gplots")
+##' # expression data
+##' heatmap.2(test_data, scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"))
+##' # correlations
+##' heatmap.2(cor(t(test_data)), scale = "none", trace = "none", col = colorpanel(50, "white", "red"))
+##' # expected correlations (Σ)
+##' sigma_matrix <- make_sigma_mat_graph(graph_test, cor = 0.8)
+##' heatmap.2(make_sigma_mat_graph(graph_test, cor = 0.8), scale = "none", trace = "none", col = colorpanel(50, "white", "red"))
+##' 
+##' # construct a synthetic graph network
[…]

### 2020-06-10 R/laplacian.R

@@ -16,10 +16,38 @@
 ##' @importClassesFrom Matrix dgCMatrix
 ##' @examples 
 ##' 
-##' library("igraph")
+##' # construct a synthetic graph module
+##' library("igraph") 
 ##' graph_test_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
 ##' graph_test <- graph.edgelist(graph_test_edges, directed = TRUE)
+##' # compute Laplacian matrix for toy example
 ##' laplacian_matrix <- make_laplacian_graph(graph_test)
+##' laplacian_matrix
+##' 
+##' # compute Laplacian matrix from adjacency matrix
+##' adjacency_matrix <- make_adjmatrix_graph(graph_test)
+##' laplacian_matrix <- make_laplacian_adjmat(adjacency_matrix)
+##' laplacian_matrix
+##' 
+##' # construct a synthetic graph network
+##' graph_structure_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"), c("D", "E"),
+##'                                c("D", "F"), c("F", "G"), c("F", "I"), c("H", "I"))
+##' graph_structure <- graph.edgelist(graph_structure_edges, directed = TRUE)
+##' # compute Laplacian matrix for toy network
+##' graph_structure_laplacian_matrix <- make_laplacian_graph(graph_structure)
+##' graph_structure_laplacian_matrix
+##'  
+##' # import graph from package for reactome pathway
+##' # TGF-β receptor signaling activates SMADs (R-HSA-2173789)
+##' TGFBeta_Smad_graph <- identity(TGFBeta_Smad_graph)
+##' 
+##' # compute Laplacian matrix for TGF-β receptor signaling activates SMADs
+##' TGFBeta_Smad_laplacian_matrix <- make_laplacian_graph(TGFBeta_Smad_graph)
+##' dim(TGFBeta_Smad_laplacian_matrix)
+##' TGFBeta_Smad_laplacian_matrix[1:12, 1:12]
+##' # visualise matrix
+##' library("gplots")
+##' heatmap.2(TGFBeta_Smad_laplacian_matrix, scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"))
 ##' 
 ##' @return An Laplacian matrix compatible with generating an expression matrix
 ##' 

### 2020-06-10 R/plot_directed.R

@@ -55,30 +55,31 @@
 ##'
 ##' #generate example graphs
 ##' library("igraph")
-##' graph_test4_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"), c("D", "E"),
+##' graph_structure_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"), c("D", "E"),
 ##'                            c("D", "F"), c("F", "G"), c("F", "I"), c("H", "I"))
-##' graph_test4 <- graph.edgelist(graph_test4_edges, directed = TRUE)
+##' graph_structure <- graph.edgelist(graph_structure_edges, directed = TRUE)
 ##'
 ##' #plots with igraph defaults
-##' plot(graph_test4, layout = layout.fruchterman.reingold)
-##' plot(graph_test4, layout = layout.kamada.kawai)
+##' plot(graph_structure, layout = layout.fruchterman.reingold)
+##' plot(graph_structure, layout = layout.kamada.kawai)
 ##'
 ##' #plots with scalar states
-##' plot_directed(graph_test4, state="activating")
-##' plot_directed(graph_test4, state="inhibiting")
+##' plot_directed(graph_structure, state="activating")
+##' plot_directed(graph_structure, state="inhibiting")
 ##'
 ##' #plots with vector states
-##' plot_directed(graph_test4, state=c(1, 1, 1, 1, -1, 1, 1, 1))
-##' plot_directed(graph_test4, state=c(1, 1, -1, 1, -1, 1, -1, 1))
+##' plot_directed(graph_structure, state = c(1, 1, 1, 1, -1, 1, 1, 1))
+##' plot_directed(graph_structure, state = c(1, 1, -1, 1, -1, 1, -1, 1))
+##' plot_directed(graph_structure, state = c(1, 1, -1, 1, 1, 1, 1, -1)
 ##' 
 ##' #plots states with graph attributes
-##' E(graph_test4)$state <- 1
-##' plot_directed(graph_test4)
-##' E(graph_test4)$state <- c(1, 1, -1, 1, -1, 1, -1, 1)
-##' plot_directed(graph_test4)
+##' E(graph_structure)$state <- 1
+##' plot_directed(graph_structure)
+##' E(graph_structure)$state <- c(1, 1, -1, 1, -1, 1, -1, 1)
+##' plot_directed(graph_structure)
 ##'
 ##' #plot layout customised
[…]

### 2020-06-10 R/sigma.R

@@ -1,9 +1,9 @@
 ##' @name make_sigma
 ##' @rdname make_sigma
 ##'
-##' @title Generate Sigma Matrix
+##' @title Generate Sigma (Σ) Matrix
 ##'
-##' @description Compute the Sigma matrix from an \code{\link[igraph]{igraph}} structure 
+##' @description Compute the Sigma (Σ) matrix from an \code{\link[igraph]{igraph}} structure 
 ##' or pre-computed matrix. These are compatible with \code{\link[mvtnorm]{rmvnorm}} and
 ##' \code{\link[graphsim]{generate_expression}}.
 ##'
@@ -22,11 +22,106 @@
 ##' @import igraph
 ##' @examples 
 ##' 
+##' # construct a synthetic graph module
 ##' library("igraph")
 ##' graph_test_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
 ##' graph_test <- graph.edgelist(graph_test_edges, directed = TRUE)
+##' # compute sigma (Σ) matrix for toy example
+##' sigma_matrix <- make_sigma_mat_graph(graph_test, cor = 0.8)
+##' sigma_matrix
+##' 
+##' # compute sigma (Σ) matrix  from adjacency matrix for toy example
 ##' adjacency_matrix <- make_adjmatrix_graph(graph_test)
 ##' sigma_matrix <- make_sigma_mat_adjmat(adjacency_matrix, cor = 0.8)
+##' sigma_matrix
+##' 
+##' # compute sigma (Σ) matrix from shared edges for toy example
+##' common_link_matrix <- make_commonlink_graph(graph_test)
+##' sigma_matrix <- make_sigma_mat_comm(common_link_matrix, cor = 0.8)
+##' sigma_matrix
+##' 
+##' # compute sigma (Σ) matrix from Laplacian for toy example
+##' laplacian_matrix <- make_laplacian_graph(graph_test)
+##' sigma_matrix <- make_sigma_mat_laplacian(laplacian_matrix, cor = 0.8)
+##' sigma_matrix
+##' 
+##' # compute sigma (Σ) matrix from distance matrix for toy example
+##' distance_matrix <- make_distance_graph(graph_test, absolute = FALSE)
+##' sigma_matrix <- make_sigma_mat_dist_adjmat(distance_matrix, cor = 0.8)
+##' sigma_matrix
+##' 
+##' # compute sigma (Σ) matrix from geometric distance directly from toy example graph
[…]

### 2020-06-10 R/state.R

@@ -5,7 +5,9 @@
 ##' 
 ##' @description Functions to compute the matrix of states (1 for activating and -1 for inhibiting) 
 ##' for link signed correlations, from a vector of edge states to a signed adjacency matrix for use
-##' in \code{\link[graphsim]{generate_expression}}.
+##' in \code{\link[graphsim]{generate_expression}}. This resolves edge states to determine the sign
+##' of all correlations between nodes in a network. These are computed interally for sigma matrices
+##' as required.
 ##' @param graph An \code{\link[igraph]{igraph}} object. May be directed or weighted as long as
 ##' a shortest path can be computed.
 ##' @param state numeric vector. Vector of length E(graph). Sign used to calculate state matrix,
@@ -21,10 +23,62 @@
 ##' @import igraph
 ##' @examples 
 ##' 
+##' # construct a synthetic graph module
+##' library("igraph")
+##' graph_test_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
+##' graph_test <- graph.edgelist(graph_test_edges, directed = TRUE)
+##' 
+##'  # compute state matrix for toy example
+##' state_matrix <- make_state_matrix(graph_test)
+##' 
+##' # construct a synthetic graph network
+##' graph_structure_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"), c("D", "E"),
+##'                                c("D", "F"), c("F", "G"), c("F", "I"), c("H", "I"))
+##' graph_structure <- graph.edgelist(graph_structure_edges, directed = TRUE)
+##' 
+##' # compute state matrix for toy network
+##' graph_structure_state_matrix <- make_state_matrix(graph_structure)
+##' graph_structure_state_matrix
+##' 
+##' # compute state matrix for toy network with inhibitions
+##' edge_state <- c(1, 1, -1, 1, 1, 1, 1, -1)
+##' # edge states are a variable
+##' graph_structure_state_matrix <- make_state_matrix(graph_structure, state = edge_state)
+##' graph_structure_state_matrix
+##' 
+##' # compute state matrix for toy network with inhibitions
+##' E(graph_structure)$state <- c(1, 1, -1, 1, 1, 1, 1, -1)
+##' # edge states are a graph attribute
[…]

### 2020-06-18 R/adjmat.R
format documentation

@@ -4,7 +4,8 @@
 ##'
 ##' @title Generate Adjacency Matrix
 ##' 
-##' @description Compute the adjacency matrix of a (directed) \code{\link[igraph]{igraph}} structure, preserving node/column/row names (and direction).
+##' @description Compute the adjacency matrix of a (directed) \code{\link[igraph]{igraph}}
+##' structure, preserving node/column/row names (and direction).
 ##'
 ##' @param graph An \code{\link[igraph]{igraph}} object. May be directed or weighted.
 ##' @param directed logical. Whether directed information is passed to the adjacency matrix.

### 2020-06-18 R/commonlink.R

@@ -3,7 +3,8 @@
 ##'
 ##' @title Generate Common Link Matrix
 ##'
-##' @description Compute the common link matrix of a (directed) \code{\link[igraph]{igraph}} structure, preserving node / column / row names (and direction).
+##' @description Compute the common link matrix of a (directed) \code{\link[igraph]{igraph}}
+##' structure, preserving node / column / row names (and direction).
 ##'
 ##' @param adj_mat precomputed adjacency matrix.
 ##' @param graph An \code{\link[igraph]{igraph}} object. May be directed or weighted.
@@ -44,7 +45,8 @@
 ##' TGFBeta_Smad_common_link_matrix[1:12, 1:12]
 ##' # visualise matrix
 ##' library("gplots")
-##' heatmap.2(TGFBeta_Smad_common_link_matrix, scale = "none", trace = "none", col = colorpanel(50, "white", "red"))
+##' heatmap.2(TGFBeta_Smad_common_link_matrix, scale = "none", trace = "none",
+##'           col = colorpanel(50, "white", "red"))
 ##' 
 ##' @return An integer matrix of number of links shared between nodes
 ##' 

### 2020-06-18 R/distance.R

@@ -3,12 +3,16 @@
 ##'
 ##' @title Generate Distance Matrix
 ##'
-##' @description Compute the distance matrix of using shortest paths of a (directed) \code{\link[igraph]{igraph}} structure, normalising by the diameter of the network, preserving node/column/row names (and direction).
+##' @description Compute the distance matrix of using shortest paths of a (directed)
+##' \code{\link[igraph]{igraph}} structure, normalising by the diameter of the network,
+##' preserving node/column/row names (and direction).
 ##'
 ##' @param mat precomputed adjacency or commonlink matrix.
 ##' @param graph An \code{\link[igraph]{igraph}} object. May be directed or weighted.
 ##' @param directed logical. Whether directed information is passed to the distance matrix.
-##' @param absolute logical. Whether distances are scaled as the absolute difference from the diameter (maximum possible). Defaults to TRUE. The alternative is to calculate a relative difference from the diameter for a geometric decay in distance.
+##' @param absolute logical. Whether distances are scaled as the absolute difference
+##' from the diameter (maximum possible). Defaults to TRUE. The alternative is to
+##' calculate a relative difference from the diameter for a geometric decay in distance.
 ##' @keywords graph network igraph adjacency
 ##' @importFrom igraph as_adjacency_matrix
 ##' @import igraph
@@ -45,13 +49,17 @@
 ##' graph_structure_distance_matrix_geom
 ##' # visualise matrix
 ##' library("gplots")
-##' heatmap.2(graph_structure_distance_matrix_geom, scale = "none", trace = "none", col = colorpanel(50, "white", "red"))
+##' heatmap.2(graph_structure_distance_matrix_geom, scale = "none", trace = "none",
+##'           col = colorpanel(50, "white", "red"))
 ##' # compute nodes with relationships between nodes (arithmetically decreasing)
-##' graph_structure_distance_matrix_abs <- make_distance_adjmat(graph_structure_adjacency_matrix, absolute = TRUE)
+##' graph_structure_distance_matrix_abs <- make_distance_adjmat(graph_structure_adjacency_matrix,
+##'                                                             absolute = TRUE)
 ##' graph_structure_distance_matrix_abs
 ##' # visualise matrix
 ##' library("gplots")
[…]

### 2020-06-18 R/generate.R

@@ -22,7 +22,13 @@
 ##' @param graph An \code{\link[igraph]{igraph}} object. May must be 
 ##' directed if states are used.
 ##' @param state numeric vector. Vector of length E(graph). Sign used
-##'  to calculate state matrix, may be an integer state or inferred directly from expected correlations for each edge. May be applied a scalar across all edges or as a vector for each edge respectively. May also be entered as text for "activating" or "inhibiting" or as integers for activating (0,1) or inhibiting (-1,2). Compatible with inputs for \code{\link[graphsim]{plot_directed}}. Also takes a pre-computed state matrix from \code{\link[graphsim]{make_state_matrix}} if applied to the same graph multiple times.
+##' to calculate state matrix, may be an integer state or inferred directly
+##' from expected correlations for each edge. May be applied a scalar across
+##' all edges or as a vector for each edge respectively. May also be entered
+##' as text for "activating" or "inhibiting" or as integers for activating (0,1)
+##' or inhibiting (-1,2). Compatible with inputs for \code{\link[graphsim]{plot_directed}}.
+##' Also takes a pre-computed state matrix from \code{\link[graphsim]{make_state_matrix}}
+##' if applied to the same graph multiple times.
 ##' @param cor numeric. Simulated maximum correlation/covariance of two 
 ##' adjacent nodes. Default to 0.8.
 ##' @param mean mean value of each simulated gene. Defaults to 0.
@@ -61,12 +67,16 @@
 ##' ##' # visualise matrix
 ##' library("gplots")
 ##' # expression data
-##' heatmap.2(test_data, scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"))
+##' heatmap.2(test_data, scale = "none", trace = "none",
+##'           col = colorpanel(50, "blue", "white", "red"))
 ##' # correlations
-##' heatmap.2(cor(t(test_data)), scale = "none", trace = "none", col = colorpanel(50, "white", "red"))
+##' heatmap.2(cor(t(test_data)), scale = "none", trace = "none",
+##'           col = colorpanel(50, "white", "red"))
 ##' # expected correlations (Σ)
[…]

### 2020-06-18 R/sigma.R

@@ -14,12 +14,22 @@
 ##'
 ##' @param mat precomputed adjacency, laplacian, commonlink, or scaled distance matrix.
 ##' @param graph An \code{\link[igraph]{igraph}} object. May be directed or weighted.
-##' @param state numeric vector. Vector of length E(graph). Sign used to calculate state matrix, may be an integer state or inferred directly from expected correlations for each edge. May be applied a scalar across all edges or as a vector for each edge respectively. May also be entered as text for "activating" or "inhibiting" or as integers for activating (0,1) or inhibiting (-1,2). Compatible with inputs for \code{\link[graphsim]{plot_directed}}. Also takes a pre-computed state matrix from \code{\link[graphsim]{make_state_matrix}} if applied to the same graph multiple times.
-##' @param cor numeric. Simulated maximum correlation/covariance of two adjacent nodes. Default to 0.8.
-##' @param sd	 standard deviations of each gene. Defaults to 1. May be entered as a scalar applying to all genes or a vector with a separate value for each. 
+##' @param state numeric vector. Vector of length E(graph). Sign used to calculate 
+##' state matrix, may be an integer state or inferred directly from expected correlations
+##' for each edge. May be applied a scalar across all edges or as a vector for each edge
+##' respectively. May also be entered as text for "activating" or "inhibiting" or as
+##' integers for activating (0,1) or inhibiting (-1,2). Compatible with inputs for 
+##' \code{\link[graphsim]{plot_directed}}. Also takes a pre-computed state matrix from
+##' \code{\link[graphsim]{make_state_matrix}} if applied to the same graph multiple times.
+##' @param cor numeric. Simulated maximum correlation/covariance of two adjacent nodes.
+##' Default to 0.8.
+##' @param sd	 standard deviations of each gene. Defaults to 1. May be entered as a scalar
+##' applying to all genes or a vector with a separate value for each. 
 ##' @param directed logical. Whether directed information is passed to the distance matrix.
[…]

### 2020-06-18 R/state.R

@@ -68,16 +68,25 @@
 ##' state_matrix_TFGBeta_Smad <- make_state_matrix(TGFBeta_Smad_graph)
 ##' # visualise matrix
 ##' library("gplots")
-##' heatmap.2(state_matrix_TFGBeta_Smad , scale = "none", trace = "none", dendrogram = "none", Rowv = FALSE, Colv = FALSE, col = colorpanel(50, "blue", "white", "red"))
+##' heatmap.2(state_matrix_TFGBeta_Smad , scale = "none", trace = "none",
+##'           dendrogram = "none", Rowv = FALSE, Colv = FALSE,
+##'           col = colorpanel(50, "blue", "white", "red"))
 ##' 
 ##' # compare the states to the sign of expected correlations in the sigma matrix
-##' sigma_matrix_TFGBeta_Smad_inhib <- make_sigma_mat_dist_graph(TGFBeta_Smad_graph, cor = 0.8, absolute = FALSE)
+##' sigma_matrix_TFGBeta_Smad_inhib <- make_sigma_mat_dist_graph(TGFBeta_Smad_graph,
+##'                                                              cor = 0.8,
+##'                                                              absolute = FALSE)
 ##' # visualise matrix
-##' heatmap.2(sigma_matrix_TFGBeta_Smad_inhib, scale = "none", trace = "none", dendrogram = "none", Rowv = FALSE, Colv = FALSE, col = colorpanel(50, "blue", "white", "red"))
+##' heatmap.2(sigma_matrix_TFGBeta_Smad_inhib,
+##'           scale = "none", trace = "none",
+##'           dendrogram = "none", Rowv = FALSE, Colv = FALSE,
+##'           col = colorpanel(50, "blue", "white", "red"))
 ##' 
 ##' # compare the states to the sign of final correlations in the simulated matrix
 ##' TFGBeta_Smad_data <- generate_expression(100, TGFBeta_Smad_graph, cor = 0.8)
-##' heatmap.2(cor(t(TFGBeta_Smad_data)), scale = "none", trace = "none", dendrogram = "none", Rowv = FALSE, Colv = FALSE, col = colorpanel(50, "blue", "white", "red"))
+##' heatmap.2(cor(t(TFGBeta_Smad_data)), scale = "none", trace = "none",
+##'           dendrogram = "none", Rowv = FALSE, Colv = FALSE,
+##'           col = colorpanel(50, "blue", "white", "red"))
 ##' 
 ##' 
 ##' @return An integer matrix indicating the resolved state

### 2020-06-18 R/laplacian.R
correct documentation for plotting function

@@ -4,7 +4,8 @@
 ##'
 ##' @title Generate Laplacian Matrix
 ##' 
-##' @description Compute the Laplacian matrix of a (directed) \code{\link[igraph]{igraph}} structure, preserving node/column/row names (and direction).
+##' @description Compute the Laplacian matrix of a (directed) \code{\link[igraph]{igraph}}
+##' structure, preserving node/column/row names (and direction).
 ##'
 ##' @param mat precomputed adjacency matrix.
 ##' @param graph An \code{\link[igraph]{igraph}} object. May be directed or weighted.
@@ -47,7 +48,8 @@
 ##' TGFBeta_Smad_laplacian_matrix[1:12, 1:12]
 ##' # visualise matrix
 ##' library("gplots")
-##' heatmap.2(TGFBeta_Smad_laplacian_matrix, scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"))
+##' heatmap.2(TGFBeta_Smad_laplacian_matrix, scale = "none", trace = "none",
+##'           col = colorpanel(50, "blue", "white", "red"))
 ##' 
 ##' @return An Laplacian matrix compatible with generating an expression matrix
 ##' 

### 2020-06-18 R/plot_directed.R

@@ -9,8 +9,9 @@
 #   Test Package:              'Ctrl + Shift + T'
 ##' @name plot_directed
 ##' @rdname plot_directed
+##' @aliases plot.directed
 ##'
-##' @title Extensions to iGraph for Customising plots
+##' @title Extensions to igraph for Customising plots
 ##'
 ##' @description Functions to plot_directed or graph structures including customised colours, layout, states, arrows. Uses graphs functions as an extension of \code{\link[igraph]{igraph}}. Designed for plotting directed graphs.
 ##'
@@ -53,32 +54,32 @@
 ##' @import igraph graphics
 ##' @examples
 ##'
-##' #generate example graphs
+##' # generate example graphs
 ##' library("igraph")
 ##' graph_structure_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"), c("D", "E"),
 ##'                            c("D", "F"), c("F", "G"), c("F", "I"), c("H", "I"))
 ##' graph_structure <- graph.edgelist(graph_structure_edges, directed = TRUE)
 ##'
-##' #plots with igraph defaults
+##' # plots with igraph defaults
 ##' plot(graph_structure, layout = layout.fruchterman.reingold)
 ##' plot(graph_structure, layout = layout.kamada.kawai)
 ##'
-##' #plots with scalar states
+##' # plots with scalar states
 ##' plot_directed(graph_structure, state="activating")
 ##' plot_directed(graph_structure, state="inhibiting")
 ##'
-##' #plots with vector states
+##' # plots with vector states
 ##' plot_directed(graph_structure, state = c(1, 1, 1, 1, -1, 1, 1, 1))
 ##' plot_directed(graph_structure, state = c(1, 1, -1, 1, -1, 1, -1, 1))
-##' plot_directed(graph_structure, state = c(1, 1, -1, 1, 1, 1, 1, -1)
+##' plot_directed(graph_structure, state = c(1, 1, -1, 1, 1, 1, 1, -1))
 ##' 
-##' #plots states with graph attributes
+##' # plots states with graph attributes
 ##' E(graph_structure)$state <- 1
 ##' plot_directed(graph_structure)
 ##' E(graph_structure)$state <- c(1, 1, -1, 1, -1, 1, -1, 1)
 ##' plot_directed(graph_structure)
 ##'
-##' #plot layout customised
+##' # plot layout customised
[…]

### 2020-06-18 tests/testthat/test_sigma.R
test plotting functions

@@ -74,7 +74,7 @@ test_that("Generate Sigma matrix from distance matrix", {
 
 ## all with graph
 
-test_that("Generate distance relation matrix matrix from graph structure", {
+test_that("Generate Sigma matrix  from graph structure", {
   graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
   graph_test1 <- graph.edgelist(graph_test1_edges, directed = TRUE)
   adjacency_matrix1 <- make_adjmatrix_graph(graph_test1)
@@ -92,7 +92,7 @@ test_that("Generate distance relation matrix matrix from graph structure", {
 })
 
 
-test_that("Generate distance relation matrix matrix from graph structure", {
+test_that("Generate Sigma matrix  from graph structure", {
   graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
   graph_test1 <- graph.edgelist(graph_test1_edges, directed = TRUE)
   adjacency_matrix1 <- make_adjmatrix_graph(graph_test1)
@@ -109,7 +109,7 @@ test_that("Generate distance relation matrix matrix from graph structure", {
   expect_true(all(sigma_matrix1 %in% c(0, 1, 0.8/2^c(0:100))))
 })
 
-test_that("Generate distance relation matrix matrix from graph structure", {
+test_that("Generate Sigma matrix  from graph structure", {
   graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))
   graph_test1 <- graph.edgelist(graph_test1_edges, directed = TRUE)
   adjacency_matrix1 <- make_adjmatrix_graph(graph_test1)
@@ -124,4 +124,37 @@ test_that("Generate distance relation matrix matrix from graph structure", {
   expect_true(all(adjacency_matrix1 == cbind(c(0, 1, 0, 0), c(1, 0, 1, 1), c(0, 1, 0, 0), c(0, 1, 0, 0))))
   expect_true(all(sigma_matrix1 == cbind(c(1, 0, 0.8, 0.8), c(0, 1, 0, 0), c(0.8, 0, 1, 0.8), c(0.8, 0, 0.8, 1))))
[…]

### 2020-06-24 paper/paper.Rmd
test Mac OS on Travis
update paper introduction text (biorXiv version)

@@
 SERGIO [@Dibaeinia2019] and dyngen [@Cannoodt2020] build on
-this to include gene regulatory networks and multimodality
+thisby adding gene regulatory networks and multimodality
 respectively. These methods have been designed based on known
-or modelled reaction states and added stochasticity to determinisitic
+or synthetic reaction states and added stochasticity to deterministic
 relationships. However, it is computationally-intensive to model
@@
 In some cases, it is only necessary to model the statistical
-variability of RNA-Seq data to evaluate methods with these
-correlation structures.
+variability and ``noise'' of RNA-Seq data to evaluate methods
+with multivariate correlation structures.
 

### 2020-06-24 paper/paper.Rmd
update paper formatting (biorXiv version)

@@
 $\Sigma$ with negative correlations).\
+
 Computing the nearest positive definite matrix is necessary to ensure
@@
 \texttt{igraph} objects in and is available in the and packages. The
-\texttt{plot_directed} function enables customisation of plot parameters for
+\texttt{plot\_directed} function enables customisation of plot parameters for
 each node or edge, and mixed (directed) edge types for indicating

### 2020-06-24 paper/paper.Rmd
update figure placement settings

@@
   - \usepackage{caption}
- # - \usepackage{subfig}
+  - \include{header.tex}
 ---
@@
 knitr::opts_chunk$set(fig.cap = "", fig.path = "Plot")
+knitr::opts_chunk$set(fig.pos = "!htbp")
 options(width = 68, cli.unicode = FALSE, cli.width = 68)
@@
 ```{r simulation_activating_hide, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.375\\linewidth', out.width='.375\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, eval=FALSE, warning=FALSE}
-# activating graph
-state <- rep(1, length(E(graph)))
-plot_directed(graph, state=state, layout = layout.kamada.kawai,
-  cex.node=2, cex.arrow=4, arrow_clip = 0.2)
-
 #plot relationship matrix

### 2020-06-24 paper/paper.Rmd
update figure placement settings

@@
   - \include{header.tex}
+  - \let\origfigure\figure |
+  \let\endorigfigure\endfigure |
+\renewenvironment{figure}[1][2] { |
+    \expandafter\origfigure\expandafter[!htbp]
+} { |
+    \endorigfigure |
+}
 ---

### 2020-06-24 paper/paper.Rmd
update figure placement settings

--- 2020-06-24 paper/paper.Rmd
@@
   - \include{header.tex}
+  - \let\origfigure\figure \let\endorigfigure\endfigure \renewenvironment{figure}[1][2] { \expandafter\origfigure\expandafter[!htbp] } { \endorigfigure }
 ---

### 2020-06-24 paper/paper.Rmd
update figure placement settings

--- 2020-06-24 paper/paper.Rmd
@@
   - \usepackage{caption}
-  - \include{header.tex}
+  - \let\origfigure\figure
+  - \let\endorigfigure\endfigure
+  - \renewenvironment{figure}[1][2] { \expandafter\origfigure\expandafter[!htbp] } { \endorigfigure }
 ---
@@
 knitr::opts_chunk$set(fig.cap = "", fig.path = "Plot")
-knitr::opts_chunk$set(fig.pos = "!htbp")
+#knitr::opts_chunk$set(fig.pos = "!htbp")
 options(width = 68, cli.unicode = FALSE, cli.width = 68)

### 2020-06-24 paper/paper.Rmd
update figure placement settings

@@
   - \usepackage{caption}
-  - \let\origfigure\figure
-  - \let\endorigfigure\endfigure
-  - \renewenvironment{figure}[1][2] { \expandafter\origfigure\expandafter[!htbp] } { \endorigfigure }
 ---

### 2020-06-24 paper/paper.Rmd
update figure placement settings

@@
     keep_md: TRUE
-    keep_tex: TRUE
+    #keep_tex: TRUE
     fig_caption: yes
@@
 knitr::opts_chunk$set(fig.cap = "", fig.path = "Plot")
-#knitr::opts_chunk$set(fig.pos = "!htbp")
+knitr::opts_chunk$set(fig.pos = "!htbp")
 options(width = 68, cli.unicode = FALSE, cli.width = 68)
@@
 
+The following changes are needed to handle inhibitions:
+
 ```{r, include = FALSE}
@@
 functions as as an "inhibitor SMAD" with competitively inhibits *SMAD4*.
+
+We can import the graph structure into R as follows and run simulations as above:
 

### 2020-06-30 R/commonlink.R
update simulation vignettes for clarity

@@ -4,7 +4,9 @@
 ##' @title Generate Common Link Matrix
 ##'
 ##' @description Compute the common link matrix of a (directed) \code{\link[igraph]{igraph}}
-##' structure, preserving node / column / row names (and direction).
+##' structure, preserving node / column / row names (and direction). We can compute the common 
+##' inks between each pair of nodes. This shows how many nodes are mutually connected to both
+##' of the nodes in the matrix (how many paths of length 2 exist between them).
 ##'
 ##' @param adj_mat precomputed adjacency matrix.
 ##' @param graph An \code{\link[igraph]{igraph}} object. May be directed or weighted.

### 2020-06-30 paper/paper.Rmd
increment version for major release
update manscript with feedback

@@
     fig_crop: no
-    keep_md: TRUE
-    #keep_tex: TRUE
+    #keep_md: TRUE
+    keep_tex: TRUE
     fig_caption: yes
@@
 analysis. As a result, databases that describe biological pathways 
-are increasingly relied on to assist with the interpretation of results
+are increasingly uesad to assist with the interpretation of results
 from large-scale genomics studies. Incorporating information from 
@@
 such methods, it is important to gain an accurate assessment of their 
-performance, with simulation-based validation studies a popular choice.
+performance. Simulation-based validation studies are frequently used
+for this.
 This necessitates the use of simulated data that correctly accounts for
@@
 well-defined, experimentally-validated, and are available in public
-databases [@Reactome]. As a result, network analysis techniques could
+databases [@Reactome]. As a result, network analysis techniques can
 play an important role in furthering our understanding of biological
@@
 enhancers, and sites of epigenetic marks or chromatin structures
-[@Barabasi2004; @Yamaguchi2007]. Inference of these regulatory
-interactions for genomics investigations has the potential to radically
+[@Barabasi2004; @Yamaguchi2007]. Inference using these regulatory
+interactions genomic analysis has the potential to radically
 expand the range of candidate biological pathways to be further
 explored, or to improve the accuracy of bioinformatics and functional
-genomic analysis. A number of methods have already been developed to
-utilise timecourse gene expression data [@Arner2015; @Yamaguchi2007]
+genomic analysis. A number of methods have been developed to
+utilise timecourse gene expression data [@Yamaguchi2007; @Arner2015]
 using gene regulatory modules in state-space models and recursive vector
@@
 to gene regulation and networks at the genome-wide scale have led to
-novel biological insights [@Arner2015; @Komatsu2013]. However, inference
+novel biological insights [@Arner2015; @Komatsu2013], however, inference
 of regulatory networks has thus far primarily relied on experimental
@@
 
-Simulated datasets with pathway structure are of particular interest for
-benchmarking regulatory network inference techniques and methods
-being developed for genomics data with complex biological 
+Simulating datasets that account for pathway structure are of particular
+interest for benchmarking regulatory network inference techniques
+and methods being developed for genomics data containing complex biological 
 interactions [@Schaffter2011; @Saelens2019]. Dynamical models using
@@
 specifically for benchmarking gene regulatory network inference techniques.
-There is renewed interest in modelling biological pathways and simulating
+There is also renewed interest in modelling biological pathways and simulating
 data for benchmarking due to the emergence of single-cell genomics
-technologies and the plethora of bioinformatics techniques developed
+technologies and the growing number of bioinformatics techniques developed
 to use this data [@Zappia2017; @Saelens2019]. Packages such as 'splatter'
@@
 SERGIO [@Dibaeinia2019] and dyngen [@Cannoodt2020] build on
-thisby adding gene regulatory networks and multimodality
+this by adding gene regulatory networks and multimodality
 respectively. These methods have been designed based on known
-or synthetic reaction states and added stochasticity to deterministic
-relationships. However, it is computationally-intensive to model
-these reactions at scale over iterations in benchmarking.
+deterministic relationships or synthetic reaction states,
+to which stochasticity is then added.
+However, it is computationally-intensive to model
+these reactions at scale or run many iterations for benchmarking.
 In some cases, it is only necessary to model the statistical
-variability and ``noise'' of RNA-Seq data to evaluate methods
-with multivariate correlation structures.
+variability and "noise" of RNA-Seq data in order to evaluate
+methods in the presence of multivariate correlation structures.
 
@@
 simulation of expression profiles that approximate the log-transformed
-and normalised data from microarray and bulk or single-cell RNA-Seq
-experiments. This procedure has been released as the package to enable
-the generation of simulated gene expression datasets containing pathway
-relationships from a known underlying network. These simulated datasets
-can be used to evaluate various bioinformatics methodologies, including
+and normalised data from microarray studies, as well as, bulk or single-cell RNA-Seq
+experiments. This procedure has been released as the \texttt{graphsim}
+package to enable the generation of simulated gene expression datasets
+containing pathway relationships from a known underlying network.
+These simulated datasets can be used to evaluate various bioinformatics
+methodologies, including
 statistical and network inference procedures.
@@
 approach has also been applied for modelling UMI-based count data from
-single-cell RNA-Seq experiments in the package [@Wang2018].
+single-cell RNA-Seq experiments in the \texttt{DESCEND} R package [@Wang2018].
 
@@
 Computing the nearest positive definite matrix is necessary to ensure
-that the variance-covariance matrix could be inverted when used as a
+that the variance-covariance matrix can be inverted when used as a
 parameter in multivariate normal simulations, particularly when negative
 correlations are included for inhibitions (as shown below). Matrices
-that could not be inverted occurred rarely with biologically plausible
+that cannot be inverted occur rarely with biologically plausible
 graph structures but this approach allows for the computation of a
@@
 incomplete or contains loops. When required, the nearest positive
-definite matrix is computed using the `nearPD` function of the \texttt{Matrix} R package
-[@Matrix] to perform Higham's algorithm [@Higham2002] on
-variance-covariance matrices. The \texttt{graphsim} package gives a warning when this
-occurs.
+definite matrix is computed using the `nearPD` function of
+the \texttt{Matrix} R package [@Matrix] to perform
+Higham's algorithm [@Higham2002] on variance-covariance matrices.
+The \texttt{graphsim} package gives a warning when this occurs.
 
@@
 expression data by sampling from a multivariate normal distribution
-using the R package [@Genz2009; @mvtnorm]. The graph structure
+using the \texttt{mvtnorm} R package [@Genz2009; @mvtnorm]. The graph structure
 visualisation in
@@
 reference="fig:simple_graph"} was specifically developed for (directed)
-\texttt{igraph} objects in and is available in the and packages. The
+\texttt{igraph} objects in and is available in the \texttt{graphsim} package. The
 \texttt{plot\_directed} function enables customisation of plot parameters for
@@
 ```{r, eval = FALSE}
-#install packages required (once per machine)
+#install packages required (once per computer)
 install.packages("graphsim")
@@
 A constructed graph structure used as an example to demonstrate the simulation procedure in Figures 2 and 3.
-Activating links are denoted by black arrows and inhibiting links by red edges.
-Inhibiting edges have been highlighted in red.', echo=FALSE}
+Activating links are denoted by black arrows and inhibiting links by red edges.', echo=FALSE}
 #generate graph structure
@@
 
-#simulated data
+#simulate data
 expr <- generate_expression(100, graph, cor = 0.8, mean = 0,
@@
 An example of a graph structure (a) that has been used to derive a relationship matrix (b), $\\Sigma$  matrix (c) and correlation structure (d) from the relative distances between the nodes.
-Non-negative values are coloured white to red from $0$ to $1$.
-This $\\Sigma$ matrix has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
+Non-negative values are coloured white to red from $0$ to $1$ (e).
+The $\\Sigma$ matrix has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
 Here genes with closer relationships in the pathway structure show a higher correlation between simulated values.', warning=FALSE, echo=FALSE}
@@
 heatmap.2(make_distance_graph(graph, absolute = FALSE),
-  scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
+  scale = "none", trace = "none", key = FALSE,
+  col = colorpanel(50, "white", "red"),
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
@@
 heatmap.2(make_sigma_mat_dist_graph(graph, cor = 0.8, absolute = FALSE),
-scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
+scale = "none", trace = "none", key = FALSE,
+col = colorpanel(50, "white", "red"),
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
 mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3.5, at=0, adj=0.5, cex=1.75)
-#simulated data
+#simulate data
 expr <- generate_expression(100, graph, cor = 0.8, mean = 0,
@@
 #plot simulated correlations
-heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
+heatmap.2(cor(t(expr)), scale = "none", trace = "none", key = FALSE, 
+          col = colorpanel(50, "white", "red"),
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
@@
 #plot simulated expression data
-heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+heatmap.2(expr, scale = "none", trace = "none", key = FALSE, col = bluered(50),
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
@@
 #plot simulated expression data
-heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
+heatmap.2(expr, scale = "none", trace = "none", key = FALSE, col = bluered(50),
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "", mar = c(6, 6), keysize = 1)
@@
 ```{r simulation_inhibiting, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.375\\linewidth', out.width='.375\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from graph structure with inhibitions}.
-An example of a graph structure (a), that has been used to derive a relationship matrix (b), $\\Sigma$ matrix (c), and correlation structure (d), from the relative distances between the nodes.
+An example of a graph structure (a), that has been used to derive a relationship matrix (b), $\\Sigma$ matrix (c), and correlation structure (d), from the relative distances between the nodes (e).
 These values are coloured blue to red from $-1$ to $1$.
@@
 heatmap.2(make_distance_graph(graph, absolute = FALSE),
-  scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
+  scale = "none", trace = "none", key = FALSE, col = colorpanel(50, "white", "red"),
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
@@
 heatmap.2(make_sigma_mat_dist_graph(graph, state, cor = 0.8, absolute = FALSE),
-scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
+scale = "none", trace = "none", key = FALSE, col = colorpanel(50, "blue", "white", "red"),
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
@@
 #plot simulated correlations
-heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
+heatmap.2(cor(t(expr)), scale = "none", trace = "none", key = FALSE, col = colorpanel(50, "blue", "white", "red"),
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
@@
 reference="fig:simulation_smad"}) on a pathway structure for a known
-biological pathway (from reactome R-HSA-2173789) of TGF-$\beta$ receptor
-signaling activates SMADs
+biological pathway (Reactome pathway R-HSA-2173789): "TGF-$\beta$ receptor
+signaling activates SMADs"
 (Figure [4a](#fig:simulation_smad:first){reference-type="ref"
 reference="fig:simulation_smad:first"}) derived from the Reactome
-database version 52 `[@Reactome]`. Distances are calculated in the same
+database version 52 [@Reactome]. Distances are calculated in the same
 manner as before
@@
 and [4d](#fig:simulation_smad:fifth){reference-type="ref"
-reference="fig:simulation_smad:fifth"}). This shows that
+reference="fig:simulation_smad:fifth"}). This shows that the
 multivariate normal distribution can be sampled to generate simulated
@@
 reference="fig:simulation_smad:fourth"}). Here *SMAD7* exhibits
-negative correlations with the other SMADs consistent with it's
-functions as as an "inhibitor SMAD" with competitively inhibits *SMAD4*.
+negative correlations with the other SMADs consistent with its
+functions as an "inhibitor SMAD" which competitively inhibits *SMAD4*.
 
@@
 ```{r simulation_smad, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.375\\linewidth', out.width='.375\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a biological pathway graph structure}.
-The graph structure (a) of a known biological pathway, the TGF-$\\beta$ receptor signaling activates SMADs (R-HSA-2173789), was used to derive a relationship matrix (b), $\\Sigma$ matrix (c) and correlation structure (d) from the relative distances between the nodes.
-These values are coloured blue to red from $-1$ to $1$.
+The graph structure (a) of a known biological pathway,  "TGF-$\\beta$ receptor signaling activates SMADs" (R-HSA-2173789), was used to derive a relationship matrix (b), $\\Sigma$ matrix (c) and correlation structure (d) from the relative distances between the nodes.
+These values are coloured blue to red from $-1$ to $1$ (e).
 This has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution.
@@
 heatmap.2(make_distance_graph(graph, absolute = FALSE),
-  scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
+  scale = "none", trace = "none", key = FALSE, col = colorpanel(50, "white", "red"),
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
@@
 heatmap.2(make_sigma_mat_dist_graph(graph, state, cor = 0.8, absolute = FALSE),
-scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
+scale = "none", trace = "none", key = FALSE, col = colorpanel(50, "blue", "white", "red"),
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
@@
 #plot simulated correlations
-heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
+heatmap.2(cor(t(expr)), scale = "none", trace = "none", key = FALSE, col = colorpanel(50, "blue", "white", "red"),
 colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
@@
 expression data within a graph network to test genomic analysis techniques.
-Correlation structure can be included into datasets generated
-when testing whether true positive genes or samples can be detected
+Correlation structure can be included in datasets generated
+for testing whether true positive genes or samples can be detected
 in a sample with the background of complex pathway structure.
@@
 but could in principle be adapted to modelling single-cell or
-different modalities such as epigenomics data.
+different modalities such as genome-wide epigenetic data.
 
@@
 
-The results in this paper were obtained using R 3.6.1 with the \texttt{igraph} 1.2.4.1
-\texttt{Matrix} 1.2-17, \texttt{matrixcalc} 1.0-3, and \texttt{mvtnorm} 1.0-11 packages. R itself and all dependent packages
+The results in this paper were obtained using R 4.0.2 with the \texttt{igraph} 1.2.5
+\texttt{Matrix} 1.2-17,
+\texttt{matrixcalc} 1.0-3, and \texttt{mvtnorm} 1.1-1 packages.
+R itself and all dependent packages
 used are available from the Comprehensive Archive Network (CRAN) at
-\url{https://CRAN.R-project.org}. The \texttt{graphsim} package presented can be
-installed from CRAN and the issues can  be reported to the development version on GitHub (\url{https://github.com/TomKellyGenetics/graphsim}).
-This package is included in the library on GitHub (\url{https://github.com/TomKellyGenetics/igraph.extensions}) which installs various
+\url{https://CRAN.R-project.org}. The \texttt{graphsim} 1.0.0 package
+presented can be installed from CRAN and the issues can  be reported to
+the development version on GitHub (\url{https://github.com/TomKellyGenetics/graphsim}).
+This package is included in the \texttt{igraph.extensions} library on GitHub (\url{https://github.com/TomKellyGenetics/igraph.extensions})
+which installs various
 tools for \texttt{igraph} analysis. This software is cross-platform and
@@
 information and examples on running functions released in the package.
-The package (\texttt{graphsim} 0.1.2) has been released on CRAN and will be updated.
+Updates to the  package (\texttt{graphsim} 1.0.0) will be released on CRAN.
+
+Complete examples of code needed to produce the figures in this paper are
+available in the Rmarkdown version in the package GitHub repository
+ (\url{https://github.com/TomKellyGenetics/graphsim/paper}).
+ Further details are available in the vignettes as well.
 

### 2020-07-01 R/distance.R
tests passing

@@ -121,6 +121,24 @@ make_distance_adjmat <- function(mat, directed = FALSE, absolute = FALSE){
 }
 
 ##' @rdname make_distance
+##' @export
+make_distance_comm <- function(mat, directed = FALSE, absolute = FALSE){
+  diag(mat) <- 0
+  if (absolute){
+    mat <- mat / max (mat)
+    diag(mat) <- 1
+  } else {
+    diam <- max(mat)
+    mat[mat == 0] <- NA
+    mat <- 1^-1/mat
+    mat[is.na(mat)] <- 0
+    diag(mat) <- 1
+  }
+  rownames(mat) <- colnames(mat) <- names(V(graph))
+  return(mat)
+}
+
+##' @rdname make_distance
 ##' @importFrom igraph graph_from_adjacency_matrix laplacian_matrix
 ##' @export
 make_distance_laplacian <- function(mat, directed = FALSE, absolute = FALSE){

### 2020-07-01 tests/testthat/test_adjmatrix.R

@@ -1,6 +1,6 @@
 library("graphsim")
 library("igraph")
-context("Make Adjancency Matrix")
+context("Make Adjacency Matrix")
 
 test_that("Generate adjacency matrix from graph structure", {
   graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))

### 2020-07-01 tests/testthat/test_distance.R

@@ -27,10 +27,10 @@ test_that("Generate distance relation matrix from commonlink matrix", {
   expect_true(all(diag(distance_matrix1) == rep(1, nrow(adjacency_matrix1))))
   expect_equal(nrow(distance_matrix1), length(V(graph_test1)))
   expect_equal(ncol(distance_matrix1), length(V(graph_test1)))
-  expect_equal(sum(distance_matrix1), 10)
+  expect_equal(sum(distance_matrix1), length(V(graph_test1)))
   expect_true(all(adjacency_matrix1 == cbind(c(0, 1, 0, 0), c(1, 0, 1, 1), c(0, 1, 0, 0), c(0, 1, 0, 0))))
   expect_true(all(common_link_matrix1 == cbind(c(1, 0, 1, 1), c(0, 3, 0, 0), c(1, 0, 1, 1), c(1, 0, 1, 1))))
-  expect_true(all(distance_matrix1 == cbind(c(1, 0, 1, 1), c(0, 1, 0, 0), c(1, 0, 1, 1), c(1, 0, 1, 1))))
+  expect_true(all(distance_matrix1 == diag(4)))
   expect_true(all(distance_matrix1 %in% c(0, 1/2^c(0:100))))
 })
 

### 2020-07-01 tests/testthat/test_laplacian.R

@@ -1,6 +1,6 @@
 library("graphsim")
 library("igraph")
-context("Make Commonlink Matrix")
+context("Make Laplacian Matrix")
 
 test_that("Generate Laplacian matrix from adjacency matrix", {
   graph_test1_edges <- rbind(c("A", "B"), c("B", "C"), c("B", "D"))

### 2020-07-01 R/adjmat.R
remove links to aliases

@@ -4,10 +4,10 @@
 ##'
 ##' @title Generate Adjacency Matrix
 ##' 
-##' @description Compute the adjacency matrix of a (directed) \code{\link[igraph]{igraph}}
+##' @description Compute the adjacency matrix of a (directed) \code{\link[igraph]{igraph-package}}
 ##' structure, preserving node/column/row names (and direction).
 ##'
-##' @param graph An \code{\link[igraph]{igraph}} object. May be directed or weighted.
+##' @param graph An \code{\link[igraph]{igraph-package}} object. May be directed or weighted.
 ##' @param directed logical. Whether directed information is passed to the adjacency matrix.
 ##' @keywords graph network igraph adjacency
 ##' @importFrom igraph as_adjacency_matrix graph.edgelist as.undirected

### 2020-07-01 R/commonlink.R

@@ -3,13 +3,13 @@
 ##'
 ##' @title Generate Common Link Matrix
 ##'
-##' @description Compute the common link matrix of a (directed) \code{\link[igraph]{igraph}}
+##' @description Compute the common link matrix of a (directed) \code{\link[igraph]{igraph-package}}
 ##' structure, preserving node / column / row names (and direction). We can compute the common 
 ##' inks between each pair of nodes. This shows how many nodes are mutually connected to both
 ##' of the nodes in the matrix (how many paths of length 2 exist between them).
 ##'
 ##' @param adj_mat precomputed adjacency matrix.
-##' @param graph An \code{\link[igraph]{igraph}} object. May be directed or weighted.
+##' @param graph An \code{\link[igraph]{igraph-package}} object. May be directed or weighted.
 ##' @param directed logical. Whether directed information is passed to the adjacency matrix.
 ##' @keywords graph network igraph neighbourhood
 ##' @import igraph

### 2020-07-01 R/distance.R

@@ -4,11 +4,11 @@
 ##' @title Generate Distance Matrix
 ##'
 ##' @description Compute the distance matrix of using shortest paths of a (directed)
-##' \code{\link[igraph]{igraph}} structure, normalising by the diameter of the network,
+##' \code{\link[igraph]{igraph-package}} structure, normalising by the diameter of the network,
 ##' preserving node/column/row names (and direction).
 ##'
 ##' @param mat precomputed adjacency or commonlink matrix.
-##' @param graph An \code{\link[igraph]{igraph}} object. May be directed or weighted.
+##' @param graph An \code{\link[igraph]{igraph-package}} object. May be directed or weighted.
 ##' @param directed logical. Whether directed information is passed to the distance matrix.
 ##' @param absolute logical. Whether distances are scaled as the absolute difference
 ##' from the diameter (maximum possible). Defaults to TRUE. The alternative is to

### 2020-07-01 R/generate.R

@@ -4,22 +4,22 @@
 ##' @title Generate Simulated Expression
 ##'
 ##' @description Compute simulated continuous expression data from a graph 
-##' network structure. Requires an \code{\link[igraph]{igraph}} pathway 
+##' network structure. Requires an \code{\link[igraph]{igraph-package}} pathway 
 ##' structure and a matrix of states (1 for activating and -1 for 
 ##' inhibiting) for link signed correlations, from a vector of edge states 
 ##' to a signed adjacency matrix for use in 
 ##' \code{\link[graphsim]{generate_expression}}. 
 ##' Uses graph structure to pass a sigma covariance matrix from 
-##' \code{\link[graphsim]{make_sigma_mat_dist_graph}} or 
-##' \code{\link[graphsim]{make_sigma_mat_graph}} on to 
-##' \code{\link[mvtnorm]{rmvnorm}}. By default data is generated with a mean of
+##' \code{\link[graphsim]{make_sigma}} or 
+##' \code{\link[graphsim]{make_sigma}} on to 
+##' \code{\link[mvtnorm]{Mvnorm}}. By default data is generated with a mean of
 ##'  0 and standard deviation of 1 for each gene (with correlations between 
 ##'  derived from the graph structure).
 ##'
 ##' @param n number of observations (simulated samples).
 ##' @param mat precomputed adjacency, laplacian, commonlink, or scaled 
 ##' distance matrix.
-##' @param graph An \code{\link[igraph]{igraph}} object. May must be 
+##' @param graph An \code{\link[igraph]{igraph-package}} object. May must be 
 ##' directed if states are used.
 ##' @param state numeric vector. Vector of length E(graph). Sign used
 ##' to calculate state matrix, may be an integer state or inferred directly
@@ -38,12 +38,12 @@
 ##' May be entered as a scalar applying to 
 ##' all genes or a vector with a separate value for each.
 ##' @param dist logical. Whether a graph distance 
-##' (\code{\link[graphsim]{make_sigma_mat_dist_graph}}) or derived matrix
-##'  (\code{\link[graphsim]{make_sigma_mat_graph}}) is used to compute the
+##' (\code{\link[graphsim]{make_sigma}}) or derived matrix
+##'  (\code{\link[graphsim]{make_sigma}}) is used to compute the
 ##'   sigma matrix.
[…]

### 2020-07-01 R/laplacian.R

@@ -4,11 +4,11 @@
 ##'
 ##' @title Generate Laplacian Matrix
 ##' 
-##' @description Compute the Laplacian matrix of a (directed) \code{\link[igraph]{igraph}}
+##' @description Compute the Laplacian matrix of a (directed) \code{\link[igraph]{igraph-package}}
 ##' structure, preserving node/column/row names (and direction).
 ##'
 ##' @param mat precomputed adjacency matrix.
-##' @param graph An \code{\link[igraph]{igraph}} object. May be directed or weighted.
+##' @param graph An \code{\link[igraph]{igraph-package}} object. May be directed or weighted.
 ##' @param directed logical. Whether directed information is passed to the Laplacian matrix.
 ##' @keywords graph network igraph Laplacian
 ##' @importFrom igraph laplacian_matrix graph.edgelist as.undirected graph_from_adjacency_matrix

### 2020-07-01 R/plot_directed.R

@@ -13,9 +13,9 @@
 ##'
 ##' @title Extensions to igraph for Customising plots
 ##'
-##' @description Functions to plot_directed or graph structures including customised colours, layout, states, arrows. Uses graphs functions as an extension of \code{\link[igraph]{igraph}}. Designed for plotting directed graphs.
+##' @description Functions to plot_directed or graph structures including customised colours, layout, states, arrows. Uses graphs functions as an extension of \code{\link[igraph]{igraph-package}}. Designed for plotting directed graphs.
 ##'
-##' @param graph An \code{\link[igraph]{igraph}} object. Must be directed with known states.
+##' @param graph An \code{\link[igraph]{igraph-package}} object. Must be directed with known states.
 ##' @param state character or integer. Defaults to "activating" if no "state" edge attribute 
 ##' found. May be applied a scalar across all edges or as a vector for each edge respectively. 
 ##' Accepts non-integer values for weighted edges provided that the sign indicates whether links

### 2020-07-01 R/sigma.R

@@ -3,8 +3,8 @@
 ##'
 ##' @title Generate Sigma (\eqn{\Sigma}) Matrix
 ##'
-##' @description Compute the Sigma (\eqn{\Sigma}) matrix from an \code{\link[igraph]{igraph}} structure 
-##' or pre-computed matrix. These are compatible with \code{\link[mvtnorm]{rmvnorm}} and
+##' @description Compute the Sigma (\eqn{\Sigma}) matrix from an \code{\link[igraph]{igraph-package}} structure 
+##' or pre-computed matrix. These are compatible with \code{\link[mvtnorm]{Mvnorm}} and
 ##'  \code{\link[graphsim]{generate_expression}}.
 ##' By default data is generated with a mean of 0 and standard deviation of 1 for 
 ##' each gene (with correlations between derived from the graph structure).
@@ -13,7 +13,7 @@
 ##' between each gene in the output from \code{\link[graphsim]{generate_expression}}.
 ##'
 ##' @param mat precomputed adjacency, laplacian, commonlink, or scaled distance matrix.
-##' @param graph An \code{\link[igraph]{igraph}} object. May be directed or weighted.
+##' @param graph An \code{\link[igraph]{igraph-package}} object. May be directed or weighted.
 ##' @param state numeric vector. Vector of length E(graph). Sign used to calculate 
 ##' state matrix, may be an integer state or inferred directly from expected correlations
 ##' for each edge. May be applied a scalar across all edges or as a vector for each edge

### 2020-07-01 R/state.R

@@ -8,7 +8,7 @@
 ##' in \code{\link[graphsim]{generate_expression}}. This resolves edge states to determine the sign
 ##' of all correlations between nodes in a network. These are computed interally for sigma matrices
 ##' as required.
-##' @param graph An \code{\link[igraph]{igraph}} object. May be directed or weighted as long as
+##' @param graph An \code{\link[igraph]{igraph-package}} object. May be directed or weighted as long as
 ##' a shortest path can be computed.
 ##' @param state numeric vector. Vector of length E(graph). Sign used to calculate state matrix,
 ##' may be an integer state or inferred directly from expected correlations for each edge. May be

### 2020-07-01 R/adjmat.R
correct links in documentation (link to file) https://github.com/r-lib/roxygen2/issues/707

@@ -4,10 +4,10 @@
 ##'
 ##' @title Generate Adjacency Matrix
 ##' 
-##' @description Compute the adjacency matrix of a (directed) \code{\link[igraph]{igraph-package}}
+##' @description Compute the adjacency matrix of a (directed) \code{\link[igraph:aaa-igraph-package]{igraph}}
 ##' structure, preserving node/column/row names (and direction).
 ##'
-##' @param graph An \code{\link[igraph]{igraph-package}} object. May be directed or weighted.
+##' @param graph An \code{\link[igraph:aaa-igraph-package]{igraph}} object. May be directed or weighted.
 ##' @param directed logical. Whether directed information is passed to the adjacency matrix.
 ##' @keywords graph network igraph adjacency
 ##' @importFrom igraph as_adjacency_matrix graph.edgelist as.undirected

### 2020-07-01 R/commonlink.R

@@ -3,13 +3,13 @@
 ##'
 ##' @title Generate Common Link Matrix
 ##'
-##' @description Compute the common link matrix of a (directed) \code{\link[igraph]{igraph-package}}
+##' @description Compute the common link matrix of a (directed) \code{\link[igraph:aaa-igraph-package]{igraph}}
 ##' structure, preserving node / column / row names (and direction). We can compute the common 
 ##' inks between each pair of nodes. This shows how many nodes are mutually connected to both
 ##' of the nodes in the matrix (how many paths of length 2 exist between them).
 ##'
 ##' @param adj_mat precomputed adjacency matrix.
-##' @param graph An \code{\link[igraph]{igraph-package}} object. May be directed or weighted.
+##' @param graph An \code{\link[igraph:aaa-igraph-package]{igraph}} object. May be directed or weighted.
 ##' @param directed logical. Whether directed information is passed to the adjacency matrix.
 ##' @keywords graph network igraph neighbourhood
 ##' @import igraph

### 2020-07-01 R/distance.R

@@ -4,11 +4,11 @@
 ##' @title Generate Distance Matrix
 ##'
 ##' @description Compute the distance matrix of using shortest paths of a (directed)
-##' \code{\link[igraph]{igraph-package}} structure, normalising by the diameter of the network,
+##' \code{\link[igraph:aaa-igraph-package]{igraph}} structure, normalising by the diameter of the network,
 ##' preserving node/column/row names (and direction).
 ##'
 ##' @param mat precomputed adjacency or commonlink matrix.
-##' @param graph An \code{\link[igraph]{igraph-package}} object. May be directed or weighted.
+##' @param graph An \code{\link[igraph:aaa-igraph-package]{igraph}} object. May be directed or weighted.
 ##' @param directed logical. Whether directed information is passed to the distance matrix.
 ##' @param absolute logical. Whether distances are scaled as the absolute difference
 ##' from the diameter (maximum possible). Defaults to TRUE. The alternative is to

### 2020-07-01 R/laplacian.R

@@ -4,11 +4,11 @@
 ##'
 ##' @title Generate Laplacian Matrix
 ##' 
-##' @description Compute the Laplacian matrix of a (directed) \code{\link[igraph]{igraph-package}}
+##' @description Compute the Laplacian matrix of a (directed) \code{\link[igraph:aaa-igraph-package]{igraph}}
 ##' structure, preserving node/column/row names (and direction).
 ##'
 ##' @param mat precomputed adjacency matrix.
-##' @param graph An \code{\link[igraph]{igraph-package}} object. May be directed or weighted.
+##' @param graph An \code{\link[igraph:aaa-igraph-package]{igraph}} object. May be directed or weighted.
 ##' @param directed logical. Whether directed information is passed to the Laplacian matrix.
 ##' @keywords graph network igraph Laplacian
 ##' @importFrom igraph laplacian_matrix graph.edgelist as.undirected graph_from_adjacency_matrix

### 2020-07-01 R/plot_directed.R

@@ -13,9 +13,9 @@
 ##'
 ##' @title Extensions to igraph for Customising plots
 ##'
-##' @description Functions to plot_directed or graph structures including customised colours, layout, states, arrows. Uses graphs functions as an extension of \code{\link[igraph]{igraph-package}}. Designed for plotting directed graphs.
+##' @description Functions to plot_directed or graph structures including customised colours, layout, states, arrows. Uses graphs functions as an extension of \code{\link[igraph:aaa-igraph-package]{igraph}}. Designed for plotting directed graphs.
 ##'
-##' @param graph An \code{\link[igraph]{igraph-package}} object. Must be directed with known states.
+##' @param graph An \code{\link[igraph:aaa-igraph-package]{igraph}} object. Must be directed with known states.
 ##' @param state character or integer. Defaults to "activating" if no "state" edge attribute 
 ##' found. May be applied a scalar across all edges or as a vector for each edge respectively. 
 ##' Accepts non-integer values for weighted edges provided that the sign indicates whether links
@@ -25,7 +25,7 @@
 ##'  package \url{https://github.com/TomKellyGenetics/graphsim}. Vector input is supported 
 ##' @param labels character vector. For labels to plot nodes. Defaults to vertex names in 
 ##' graph object. Entering "" would yield unlabelled nodes.
-##' @param layout function. Layout function as selected from \code{\link[igraph]{layout_}}. 
+##' @param layout function. Layout function as selected from \code{\link[igraph:aaa-igraph-package]{layout_}}. 
 ##' Defaults to layout.fruchterman.reingold. Alternatives include layout.kamada.kawai, 
 ##' layout.reingold.tilford, layout.sugiyama, and layout.davidson.harel. A 2-column 
 ##' layout matrix giving x and y co-ordinates of each node can be given.

### 2020-07-01 R/state.R

@@ -8,7 +8,7 @@
 ##' in \code{\link[graphsim]{generate_expression}}. This resolves edge states to determine the sign
 ##' of all correlations between nodes in a network. These are computed interally for sigma matrices
 ##' as required.
-##' @param graph An \code{\link[igraph]{igraph-package}} object. May be directed or weighted as long as
+##' @param graph An \code{\link[igraph:aaa-igraph-package]{igraph}} object. May be directed or weighted as long as
 ##' a shortest path can be computed.
 ##' @param state numeric vector. Vector of length E(graph). Sign used to calculate state matrix,
 ##' may be an integer state or inferred directly from expected correlations for each edge. May be

### 2020-07-01 R/generate.R
correct links in documentation (link to file) https://github.com/r-lib/roxygen2/issues/707

@@ -4,22 +4,22 @@
 ##' @title Generate Simulated Expression
 ##'
 ##' @description Compute simulated continuous expression data from a graph 
-##' network structure. Requires an \code{\link[igraph]{igraph-package}} pathway 
+##' network structure. Requires an \code{\link[igraph:aaa-igraph-package]{igraph}} pathway 
 ##' structure and a matrix of states (1 for activating and -1 for 
 ##' inhibiting) for link signed correlations, from a vector of edge states 
 ##' to a signed adjacency matrix for use in 
 ##' \code{\link[graphsim]{generate_expression}}. 
 ##' Uses graph structure to pass a sigma covariance matrix from 
-##' \code{\link[graphsim]{make_sigma}} or 
-##' \code{\link[graphsim]{make_sigma}} on to 
-##' \code{\link[mvtnorm]{Mvnorm}}. By default data is generated with a mean of
+##' \code{\link[graphsim:make_sigma]{make_sigma_mat_graph}} or 
+##' \code{\link[graphsim:make_sigma]{make_sigma_mat_dist_graph}} on to 
+##' \code{\link[mvtnorm:Mvnorm]{rmvnorm}}. By default data is generated with a mean of
 ##'  0 and standard deviation of 1 for each gene (with correlations between 
 ##'  derived from the graph structure).
 ##'
 ##' @param n number of observations (simulated samples).
 ##' @param mat precomputed adjacency, laplacian, commonlink, or scaled 
 ##' distance matrix.
-##' @param graph An \code{\link[igraph]{igraph-package}} object. May must be 
+##' @param graph An \code{\link[igraph:aaa-igraph-package]{igraph}} object. May must be 
 ##' directed if states are used.
 ##' @param state numeric vector. Vector of length E(graph). Sign used
 ##' to calculate state matrix, may be an integer state or inferred directly

### 2020-07-01 R/sigma.R

@@ -3,8 +3,8 @@
 ##'
 ##' @title Generate Sigma (\eqn{\Sigma}) Matrix
 ##'
-##' @description Compute the Sigma (\eqn{\Sigma}) matrix from an \code{\link[igraph]{igraph-package}} structure 
-##' or pre-computed matrix. These are compatible with \code{\link[mvtnorm]{Mvnorm}} and
+##' @description Compute the Sigma (\eqn{\Sigma}) matrix from an \code{\link[igraph:aaa-igraph-package]{igraph}} structure 
+##' or pre-computed matrix. These are compatible with \code{\link[mvtnorm:Mvnorm]{rmvnorm}} and
 ##'  \code{\link[graphsim]{generate_expression}}.
 ##' By default data is generated with a mean of 0 and standard deviation of 1 for 
 ##' each gene (with correlations between derived from the graph structure).
@@ -13,7 +13,7 @@
 ##' between each gene in the output from \code{\link[graphsim]{generate_expression}}.
 ##'
 ##' @param mat precomputed adjacency, laplacian, commonlink, or scaled distance matrix.
-##' @param graph An \code{\link[igraph]{igraph-package}} object. May be directed or weighted.
+##' @param graph An \code{\link[igraph:aaa-igraph-package]{igraph}} object. May be directed or weighted.
 ##' @param state numeric vector. Vector of length E(graph). Sign used to calculate 
 ##' state matrix, may be an integer state or inferred directly from expected correlations
 ##' for each edge. May be applied a scalar across all edges or as a vector for each edge

### 2020-07-04 paper/paper.Rmd
update computational details https://github.com/openjournals/joss-reviews/issues/2161

@@
 
+Complete examples of code needed to produce the figures in this paper are
+available in the Rmarkdown version in the package GitHub repository
+ (\url{https://github.com/TomKellyGenetics/graphsim}).
+ Further details are available in the vignettes as well.
+
 The results in this paper were obtained using R 4.0.2 with the \texttt{igraph} 1.2.5
@@
 \url{https://CRAN.R-project.org}. The \texttt{graphsim} 1.0.0 package
-presented can be installed from CRAN and the issues can  be reported to
-the development version on GitHub (\url{https://github.com/TomKellyGenetics/graphsim}).
+can be installed from CRAN and the issues can  be reported to
+the development version on GitHub.
 This package is included in the \texttt{igraph.extensions} library on GitHub (\url{https://github.com/TomKellyGenetics/igraph.extensions})
@@
 compatible with installations on Windows, Mac, and Linux operating
-systems. The package GitHub repository also contains vignettes with more
-information and examples on running functions released in the package.
+systems. 
 Updates to the  package (\texttt{graphsim} 1.0.0) will be released on CRAN.
-
-Complete examples of code needed to produce the figures in this paper are
-available in the Rmarkdown version in the package GitHub repository
- (\url{https://github.com/TomKellyGenetics/graphsim/paper}).
- Further details are available in the vignettes as well.
 

### 2020-07-07 paper/paper.Rmd
correct grammar in paper

@@
 simulation of expression profiles that approximate the log-transformed
-and normalised data from microarray studies, as well as, bulk or single-cell RNA-Seq
+and normalised data from microarray studies, as well as bulk or single-cell RNA-Seq
 experiments. This procedure has been released as the \texttt{graphsim}

### 2020-07-09 paper/paper.Rmd
copyediting paper https://github.com/openjournals/joss-reviews/issues/2161

@@
 analysis. As a result, databases that describe biological pathways 
-are increasingly uesad to assist with the interpretation of results
+are increasingly used to assist with the interpretation of results
 from large-scale genomics studies. Incorporating information from 

### 2020-07-13 R/graphsim-package.R
clarify authorship and contributions (based on ggplot2 and dplyr)

@@ -9,8 +9,21 @@
 #' @aliases graphsim-package graphsim
 #' @docType package
 #' 
-#' @author \bold{Maintainer}:  Tom Kelly \email{tom.kelly@@riken.jp} (\href{https://orcid.org/0000-0003-3904-6690}{ORCID})
-#' @author Mik Black (\href{0000-0003-1174-6054}{ORCID})
+#' @author \bold{Maintainer}:  Tom Kelly  \email{tom.kelly@@riken.jp}
+#' 
+#' Authors:
+#' \itemize{
+#' \item Tom Kelly (RIKEN IMS) \href{https://orcid.org/0000-0003-3904-6690}{ORCID})
+#' \item Mik Black (Otago University) (\href{0000-0003-1174-6054}{ORCID})
+#' }
+#' 
+#' Reviewers:
+#' \itemize{
+#' \item Cory Brunson (UConn) (\href{0000-0003-3126-9494}{ORCID})
+#' \item Mik Black (Otago University) (\href{0000-0003-1174-6054}{ORCID})
+#' }
+#' 
+#' Editor: Mark Jensen (Frederick National Laboratory for Cancer Research)
 #' 
 #' @details This package provides functions to develop simulated continuous data 
 #' (e.g., gene expression) from a sigma covariance matrix derived from a 

### 2020-07-13 R/graphsim-package.R
add see also links to package help

@@ -100,4 +100,31 @@
 #'      journal = {Journal of Open Source Software} 
 #'    }
 #'  }
+#'  
+#' @seealso 
+#'  
+#'  Publication at \emph{Journal of Open Source Software}:
+#'  
+#'  \itemize{
+#'  \item \url{https://doi.org/10.21105/joss.02161}
+#'  }
+#'  
+#'  GitHub repository:
+#'  
+#'  \itemize{
+#'  \item \url{https://github.com/TomKellyGenetics/graphsim/}
+#'  }
+#'  
+#'  Report bugs:
+#'  
+#'  \itemize{
+#'  \item \url{https://github.com/TomKellyGenetics/graphsim/issues}
+#'  }
+#'  
+#'  Contributions:
+#'  
+#'  \itemize{
+#'  \item \url{https://github.com/TomKellyGenetics/graphsim/blob/master/CONTRIBUTING.md}
+#'  }
+#'  
 NULL

### 2020-07-13 R/graphsim-package.R
example of expression dataset

@@ -52,7 +52,49 @@
 #' distribution.
 #' 
 #' @section Getting Started:
-#' lorem ipsum (generate expression)
+#' The \code{\link[graphsim]{generate_expression}} function is a wrapper 
+#' around all necessary functions to give a final simulated dataset.
+#' 
+#' Here we set up an example graph object using the 
+#' \code{\link[igraph:aaa-igraph-package]{igraph}} package.
+#' 
+#' \preformatted{
+#' library("igraph")
+#' graph_structure_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"),c("D", "E"),
+#'                                c("D", "F"), c("F", "G"), c("F", "I"), c("H", "I"))
+#' graph_structure <- graph.edgelist(graph_structure_edges, directed = TRUE)
+#' }
+#' 
+#' Then we can call \code{\link[graphsim]{generate_expression}} to return
+#' the simulated data based on the relationships defined in the graph
+#' structure. Various options are available to fine-tune this.
+#' 
+#' \preformatted{
+#' expr <- generate_expression(100, graph_structure,
+#'                             cor = 0.8,
+#'                             mean = 0,
+#'                             sd = 1,
+#'                             comm = FALSE,
+#'                             dist = TRUE,
+#'                             absolute = FALSE,
+#'                             laplacian = FALSE)
+#' }
+#' 
+#' Here we can see the final result.
+#' 
+#' \preformatted{
+#' library("gplots")
+#' heatmap.2(expr,
+#'           scale = "none",
+#'           trace = "none",
+#'           col = bluered(50),
+#'           colsep = 1:4,
+#'           rowsep = 1:4)
+#' }
+#' 
+#' Input with an adjacency matrix is available using the
+#' \code{\link[graphsim:generate_expression]{generate_expression_mat}}
+#' function.
 #' 
 #' @section Creating Input Data:
 #' lorem ipsum (igraph / matrix functions / import from pkg data)

### 2020-07-13 R/graphsim-package.R
describe outputs

@@ -83,6 +83,8 @@
 #' Here we can see the final result.
 #' 
 #' \preformatted{
+#' dim(expr)
+#' 
 #' library("gplots")
 #' heatmap.2(expr,
 #'           scale = "none",
@@ -91,6 +93,9 @@
 #'           colsep = 1:4,
 #'           rowsep = 1:4)
 #' }
+#' 
+#' This dataset consists of 9 rows (one for each vertex or gene)
+#' in the graph and 100 columns (one for each sample or observation).
 #' 
 #' Input with an adjacency matrix is available using the
 #' \code{\link[graphsim:generate_expression]{generate_expression_mat}}

### 2020-07-14 R/graphsim-package.R
add section on plotting to package help

@@ -299,11 +299,52 @@
 #' }
 #' 
 #' @section Examining Step-by-Step:
+#'
+#'
 #' lorem ipsum (plotting sigma and \code{cor(t(expr))})
 #' 
 #' @section Visualization:
-#' lorem ipsum (plot_directed)
+#'
+#' The \sQuote{graphsim} package comes with a built-in plotting
+#' function to display graph objects. 
 #' 
+#' \preformatted{
+#' graph_structure_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"),c("D", "E"),
+#'                                c("D", "F"), c("F", "G"), c("F", "I"), c("H", "I"))
+#' graph_structure <- graph.edgelist(graph_structure_edges, directed = TRUE)
+#' plot_directed(graph_structure, layout = layout.kamada.kawai)
+#' }
+#'
+#' This supports the \sQuote{state} parameter to display
+#' activating relationships (with positive correlations)
+#' and inhibiting or repressive relationships (with
+#' negative correlations).
+#'
+#' \preformatted{
+#' edge_states <- c(1, 1, -1, -1, 1, -1, 1, -1)
+#' graph_structure <- graph.edgelist(graph_structure_edges, directed = TRUE)
+#' plot_directed(graph_structure, state = edge_states,
+#'               col.arrow = c("darkgreen", "red")[edge_states / 2 + 1.5]
+#'               layout = layout.kamada.kawai)
+#' }
+#'
+#' These states can also be passed from the \sQuote{state} edge
+#' attribute of the graph object.
+#'
+#' \preformatted{
+#" library("scales")
+#' graph_pathway <- identity(TGFBeta_Smad_graph)
+#' edge_properties <- E(graph_pathway)$state
+#' plot_directed(graph_pathway,
+#'               col.arrow = c(alpha("navyblue", 0.25),
+#'                             alpha("red", 0.25))[edge_properties],
+#'               fill.node = c("lightblue"),
+#'               layout = layout.kamada.kawai)
+#' }
+#' 
+#' This plotting function is demonstrated in more detail
+#' in the \link[../html/plots_directed.html]{plotting} vignette.
+#'
 #' @section Further information:
 #'   The graphsim package is published in the \emph{Journal of Open Source Software}.
 #'   See the paper here for more details:

### 2020-07-14 R/graphsim-package.R
add step-by-step demonstration of computing with internal functions

@@ -26,7 +26,7 @@
 #' Editor: Mark Jensen (Frederick National Laboratory for Cancer Research)
 #' 
 #' @details This package provides functions to develop simulated continuous data 
-#' (e.g., gene expression) from a sigma covariance matrix derived from a 
+#' (e.g., gene expression) from a Sigma (\eqn{\Sigma}) covariance matrix derived from a 
 #' graph structure in \sQuote{\code{igraph}} objects. Intended to extend 
 #' \sQuote{\code{mvtnorm}} to take 'igraph' structures rather than sigma 
 #' matrices as input. This allows the use of simulated data that correctly
@@ -151,7 +151,7 @@
 #' } 
 #' 
 #' The following functions will compute matrices from an
-#' \code{\link[graphsim:make_adjmatrix]{adjcacency matrix}}:
+#' \code{\link[graphsim:make_adjmatrix]{adjacency matrix}}:
 #' 
 #' \itemize{
 #' 
@@ -300,9 +300,150 @@
 #' 
 #' @section Examining Step-by-Step:
 #'
-#'
-#' lorem ipsum (plotting sigma and \code{cor(t(expr))})
+#' These internal functions can be called to compute steps of
+#' the simulation procedure and examine the results.
 #' 
+#' 1. first we create a graph structure and define the input parameters
+#' 
+#' \preformatted{
+#' library("igraph")
+#' graph_structure_edges <- rbind(c("A", "C"), c("B", "C"), c("C", "D"),c("D", "E"),
+#'                                c("D", "F"), c("F", "G"), c("F", "I"), c("H", "I"))
+#' graph_structure <- graph.edgelist(graph_structure_edges, directed = TRUE)
+#' #sample size
+#' data.n <- 100
+#' #data distributions
+#' data.cor <- 0.75
+#' data.mean <- 3
+#' data.sd <- 1.5
+#' #inhibition states
+#' edge_states <- c(1, 1, -1, -1, 1, 1, 1, 1)
+#' }
+#' 
+#' 2. examine the relationships between the genes.
+#' 
+#' Here we can see which nodes share an edge:
+#' 
+#' \preformatted{
+#' > adjacency_matrix <- make_adjmatrix_graph(graph_structure)
[…]

### 2020-07-14 R/graphsim-package.R
complete package help page #6

@@ -444,7 +444,45 @@
 #' These functions are demonstrated in more detail
 #' in the \link[../html/simulate_expression.html]{main} vignette.
 #'  
-#' @section Visualization:
+#' @section Data Visualization:
+#' 
+#' Heatmaps can be used from the \code{\link[gplots::heatmap.2]{gplots}}
+#' package to display these simulated datasets.
+#' 
+#' \preformatted{
+#' library("gplots")
+#' heatmap.2(adjacency_matrix, scale = "none", trace = "none",
+#'           col = colorpanel(50, "white", "black"), key = FALSE)
+#'           
+#' heatmap.2(relationship_matrix, scale = "none", trace = "none",
+#'           col = colorpanel(50, "white", "red"))
+#'           
+#' heatmap.2(state_matrix, scale = "none", trace = "none",
+#'           col = colorpanel(50, "royalblue", "palevioletred"),
+#'           colsep = 1:length(V(graph_structure)),
+#'           rowsep = 1:length(V(graph_structure)))
+#' 
+#' heatmap.2(sigma_matrix, scale = "none", trace = "none",
+#'           col = colorpanel(50, "royalblue", "white", "palevioletred"),
+#'           colsep = 1:length(V(graph_structure)),
+#'           rowsep = 1:length(V(graph_structure)))
+#'           
+#' heatmap.2(expression_data, scale = "none", trace = "none",
+#'           col = colorpanel(50, "royalblue", "white", "palevioletred"),
+#'           colsep = 1:length(V(graph_structure)),
+#'          rowsep = 1:length(V(graph_structure)))
+#' 
+#' heatmap.2(cor_data, scale = "none", trace = "none",
+#'           col = colorpanel(50, "royalblue", "white", "palevioletred"),
+#'           colsep = 1:length(V(graph_structure)),
+#'           rowsep = 1:length(V(graph_structure)))
+#' }
+#' 
+#' In particular we can see here that the expected correlations
+#' show by the \code{sigma_matrix} are similar to the observed
+#' correlations in the \code{cor_data}.
+#'  
+#' @section Graph Visualization:
 #'
 #' The \sQuote{graphsim} package comes with a built-in plotting
 #' function to display graph objects. 

### 2020-07-17 R/graphsim-package.R
corrections for CRAN checks

@@ -14,13 +14,13 @@
 #' Authors:
 #' \itemize{
 #' \item Tom Kelly (RIKEN IMS) \href{https://orcid.org/0000-0003-3904-6690}{ORCID})
-#' \item Mik Black (Otago University) (\href{0000-0003-1174-6054}{ORCID})
+#' \item Mik Black (Otago University) (\href{https://orcid.org/0000-0003-1174-6054}{ORCID})
 #' }
 #' 
 #' Reviewers:
 #' \itemize{
-#' \item Cory Brunson (UConn) (\href{0000-0003-3126-9494}{ORCID})
-#' \item Robrecht Cannoodt (Ghent University) (\href{0000-0003-3641-729X}{ORCID})
+#' \item Cory Brunson (UConn) (\href{https://orcid.org/0000-0003-3126-9494}{ORCID})
+#' \item Robrecht Cannoodt (Ghent University) (\href{https://orcid.org/0000-0003-3641-729X}{ORCID})
 #' }
 #' 
 #' Editor: Mark Jensen (Frederick National Laboratory for Cancer Research)
@@ -442,11 +442,11 @@
 #' }
 #' 
 #' These functions are demonstrated in more detail
-#' in the \link[../html/simulate_expression.html]{main} vignette.
+#' in the \href{https://cran.r-project.org/web/packages/graphsim/vignettes/simulate_expression.html}{main} vignette.
 #'  
 #' @section Data Visualization:
 #' 
-#' Heatmaps can be used from the \code{\link[gplots::heatmap.2]{gplots}}
+#' Heatmaps can be used from the \code{\link[gplots:heatmap.2]{gplots}}
 #' package to display these simulated datasets.
 #' 
 #' \preformatted{
@@ -522,7 +522,7 @@
 #' }
 #' 
 #' This plotting function is demonstrated in more detail
-#' in the \link[../html/plots_directed.html]{plotting} vignette.
+#' in the \href{https://cran.r-project.org/web/packages/graphsim/vignettes/plots_directed.html}{plotting} vignette.
 #'
 #' @section Further information:
 #'   The graphsim package is published in the \emph{Journal of Open Source Software}.

### 2021-07-27 R/graphsim-package.R
CRAN release 1.0.2

@@ -547,7 +547,6 @@
 #'   A BibTeX entry for LaTeX users is: \preformatted{
 #'   @article{Kelly2020joss02161,
 #'      doi = {10.21105/joss.02161},
-#'      url ={https://doi.org/10.21105/joss.02161},
 #'      year = {2020},
 #'      publisher = {The Open Journal},
 #'      volume = {5},

### 2022-09-10 R/graphsim-package.R
update links in documentation

@@ -527,7 +527,7 @@
 #' @section Further information:
 #'   The graphsim package is published in the \emph{Journal of Open Source Software}.
 #'   See the paper here for more details:
-#'    \href{https://doi.org/10.21105/joss.02161}{doi: 10.21105/joss.02161}
+#'    \doi{10.21105/joss.02161}
 #'   
 #'   The graphsim GitHub repository is here:
 #'   \href{https://github.com/TomKellyGenetics/graphsim}{TomKellyGenetics/graphsim}
@@ -542,7 +542,7 @@
 #'   Kelly, S.T. and Black, M.A. (2020). graphsim: An R package for simulating gene
 #'   expression data from graph structures of biological pathways.
 #'   \emph{Journal of Open Source Software}, \bold{5}(51), 2161,
-#'   \url{https://doi.org/10.21105/joss.02161}
+#'   \doi{10.21105/joss.02161}
 #'   
 #'   A BibTeX entry for LaTeX users is: \preformatted{
 #'   @article{Kelly2020joss02161,
@@ -563,7 +563,7 @@
 #'  Publication at \emph{Journal of Open Source Software}:
 #'  
 #'  \itemize{
-#'  \item \url{https://doi.org/10.21105/joss.02161}
+#'  \item \doi{10.21105/joss.02161}
 #'  }
 #'  
 #'  GitHub repository:

### 2022-09-10 R/graphsim-package.R
CRAN release 1.0.3

@@ -522,7 +522,7 @@
 #' }
 #' 
 #' This plotting function is demonstrated in more detail
-#' in the \href{https://CRAN.R-project.org/package=graphsim/vignettes/plots_directed.html}{plotting} vignette.
+#' in the plots_directed.Rmd plotting vignette.
 #'
 #' @section Further information:
 #'   The graphsim package is published in the \emph{Journal of Open Source Software}.

### 2022-09-10 R/laplacian.R

@@ -13,8 +13,6 @@
 ##' @keywords graph network igraph Laplacian
 ##' @importFrom igraph laplacian_matrix graph.edgelist as.undirected graph_from_adjacency_matrix
 ##' @import igraph
-##' @importFrom Matrix Matrix
-##' @importClassesFrom Matrix dgCMatrix
 ##' 
 ##' @family graphsim functions
 ##' @family graph conversion functions
